### 2019-08-06 manuscript/template.tex
begin intro

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}		% Can be removed after putting your text content

\title{Towards Explainability of Machine Learning Models in Insurance Pricing}

%\date{September 9, 1985}	% Here you can change the date presented in the paper title
%\date{} 					% Or removing it

\author{
 Kevin Kuo \\
  \texttt{kevin@kasa.ai} \\
  %% examples of more authors
   \And
 Daniel Lupton \\
  \texttt{danlupton@gmail.com} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

\begin{document}
\maketitle

\begin{abstract}
Abstract goes here
\end{abstract}

\section{Introduction}
Risk classification for insurance rating has traditionally been done with one-way, or univariate, analysis techniques. In recent years, many insurers have moved towards using generalized linear models (GLM), a multivariate predictive modeling technique, which addresses many shortcomings of univariate approaches, and is currently considered the state of the art in insurance risk classification. At the same time, machine learning (ML) techniques such as random forests and deep neural networks have gained popularity in many industries due to their superior predictive performance over linear models. However, these ML techniques, often considered to be completely “black box”, have been less successful in gaining adoption in insurance pricing, which is a regulated discipline and requires a certain amount of transparency in models.

\end{document}

### 2019-08-28 manuscript/manuscript.Rmd

--- 2019-08-06 manuscript/template.tex
@@
-\documentclass{article}
+---
+title: |
+ Towards Explainability of Machine Learning Models in Insurance Pricing
+type: ARTICLE TEMPLATE
+author:
+  - name: Author One
+    affil: a
+    email: foo@bar.foo
+  - name: Author Two
+    affil: b
+    email: bar@foo.bar
+affiliation:
+  - num: a
+    address: |
+      Foo
+  - num: b
+    address: |
+      Bar
+bibliography: interactcadsample.bib
+appendix: appendix.tex
+abstract: |
+  Abstract tbd
+keywords: |
+  actuarial science; general insurance
+header-includes: |
+  \usepackage{hyperref}
+  \usepackage[utf8]{inputenc}
+  \def\tightlist{}
+output: rticles::tf_article
+---
 
-\usepackage{arxiv}
+# Introduction
 
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{lmodern}
-\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{lipsum}		% Can be removed after putting your text content
-
-\title{Towards Explainability of Machine Learning Models in Insurance Pricing}
-
-%\date{September 9, 1985}	% Here you can change the date presented in the paper title
-%\date{} 					% Or removing it
-
-\author{
- Kevin Kuo \\
-  \texttt{kevin@kasa.ai} \\
-  %% examples of more authors
-   \And
- Daniel Lupton \\
-  \texttt{danlupton@gmail.com} \\
-  %% \AND
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-}
-
-\begin{document}
-\maketitle
-
-\begin{abstract}
-Abstract goes here
-\end{abstract}
-
-\section{Introduction}
 Risk classification for insurance rating has traditionally been done with one-way, or univariate, analysis techniques.
@@
 However, these ML techniques, often considered to be completely “black box”, have been less successful in gaining adoption in insurance pricing, which is a regulated discipline and requires a certain amount of transparency in models.
-
-\end{document}

### 2019-08-29 manuscript/manuscript.Rmd
article outline

@@
  Towards Explainability of Machine Learning Models in Insurance Pricing
-type: ARTICLE TEMPLATE
+type: WIP DRAFT 
 author:
@@
       Bar
-bibliography: interactcadsample.bib
+bibliography: explain-pricing.bib
 appendix: appendix.tex
@@
 
-Risk classification for insurance rating has traditionally been done with one-way, or univariate, analysis techniques.
+Risk classification for property & casualty (P&C) insurance rating has traditionally been done with one-way, or univariate, analysis techniques.
 In recent years, many insurers have moved towards using generalized linear models (GLM), a multivariate predictive modeling technique, which addresses many shortcomings of univariate approaches, and is currently considered the state of the art in insurance risk classification.
-At the same time, machine learning (ML) techniques such as random forests and deep neural networks have gained popularity in many industries due to their superior predictive performance over linear models.
-However, these ML techniques, often considered to be completely “black box”, have been less successful in gaining adoption in insurance pricing, which is a regulated discipline and requires a certain amount of transparency in models.
+At the same time, machine learning (ML) techniques such as deep neural networks have gained popularity in many industries due to their superior predictive performance over linear models [@lecunDeepLearning2015].
+In fact, there is a fast growing body of literatuer on applying ML to P&C reserving [@kuoDeepTriangleDeep2018; @wuthrichMachineLearning2018; @gabrielliNeuralNetwork2019a; @gabrielliNeuralNetwork2019].
+However, these ML techniques, often considered to be completely “black box”, have been less successful in gaining adoption in pricing, which is a regulated discipline and requires a certain amount of transparency in models.
+
+If insurers can gain more insight into how ML models behave in risk classification contexts, it would increase their ability to reassure regulators and the public that accepted ratemaking principles are met.
+Being able to charge more accurate premiums would, in turn, make the risk transfer system more efficient and contribute to the betterment of society.
+In this paper, we aim to take a step towards liberating actuaries from the confines of linear models in pricing projects, by proposing a framework for explaining ML models for ratemaking that regulators, practictioners, and researchers in actuarial science can build upon.
+
+# Property and Casualty Ratemaking
+
+(overview of p/c ratemaking: what it is, maybe a bit of history of getting from univariate methods to GLMs today, point out that in this paper we're really dealing with the risk classification aspect)
+
+# Interpretability in the Ratemaking Context
+
+(literature review, settle on a definition to work with, e.g. @doshi-velezRigorousScience2017, identify maybe 3 questions about a model to answer, tie to principles on p/c ratemaking)
+
+# Applying Model Interpretation Techniques
+
+(some definitions, e.g. global/model vs. local/instance, categorize questions accordingly)
+
+## (answer each question)
+
+(for each question, propose a technique, mention alternative techniques, pros/cons, implement, interpret results)
+
+# Conclusion
+
+(conclude)

### 2019-08-30 manuscript/manuscript.Rmd

@@
 
-# Property and Casualty Ratemaking
+The rest of this paper is organized as follows: Section \ref{ratemaking} provides an overview of P&C ratemaking, Section \ref{interpretability} discusses model interpretability in the context of ratemaking and proposes specific tasks for model explanation, Section \ref{application} describes current model interpretation techniques and applies them to the tasks defined in the previous section, and Section \ref{conclusion} concludes.
+
+# Property and Casualty Ratemaking {#ratemaking}
 
@@
 
-# Interpretability in the Ratemaking Context
+# Interpretability in the Ratemaking Context {#interpretability}
 
@@
 
-# Applying Model Interpretation Techniques
+# Applying Model Interpretation Techniques {#application}
 
@@
 
-# Conclusion
+# Conclusion {#conclusion}
 

### 2019-08-30 manuscript/manuscript.Rmd
sota -> gold standard to close #10

--- 2019-08-29 manuscript/manuscript.Rmd
@@
 Risk classification for property & casualty (P&C) insurance rating has traditionally been done with one-way, or univariate, analysis techniques.
-In recent years, many insurers have moved towards using generalized linear models (GLM), a multivariate predictive modeling technique, which addresses many shortcomings of univariate approaches, and is currently considered the state of the art in insurance risk classification.
+In recent years, many insurers have moved towards using generalized linear models (GLM), a multivariate predictive modeling technique, which addresses many shortcomings of univariate approaches, and is currently considered the gold standard in insurance risk classification.
 At the same time, machine learning (ML) techniques such as deep neural networks have gained popularity in many industries due to their superior predictive performance over linear models [@lecunDeepLearning2015].
@@
 In this paper, we aim to take a step towards liberating actuaries from the confines of linear models in pricing projects, by proposing a framework for explaining ML models for ratemaking that regulators, practictioners, and researchers in actuarial science can build upon.
+
+(add outline for rest of paper)
 

### 2019-08-30 manuscript/manuscript.Rmd
typos

@@
 At the same time, machine learning (ML) techniques such as deep neural networks have gained popularity in many industries due to their superior predictive performance over linear models [@lecunDeepLearning2015].
-In fact, there is a fast growing body of literatuer on applying ML to P&C reserving [@kuoDeepTriangleDeep2018; @wuthrichMachineLearning2018; @gabrielliNeuralNetwork2019a; @gabrielliNeuralNetwork2019].
+In fact, there is a fast growing body of literature on applying ML to P&C reserving [@kuoDeepTriangleDeep2018; @wuthrichMachineLearning2018; @gabrielliNeuralNetwork2019a; @gabrielliNeuralNetwork2019].
 However, these ML techniques, often considered to be completely “black box”, have been less successful in gaining adoption in pricing, which is a regulated discipline and requires a certain amount of transparency in models.
@@
 Being able to charge more accurate premiums would, in turn, make the risk transfer system more efficient and contribute to the betterment of society.
-In this paper, we aim to take a step towards liberating actuaries from the confines of linear models in pricing projects, by proposing a framework for explaining ML models for ratemaking that regulators, practictioners, and researchers in actuarial science can build upon.
+In this paper, we aim to take a step towards liberating actuaries from the confines of linear models in pricing projects, by proposing a framework for explaining ML models for ratemaking that regulators, practitioners, and researchers in actuarial science can build upon.
 

### 2019-09-27 manuscript/manuscript.Rmd

@@
 
+-First GLM paper to appear on the syllabus was by Anderson, et al. in 2006?
+-First ML paper... tricky. "cluster analysis" first mentioned in 2011. I think that was the Robertson paper, and he used k-means to cluster the hazard groups for NCCI?
+-First real time ML appears as such is 2018 with the MAS-I and MAS-II exams
+-Machine learning algorithms in a ratemaking context are a very recent phenomenon
+-Brief hand-wavey history of machine learning
+
 # Interpretability in the Ratemaking Context {#interpretability}
@@
 (literature review, settle on a definition to work with, e.g. @doshi-velezRigorousScience2017, identify maybe 3 questions about a model to answer, tie to principles on p/c ratemaking)
+
+Within the actuarial profession, Actuarial Standard of Practice 41 ("Actuarial Communications") notes that "...another actuary qualified in the same practice area [should be able to] make an objective appraisal of the reasonableness of the actuary's work as presented in the actuarial report." (cite: ASOP 41) Underlying this requirement is an assumption that the hypothetical other actuary qualified in the same practice area is adequately familiar with the relevant techniques employed.
+Although the syllabus of basic education is constantly changing, there has at times been an assumption that all techniques and assumptions that have ever been a part of the syllabus of basic education needn't be explained from first principles in general actuarial communications, and that an actuary practicing in the same field should be able to make an objective appraisal of the results from the methods found in the syllabus.
+[This can be supported by reference to the edits to ASOP 38 over time - originally it was designed to be about all models, but they revised it to be about only models that incorporate specialized knowledge outside of the actuary's area of expertise...
+do we need this citation, though?] This is notable because, beginning with the introduction of the MAS-I and MAS-II examination in July of 2018, several machine learning models were formally included in the syllabus of basic education.
+These exams cover a wide range of topics, such as splines, clustering algorithms, decision trees, boosting, and principle components analysis.
+(cite: 2018 syllabus of basic education)
+
+Nevertheless, machine learning poses something of a special challenge for ASOP 41 for several reasons.
+Machine learning models can be very ad hoc compared to traditional statistical models.
+Because many machine learning models are deterministic, they may not admit of standard metrics for model comparison (e.g., it's not straightforward to calculate an AIC over a neural network).
+In addition, machine learning methods are often combined into ensembles that may not be easily separated and that may, as a collection, cease to resemble a single standard version of a model.
+Complicating matters still further, machine learning models can be "black boxes" insofar as the final form of response curve cannot be easily predicted and may depend heavily on the available data (which may not, in turn, be available to the reviewer).
+
+This last item raises a final interesting issue.
+Generalized linear models and their ilk are often fitted using one of a handful of standard and well-understood approaches (e.g., maximum likelihood estimation).
+However, this is not possible in general with machine learning models, as machine learning algorithms often use loss surfaces that are very complex such that it may not be feasible to calculate the global minimum of the surface.
+Certainly, closed form representations of the loss surfaces are not generally available.
+For this reason, the training phase of a machine learning model is, in many ways, just as important to one's understanding as the model form and the data on which the model is fitted.
+Because the final model result is inseparable from these three components (training method, model form, and data), it is not generally adequate to just know the method employed to make an objective appraisal of the reasonableness of the model.
+More information is necessary.
+
+Of course, these comments only apply to the actuarial profession.
+Outside of the actuarial profession, communication of results may be more challenging.
+A survey conducted by the Casualty Actuarial and Statistical Task Force of the National Association of Insurance Commissioners found that ___
+
+-NAIC surveys indicate challenges for regulators in understanding even GLMs
+
+In response to a similar survey in Sping 2017, 33 state regulators noted that it would be helpful or very helpful for the NAIC to develop information and tools to assist in reviewing rate filings based on GLMs, and 34 noted that it would be helpful to develop similar items to assist in reviewing "Other Advanced Modeling Techniques." [cite: Spring Vol I Proceedings of the NAIC]
+
+One outgrowth of this need was the development of a white paper on best practices for regulatory review of predictive models.
+The white paper focuses on review of GLMs, particularly with respect to private passenger automobile and homeowners' insurance.
+Some of the guidance offered in this regard is therefore not strictly applicable to the review of machine learning models.
+For example, as previously noted, p-values are not a concept that translates well to deterministic machine learning algorithms.
+However, among the guidance applicable to machine learning algorithms are the following (paraphrasing):
+* understand the relationship between the inputs and the expected loss or expense differences in risk,
+* determine that these input characteristics are not unfairly discriminatory, and
+* determine the extent of premium disruption among policyholders and be able to explain the source of premium disruptions. [cite: NAIC white paper]
+
+Note that this list is not exhaustive of the guidance offered by any means, but that these three items are those that are applicable that may present greater challenge for machine learning algorithms compared to GLMs.
+[By comparison, guidance to understand that the variables are predictive of loss is a common goal of GLMs and ML, and in fact ML kind of gets it "for free" since it's about minimizing prediction error rather than maximizing likelihood.
+Not sure if this bears mentioning.
+Could also note the extent to which the NAIC guidance focuses on predictiveness as opposed to p-values.
+I think it's pretty questionable.]
+
+Depending on the model, machine learning algorithms can be non-linear in the inputs.
+In areas where data are sparse, the model could generate unnatural response curves that could go undetected by a very high-level view of the model results.
+This leads to a situation where lift charts or tests on holdout data may indicate that the model is performing adequately, but it may be difficult to explain the reasons for a particular premium indicaton or a change in premium when policy characteristics change.
+
+-For this, we recommend ____ (grid search?)
+-Because the model form may be difficult to represent in the form of a single equation (and harder still to evaluate), one method of evaluating the reasonableness of model results is local approximations of the impact of different rating variables (LIME?)
 

### 2019-09-27 manuscript/manuscript.Rmd
initial changes

@@
 
-(overview of p/c ratemaking: what it is, maybe a bit of history of getting from univariate methods to GLMs today, point out that in this paper we're really dealing with the risk classification aspect)
+## History of Ratemaking
 
--First GLM paper to appear on the syllabus was by Anderson, et al. in 2006?
--First ML paper... tricky. "cluster analysis" first mentioned in 2011. I think that was the Robertson paper, and he used k-means to cluster the hazard groups for NCCI?
--First real time ML appears as such is 2018 with the MAS-I and MAS-II exams
--Machine learning algorithms in a ratemaking context are a very recent phenomenon
--Brief hand-wavey history of machine learning
+Early classification ratemaking procedures were typically univariate in nature.
+For example, [@lange_1966] notes that (at that time) most major lines of insurance used univariate methods based around the same principle: distributing an overall indication to territorial relativities or classification relativities based on the extent to which they deviated from the average experience.
+
+[@bailey_simon_1960] introduced minimum bias methods, which were expanded throughout the 60s, 70s, and 80s.
+As computing power developed, minimum bias began to give away to generalized linear models, with papers such as [@brown_1988] and [@mildenhall_1999] bridging the gap between the methods.
+
+Arguably, generalized linear models predate minimum bias procedures by a significant margin.
+The term "Generalized Linear Model" was coined by [@nelder_wedderburn_1972], but generalizations of least squares linear regression date back at least to the 1930s.
+Like minimum bias methods, GLMs did not become mainstream in actuarial science for some time.
+For example, the syllabus of basic education does not seem to include any mention of GLMs prior to [@brown_1988] in the 1990 syllabus for basic education.
+From there, GLMs seem to have received only passing mention until 2006 with the introduction of [@anderson_2005] to the syllabus.
+
+## Machine Learning in Ratemaking
+Paralelling the development of generalized linear models was the development of machine learning algorithms throughout the middle part of the 20th century.
+Detailed histories of machine learning may be found in sources such as [@Nilsson_2009] and [@wang_raj_2017].
+Consistent with GLMs, machine learning was relatively unpopular in actuarial science until the last ten years as computing power has become cheaper and more easily available and as machine learning software packages have obviated the need for developing analyses from scratch each time an analysis is performed.
+Due to the breadth of machine learning as a field, it is difficult to identify the first time it entered the CAS syllabus; however, cluster analysis (in the form of k-means) seems to have been first included in 2011 with [@robertson_2009].
+More recently, the MAS-I and MAS-II exams introduced in 2018 have included machine learning explicitly.
+
+Within the area of ratemaking, machine learning is still in its infancy.
+A significant portion of machine learning applications to ratemaking has been in the context of automobile telematics, such as [@gao_2018], [@gao_2018_2], [@gao_2019], [@roel_2018], or [@wuthrich_2017].
+Presumably this focus has been a result of the high-dimensionality and complexity of telematics data, making it a field in which the unique abilities of machine learning techniques give a clear advantage over traditional approaches.
+
+Outside of telematics, [@yang_2018] uses a gradient tree-boosting approach to capture non-linearities that would be a challenge for GLMs.
+[@henckaerts_2018] makes use of "generalized additive models" to improve predictions of GLMs.
+Many researchers, in an apparent effort to demonstrate the range of possibilities and advantages of machine learning, have approached the topic by comparing many different machine learning algorithms within a single study, such as in [@dugas_2003], [@noll_2018], [@spedicato_2018].
+These studies make use of such varied techniques as regression trees, boosting machines, support vector machines, and neural networks.
+
+## Ratemaking Process
+
+Regardless of the method employed for determining this risk of various classifications, the actual process of setting rate relativities typically involves some variation of the following steps:
+
+1. Obtain relevant policy-level data
+2. Prepare data for analysis
+3. Perform analysis on the data, employing desired method or methods to estimate needed rate relativites
+4. Select final rate relativities based on rate indications
+5. Present rates to the reglator, including explanation of the steps followed to derive the rates
+6. Answer questions from regulators regarding the method employed
+
+The focus of this paper is on steps 5 and 6.
+In particular, rate regulators are concerned with whether rates are inadequate, excessive, or unfairly discriminatory.
+In many states, rate filings that exceed certain thresholds for magnitude of rate changes or filings that make use of new or sophisticated predictive models may be subject to intense regulatory scrutiny.
+In these cases, it is necessary to be able to explain the results of the modeling process in a way that is understandable without sacrificing statistical rigor.
+
+It should be noted that communicating results is not simply a method of passing regulatory muster.
+Generating interpretable modeling output is an important - even essential - facet of model checking.
+Therefore, the techniques discussed in this paper may be viewed from the lens of providing useful information to regulators, but they should also be considered as part of a thorough vetting of any rating model.
 
@@
 
-(literature review, settle on a definition to work with, e.g. @doshi-velezRigorousScience2017, identify maybe 3 questions about a model to answer, tie to principles on p/c ratemaking)
+In this section, we attempt to develop a working definition of interpretability for ratemaking applications.
+While we will not provide a comprehensive survey of the prolific and fast evolving ML interpretability literature, we draw from it as appropriate in setting the stage for our discussion.
+There is not a standard definition of interpretability with consensus from all researchers; here are a few from frequently cited papers:
 
-Within the actuarial profession, Actuarial Standard of Practice 41 ("Actuarial Communications") notes that "...another actuary qualified in the same practice area [should be able to] make an objective appraisal of the reasonableness of the actuary's work as presented in the actuarial report." (cite: ASOP 41) Underlying this requirement is an assumption that the hypothetical other actuary qualified in the same practice area is adequately familiar with the relevant techniques employed.
+- Ability to explain or to present in understandable terms to a human [@doshi-velezRigorousScience2017];
+- The degree to which an observer can understand the cause of a decision [@biranExplanationJustification2017]; and
+- A method is interpretable if a user can correctly and efficiently predict the method’s results [@kimExamplesAre2016].
+
+We motivate our discussion by considering several aspects of interpretability.
+For the time being, one can take any of the definitions cited above.
+As we proceed through the points below, we aim to arrive at a more scoped and relevant definition of what it means to attain interpretability in a pricing model.
+(TODO: add a sentence mapping out what we'll talk about in this section)
+
+## Not All Linear Models are Interpretable
+
+We look at some examples of models that are considered by many to be interpretable.
+The GLM is probably the most oft-cited example of such a model (or rather, modeling framework) in the actuarial literature.
+Given a set of inputs, we can easily reason about what the output of the model is.
+As an illustrative example, we may have a claims severity model with age, amount of insurance (AOI), and gender as features; assuming a log link function, we have
+
+\begin{equation}
+  \log(E(\text{severity})) = \beta_0 + \beta_1 \cdot \text{age} + \beta_2 \cdot \text{AOI} + \beta_3 \cdot \text{gender}_{\text{male}}.
+\end{equation}
+
+Here, we can tell, for example, what the model would predict for the expected severity if we were to increase age by a certain amount, *all else being equal*, because the relationship between the feature and the response is simply the coefficient $\beta_1$ and the link function.
+Another commonly cited example of an interpretable model is a decision tree.
+An illustrative example is shown in (TODO: FIGURE).
+Here, the prediction is arrived at by following a sequence of if-else decisions.
+
+Now, it is worth pointing out that, when declaring that GLM or decision trees are interpretable models, we are implicitly qualifying that we are considering a reasonable number of linear features.
+In fact, the ease at which we can reason about a model rapidly declines as the number of features and interactions increase, as in the following (somewhat pathological) example:
+
+\begin{equation}\label{eq:badglm}
+  \log(E(\text{severity})) = \beta_0 + \beta_1 \cdot \text{AOI}  + \beta_2\cdot\text{age} + \beta_3\cdot\text{age}^2 + \beta_4 \cdot \text{age:AOI} + \beta_5 \cdot \text{gender}_{\text{male}}.
+\end{equation}
+
+Similarly, one can see that in (TODO: FIGURE), large and deep trees are tough to reason about.
+In other words, even when working within the framework of an "interpretable" class of models, we may still end up with something that many would consider "black box."
+
+## The Machinery is Not a Secret
+
+An occasional misconception is that we have no visibility into how some ML models *compute* predictions, which render them uninterpretable.
+Outside of proprietary algorithms, all common ML models, including neural networks, gradient boosted trees, and random forests, are well studied and have large bodies of literature documenting their inner workings.
+As an example, a fitted neural network model on tabular data is simply a composition of linear transformations followed with nonlinear activation functions.
+As in Equation \ref{eq:badglm}, one can write down the mathematical equation for calculating the prediction given some inputs, but it may be difficult for a human to reason about it.
+This is a somewhat philosophical point, as we show later in the paper that we can still provide explanations of completely "black box" models, but is important to note given that pricing models undergo regulatory scrutiny.
+
+## Explanations are Social
+
+Proposed by @hiltonConversationalProcesses1990 and interpreted by @millerExplanationArtificial2017 in the context of ML, model explanations are conversations or social interactions.
+One consequence of this identification is that explanations need to be *relevant* to the *audience*.
+In developing, filing, and operationalizing a pricing model, one needs to accommodate a variety of stakeholders, each of whom has a different set of questions, assumptions, and technical capacity.
+First, there are internal stakeholders at the company, which includes management and underwriters.
+While some of the individuals in this audience may be technical, they are likely less familiar with predictive modeling techniques than the actuaries and data scientists who build the models.
+(TODO: what's a representative question here?) Next, we have the regulators, who may have limited resources to review the models, and will focus on a specific list of questions.
+
+Within the actuarial profession, Actuarial Standard of Practice 41 ("Actuarial Communications") notes that "...another actuary qualified in the same practice area [should be able to] make an objective appraisal of the reasonableness of the actuary's work as presented in the actuarial report." [@asop_41] Underlying this requirement is an assumption that the hypothetical other actuary qualified in the same practice area is adequately familiar with the relevant techniques employed.
 Although the syllabus of basic education is constantly changing, there has at times been an assumption that all techniques and assumptions that have ever been a part of the syllabus of basic education needn't be explained from first principles in general actuarial communications, and that an actuary practicing in the same field should be able to make an objective appraisal of the results from the methods found in the syllabus.
@@
 These exams cover a wide range of topics, such as splines, clustering algorithms, decision trees, boosting, and principle components analysis.
-(cite: 2018 syllabus of basic education)
+[@cas_syllabus_2018]
 
@@
 Outside of the actuarial profession, communication of results may be more challenging.
-A survey conducted by the Casualty Actuarial and Statistical Task Force of the National Association of Insurance Commissioners found that ___
+A 2017 survey conducted by the Casualty Actuarial and Statistical Task Force of the National Association of Insurance Commissioners found that the plurality of responding regulators identified "Filing complexity and/or a lack of resources or expertise" as a key challenge that impedes their ability to review GLMs or other predictive models.
+[@naic_summer_2017] Given that machine learning algorithms are generally regarded as more complex than GLMs, this implies that the challenge of communicating machine learning model results is significant.
 
--NAIC surveys indicate challenges for regulators in understanding even GLMs
-
-In response to a similar survey in Sping 2017, 33 state regulators noted that it would be helpful or very helpful for the NAIC to develop information and tools to assist in reviewing rate filings based on GLMs, and 34 noted that it would be helpful to develop similar items to assist in reviewing "Other Advanced Modeling Techniques." [cite: Spring Vol I Proceedings of the NAIC]
-
-One outgrowth of this need was the development of a white paper on best practices for regulatory review of predictive models.
+In response to the same survey, 33 state regulators noted that it would be helpful or very helpful for the NAIC to develop information and tools to assist in reviewing rate filings based on GLMs, and 34 noted that it would be helpful to develop similar items to assist in reviewing "Other Advanced Modeling Techniques." One outgrowth of this need was the development of a white paper on best practices for regulatory review of predictive models.
 The white paper focuses on review of GLMs, particularly with respect to private passenger automobile and homeowners' insurance.
@@
 However, among the guidance applicable to machine learning algorithms are the following (paraphrasing):
+
 * understand the relationship between the inputs and the expected loss or expense differences in risk,
 * determine that these input characteristics are not unfairly discriminatory, and
-* determine the extent of premium disruption among policyholders and be able to explain the source of premium disruptions. [cite: NAIC white paper]
+* determine the extent of premium disruption among policyholders and be able to explain the source of premium disruptions. [@naic_white_paper]
 
-Note that this list is not exhaustive of the guidance offered by any means, but that these three items are those that are applicable that may present greater challenge for machine learning algorithms compared to GLMs.
-[By comparison, guidance to understand that the variables are predictive of loss is a common goal of GLMs and ML, and in fact ML kind of gets it "for free" since it's about minimizing prediction error rather than maximizing likelihood.
-Not sure if this bears mentioning.
-Could also note the extent to which the NAIC guidance focuses on predictiveness as opposed to p-values.
-I think it's pretty questionable.]
+Note that this list is not exhaustive of the guidance offered by any means, but that these three items are those that may present greater challenge for machine learning algorithms compared to GLMs.
 

### 2019-10-03 manuscript/manuscript.Rmd

@@
 
-In this section, we attempt to develop a working definition of interpretability for ratemaking applications.
-While we will not provide a comprehensive survey of the prolific and fast evolving ML interpretability literature, we draw from it as appropriate in setting the stage for our discussion.
-There is not a standard definition of interpretability with consensus from all researchers; here are a few from frequently cited papers:
-
-- Ability to explain or to present in understandable terms to a human [@doshi-velezRigorousScience2017];
-- The degree to which an observer can understand the cause of a decision [@biranExplanationJustification2017]; and
-- A method is interpretable if a user can correctly and efficiently predict the method’s results [@kimExamplesAre2016].
-
-We motivate our discussion by considering several aspects of interpretability.
-For the time being, one can take any of the definitions cited above.
-As we proceed through the points below, we aim to arrive at a more scoped and relevant definition of what it means to attain interpretability in a pricing model.
-(TODO: add a sentence mapping out what we'll talk about in this section)
-
-## Not All Linear Models are Interpretable
-
-We look at some examples of models that are considered by many to be interpretable.
-The GLM is probably the most oft-cited example of such a model (or rather, modeling framework) in the actuarial literature.
-Given a set of inputs, we can easily reason about what the output of the model is.
-As an illustrative example, we may have a claims severity model with age, amount of insurance (AOI), and gender as features; assuming a log link function, we have
-
-\begin{equation}
-  \log(E(\text{severity})) = \beta_0 + \beta_1 \cdot \text{age} + \beta_2 \cdot \text{AOI} + \beta_3 \cdot \text{gender}_{\text{male}}.
-\end{equation}
-
-Here, we can tell, for example, what the model would predict for the expected severity if we were to increase age by a certain amount, *all else being equal*, because the relationship between the feature and the response is simply the coefficient $\beta_1$ and the link function.
-Another commonly cited example of an interpretable model is a decision tree.
-An illustrative example is shown in (TODO: FIGURE).
-Here, the prediction is arrived at by following a sequence of if-else decisions.
-
-Now, it is worth pointing out that, when declaring that GLM or decision trees are interpretable models, we are implicitly qualifying that we are considering a reasonable number of linear features.
-In fact, the ease at which we can reason about a model rapidly declines as the number of features and interactions increase, as in the following (somewhat pathological) example:
-
-\begin{equation}\label{eq:badglm}
-  \log(E(\text{severity})) = \beta_0 + \beta_1 \cdot \text{AOI}  + \beta_2\cdot\text{age} + \beta_3\cdot\text{age}^2 + \beta_4 \cdot \text{age:AOI} + \beta_5 \cdot \text{gender}_{\text{male}}.
-\end{equation}
-
-Similarly, one can see that in (TODO: FIGURE), large and deep trees are tough to reason about.
-In other words, even when working within the framework of an "interpretable" class of models, we may still end up with something that many would consider "black box."
-
-## The Machinery is Not a Secret
-
-An occasional misconception is that we have no visibility into how some ML models *compute* predictions, which render them uninterpretable.
-Outside of proprietary algorithms, all common ML models, including neural networks, gradient boosted trees, and random forests, are well studied and have large bodies of literature documenting their inner workings.
-As an example, a fitted neural network model on tabular data is simply a composition of linear transformations followed with nonlinear activation functions.
-As in Equation \ref{eq:badglm}, one can write down the mathematical equation for calculating the prediction given some inputs, but it may be difficult for a human to reason about it.
-This is a somewhat philosophical point, as we show later in the paper that we can still provide explanations of completely "black box" models, but is important to note given that pricing models undergo regulatory scrutiny.
-
-## Explanations are Social
-
-Proposed by @hiltonConversationalProcesses1990 and interpreted by @millerExplanationArtificial2017 in the context of ML, model explanations are conversations or social interactions.
-One consequence of this identification is that explanations need to be *relevant* to the *audience*.
-In developing, filing, and operationalizing a pricing model, one needs to accommodate a variety of stakeholders, each of whom has a different set of questions, assumptions, and technical capacity.
-First, there are internal stakeholders at the company, which includes management and underwriters.
-While some of the individuals in this audience may be technical, they are likely less familiar with predictive modeling techniques than the actuaries and data scientists who build the models.
-(TODO: what's a representative question here?) Next, we have the regulators, who may have limited resources to review the models, and will focus on a specific list of questions.
+(literature review, settle on a definition to work with, e.g. @doshi-velezRigorousScience2017, identify maybe 3 questions about a model to answer, tie to principles on p/c ratemaking)
 

### 2019-10-04 manuscript/manuscript.Rmd
combine stuff

--- 2019-09-27 manuscript/manuscript.Rmd
@@
 
-The rest of this paper is organized as follows: Section \ref{ratemaking} provides an overview of P&C ratemaking, Section \ref{interpretability} discusses model interpretability in the context of ratemaking and proposes specific tasks for model explanation, Section \ref{application} describes current model interpretation techniques and applies them to the tasks defined in the previous section, and Section \ref{conclusion} concludes.
+The rest of this paper is organized as follows: Section \ref{ratemaking} provides an overview of P&C ratemaking, Section \ref{need} discusses the importance of interpretation, Section \ref{interpretability} discusses model interpretability in the context of ratemaking and proposes specific tasks for model explanation, Section \ref{application} describes current model interpretation techniques and applies them to the tasks defined in the previous section, and Section \ref{conclusion} concludes.
 
@@
 ## Machine Learning in Ratemaking
+
 Paralelling the development of generalized linear models was the development of machine learning algorithms throughout the middle part of the 20th century.
@@
 Therefore, the techniques discussed in this paper may be viewed from the lens of providing useful information to regulators, but they should also be considered as part of a thorough vetting of any rating model.
+
+# The Need to Understand the Black Box {#need}
+
+Within the actuarial profession, Actuarial Standard of Practice 41 ("Actuarial Communications") notes that "...another actuary qualified in the same practice area [should be able to] make an objective appraisal of the reasonableness of the actuary's work as presented in the actuarial report." [@asop_41] Underlying this requirement is an assumption that the hypothetical other actuary qualified in the same practice area is adequately familiar with the relevant techniques employed.
+Although the syllabus of basic education is constantly changing, there has at times been an assumption that all techniques and assumptions that have ever been a part of the syllabus of basic education needn't be explained from first principles in general actuarial communications, and that an actuary practicing in the same field should be able to make an objective appraisal of the results from the methods found in the syllabus.
+[This can be supported by reference to the edits to ASOP 38 over time - originally it was designed to be about all models, but they revised it to be about only models that incorporate specialized knowledge outside of the actuary's area of expertise...
+do we need this citation, though?] This is notable because, beginning with the introduction of the MAS-I and MAS-II examination in July of 2018, several machine learning models were formally included in the syllabus of basic education.
+These exams cover a wide range of topics, such as splines, clustering algorithms, decision trees, boosting, and principle components analysis.
+[@cas_syllabus_2018]
+
+Nevertheless, machine learning poses something of a special challenge for ASOP 41 for several reasons.
+Machine learning models can be very ad hoc compared to traditional statistical models.
+Because many machine learning models are deterministic, they may not admit of standard metrics for model comparison (e.g., it's not straightforward to calculate an AIC over a neural network).
+In addition, machine learning methods are often combined into ensembles that may not be easily separated and that may, as a collection, cease to resemble a single standard version of a model.
+Complicating matters still further, machine learning models can be "black boxes" insofar as the final form of response curve cannot be easily predicted and may depend heavily on the available data (which may not, in turn, be available to the reviewer).
+
+This last item raises a final interesting issue.
+Generalized linear models and their ilk are often fitted using one of a handful of standard and well-understood approaches (e.g., maximum likelihood estimation).
+However, this is not possible in general with machine learning models, as machine learning algorithms often use loss surfaces that are very complex such that it may not be feasible to calculate the global minimum of the surface.
+Certainly, closed form representations of the loss surfaces are not generally available.
+For this reason, the training phase of a machine learning model is, in many ways, just as important to one's understanding as the model form and the data on which the model is fitted.
+Because the final model result is inseparable from these three components (training method, model form, and data), it is not generally adequate to just know the method employed to make an objective appraisal of the reasonableness of the model.
+More information is necessary.
+
+Of course, these comments only apply to the actuarial profession.
+Outside of the actuarial profession, communication of results may be more challenging.
+A 2017 survey conducted by the Casualty Actuarial and Statistical Task Force of the National Association of Insurance Commissioners found that the plurality of responding regulators identified "Filing complexity and/or a lack of resources or expertise" as a key challenge that impedes their ability to review GLMs or other predictive models.
+[@naic_summer_2017] Given that machine learning algorithms are generally regarded as more complex than GLMs, this implies that the challenge of communicating machine learning model results is significant.
+
+In response to the same survey, 33 state regulators noted that it would be helpful or very helpful for the NAIC to develop information and tools to assist in reviewing rate filings based on GLMs, and 34 noted that it would be helpful to develop similar items to assist in reviewing "Other Advanced Modeling Techniques." One outgrowth of this need was the development of a white paper on best practices for regulatory review of predictive models.
+The white paper focuses on review of GLMs, particularly with respect to private passenger automobile and homeowners' insurance.
+Some of the guidance offered in this regard is therefore not strictly applicable to the review of machine learning models.
+For example, as previously noted, p-values are not a concept that translates well to deterministic machine learning algorithms.
+However, among the guidance applicable to machine learning algorithms are the following (paraphrasing):
+
+* understand the relationship between the inputs and the expected loss or expense differences in risk,
+* determine that these input characteristics are not unfairly discriminatory, and
+* determine the extent of premium disruption among policyholders and be able to explain the source of premium disruptions. [@naic_white_paper]
+
+Note that this list is not exhaustive of the guidance offered by any means, but that these three items are those that may present greater challenge for machine learning algorithms compared to GLMs.
+
+Depending on the model, machine learning algorithms can be non-linear in the inputs.
+In areas where data are sparse, the model could generate unnatural response curves that could go undetected by a very high-level view of the model results.
+This leads to a situation where lift charts or tests on holdout data may indicate that the model is performing adequately, but it may be difficult to explain the reasons for a particular premium indicaton or a change in premium when policy characteristics change.
+
+-For this, we recommend ____ (grid search?)
+-Because the model form may be difficult to represent in the form of a single equation (and harder still to evaluate), one method of evaluating the reasonableness of model results is local approximations of the impact of different rating variables (LIME?)
 
@@
 
-Within the actuarial profession, Actuarial Standard of Practice 41 ("Actuarial Communications") notes that "...another actuary qualified in the same practice area [should be able to] make an objective appraisal of the reasonableness of the actuary's work as presented in the actuarial report." [@asop_41] Underlying this requirement is an assumption that the hypothetical other actuary qualified in the same practice area is adequately familiar with the relevant techniques employed.
-Although the syllabus of basic education is constantly changing, there has at times been an assumption that all techniques and assumptions that have ever been a part of the syllabus of basic education needn't be explained from first principles in general actuarial communications, and that an actuary practicing in the same field should be able to make an objective appraisal of the results from the methods found in the syllabus.
-[This can be supported by reference to the edits to ASOP 38 over time - originally it was designed to be about all models, but they revised it to be about only models that incorporate specialized knowledge outside of the actuary's area of expertise...
-do we need this citation, though?] This is notable because, beginning with the introduction of the MAS-I and MAS-II examination in July of 2018, several machine learning models were formally included in the syllabus of basic education.
-These exams cover a wide range of topics, such as splines, clustering algorithms, decision trees, boosting, and principle components analysis.
-[@cas_syllabus_2018]
-
-Nevertheless, machine learning poses something of a special challenge for ASOP 41 for several reasons.
-Machine learning models can be very ad hoc compared to traditional statistical models.
-Because many machine learning models are deterministic, they may not admit of standard metrics for model comparison (e.g., it's not straightforward to calculate an AIC over a neural network).
-In addition, machine learning methods are often combined into ensembles that may not be easily separated and that may, as a collection, cease to resemble a single standard version of a model.
-Complicating matters still further, machine learning models can be "black boxes" insofar as the final form of response curve cannot be easily predicted and may depend heavily on the available data (which may not, in turn, be available to the reviewer).
-
-This last item raises a final interesting issue.
-Generalized linear models and their ilk are often fitted using one of a handful of standard and well-understood approaches (e.g., maximum likelihood estimation).
-However, this is not possible in general with machine learning models, as machine learning algorithms often use loss surfaces that are very complex such that it may not be feasible to calculate the global minimum of the surface.
-Certainly, closed form representations of the loss surfaces are not generally available.
-For this reason, the training phase of a machine learning model is, in many ways, just as important to one's understanding as the model form and the data on which the model is fitted.
-Because the final model result is inseparable from these three components (training method, model form, and data), it is not generally adequate to just know the method employed to make an objective appraisal of the reasonableness of the model.
-More information is necessary.
-
-Of course, these comments only apply to the actuarial profession.
-Outside of the actuarial profession, communication of results may be more challenging.
-A 2017 survey conducted by the Casualty Actuarial and Statistical Task Force of the National Association of Insurance Commissioners found that the plurality of responding regulators identified "Filing complexity and/or a lack of resources or expertise" as a key challenge that impedes their ability to review GLMs or other predictive models.
-[@naic_summer_2017] Given that machine learning algorithms are generally regarded as more complex than GLMs, this implies that the challenge of communicating machine learning model results is significant.
-
-In response to the same survey, 33 state regulators noted that it would be helpful or very helpful for the NAIC to develop information and tools to assist in reviewing rate filings based on GLMs, and 34 noted that it would be helpful to develop similar items to assist in reviewing "Other Advanced Modeling Techniques." One outgrowth of this need was the development of a white paper on best practices for regulatory review of predictive models.
-The white paper focuses on review of GLMs, particularly with respect to private passenger automobile and homeowners' insurance.
-Some of the guidance offered in this regard is therefore not strictly applicable to the review of machine learning models.
-For example, as previously noted, p-values are not a concept that translates well to deterministic machine learning algorithms.
-However, among the guidance applicable to machine learning algorithms are the following (paraphrasing):
-
-* understand the relationship between the inputs and the expected loss or expense differences in risk,
-* determine that these input characteristics are not unfairly discriminatory, and
-* determine the extent of premium disruption among policyholders and be able to explain the source of premium disruptions. [@naic_white_paper]
-
-Note that this list is not exhaustive of the guidance offered by any means, but that these three items are those that may present greater challenge for machine learning algorithms compared to GLMs.
-
-Depending on the model, machine learning algorithms can be non-linear in the inputs.
-In areas where data are sparse, the model could generate unnatural response curves that could go undetected by a very high-level view of the model results.
-This leads to a situation where lift charts or tests on holdout data may indicate that the model is performing adequately, but it may be difficult to explain the reasons for a particular premium indicaton or a change in premium when policy characteristics change.
-
--For this, we recommend ____ (grid search?)
--Because the model form may be difficult to represent in the form of a single equation (and harder still to evaluate), one method of evaluating the reasonableness of model results is local approximations of the impact of different rating variables (LIME?)
-
 # Applying Model Interpretation Techniques {#application}

### 2019-10-04 manuscript/manuscript.Rmd
fix citation style

@@
 Early classification ratemaking procedures were typically univariate in nature.
-For example, [@lange_1966] notes that (at that time) most major lines of insurance used univariate methods based around the same principle: distributing an overall indication to territorial relativities or classification relativities based on the extent to which they deviated from the average experience.
+For example, @lange_1966 notes that (at that time) most major lines of insurance used univariate methods based around the same principle: distributing an overall indication to territorial relativities or classification relativities based on the extent to which they deviated from the average experience.
 
-[@bailey_simon_1960] introduced minimum bias methods, which were expanded throughout the 60s, 70s, and 80s.
-As computing power developed, minimum bias began to give away to generalized linear models, with papers such as [@brown_1988] and [@mildenhall_1999] bridging the gap between the methods.
+@bailey_simon_1960 introduced minimum bias methods, which were expanded throughout the 60s, 70s, and 80s.
+As computing power developed, minimum bias began to give away to generalized linear models, with papers such as @brown_1988 and @mildenhall_1999 bridging the gap between the methods.
 
 Arguably, generalized linear models predate minimum bias procedures by a significant margin.
-The term "Generalized Linear Model" was coined by [@nelder_wedderburn_1972], but generalizations of least squares linear regression date back at least to the 1930s.
+The term "Generalized Linear Model" was coined by @nelder_wedderburn_1972, but generalizations of least squares linear regression date back at least to the 1930s.
 Like minimum bias methods, GLMs did not become mainstream in actuarial science for some time.
-For example, the syllabus of basic education does not seem to include any mention of GLMs prior to [@brown_1988] in the 1990 syllabus for basic education.
-From there, GLMs seem to have received only passing mention until 2006 with the introduction of [@anderson_2005] to the syllabus.
+For example, the syllabus of basic education does not seem to include any mention of GLMs prior to @brown_1988 in the 1990 syllabus for basic education.
+From there, GLMs seem to have received only passing mention until 2006 with the introduction of @anderson_2005 to the syllabus.
 
@@
 Paralelling the development of generalized linear models was the development of machine learning algorithms throughout the middle part of the 20th century.
-Detailed histories of machine learning may be found in sources such as [@Nilsson_2009] and [@wang_raj_2017].
+Detailed histories of machine learning may be found in sources such as @Nilsson_2009 and @wang_raj_2017.
 Consistent with GLMs, machine learning was relatively unpopular in actuarial science until the last ten years as computing power has become cheaper and more easily available and as machine learning software packages have obviated the need for developing analyses from scratch each time an analysis is performed.
-Due to the breadth of machine learning as a field, it is difficult to identify the first time it entered the CAS syllabus; however, cluster analysis (in the form of k-means) seems to have been first included in 2011 with [@robertson_2009].
+Due to the breadth of machine learning as a field, it is difficult to identify the first time it entered the CAS syllabus; however, cluster analysis (in the form of k-means) seems to have been first included in 2011 with @robertson_2009.
 More recently, the MAS-I and MAS-II exams introduced in 2018 have included machine learning explicitly.
@@
 Within the area of ratemaking, machine learning is still in its infancy.
-A significant portion of machine learning applications to ratemaking has been in the context of automobile telematics, such as [@gao_2018], [@gao_2018_2], [@gao_2019], [@roel_2018], or [@wuthrich_2017].
+A significant portion of machine learning applications to ratemaking has been in the context of automobile telematics, such as @gao_2018, @gao_2018_2, @gao_2019, @roel_2018, or @wuthrich_2017.
 Presumably this focus has been a result of the high-dimensionality and complexity of telematics data, making it a field in which the unique abilities of machine learning techniques give a clear advantage over traditional approaches.
 
-Outside of telematics, [@yang_2018] uses a gradient tree-boosting approach to capture non-linearities that would be a challenge for GLMs.
-[@henckaerts_2018] makes use of "generalized additive models" to improve predictions of GLMs.
-Many researchers, in an apparent effort to demonstrate the range of possibilities and advantages of machine learning, have approached the topic by comparing many different machine learning algorithms within a single study, such as in [@dugas_2003], [@noll_2018], [@spedicato_2018].
+Outside of telematics, @yang_2018 uses a gradient tree-boosting approach to capture non-linearities that would be a challenge for GLMs.
+@henckaerts_2018 makes use of "generalized additive models" to improve predictions of GLMs.
+Many researchers, in an apparent effort to demonstrate the range of possibilities and advantages of machine learning, have approached the topic by comparing many different machine learning algorithms within a single study, such as in @dugas_2003, @noll_2018, @spedicato_2018.
 These studies make use of such varied techniques as regression trees, boosting machines, support vector machines, and neural networks.
@@
 do we need this citation, though?] This is notable because, beginning with the introduction of the MAS-I and MAS-II examination in July of 2018, several machine learning models were formally included in the syllabus of basic education.
-These exams cover a wide range of topics, such as splines, clustering algorithms, decision trees, boosting, and principle components analysis.
-[@cas_syllabus_2018]
+These exams cover a wide range of topics, such as splines, clustering algorithms, decision trees, boosting, and principle components analysis [@cas_syllabus_2018].
 
@@
 Outside of the actuarial profession, communication of results may be more challenging.
-A 2017 survey conducted by the Casualty Actuarial and Statistical Task Force of the National Association of Insurance Commissioners found that the plurality of responding regulators identified "Filing complexity and/or a lack of resources or expertise" as a key challenge that impedes their ability to review GLMs or other predictive models.
-[@naic_summer_2017] Given that machine learning algorithms are generally regarded as more complex than GLMs, this implies that the challenge of communicating machine learning model results is significant.
+A 2017 survey conducted by the Casualty Actuarial and Statistical Task Force of the National Association of Insurance Commissioners found that the plurality of responding regulators identified "Filing complexity and/or a lack of resources or expertise" as a key challenge that impedes their ability to review GLMs or other predictive models [@naic_summer_2017].
+Given that machine learning algorithms are generally regarded as more complex than GLMs, this implies that the challenge of communicating machine learning model results is significant.
 
-In response to the same survey, 33 state regulators noted that it would be helpful or very helpful for the NAIC to develop information and tools to assist in reviewing rate filings based on GLMs, and 34 noted that it would be helpful to develop similar items to assist in reviewing "Other Advanced Modeling Techniques." One outgrowth of this need was the development of a white paper on best practices for regulatory review of predictive models.
+In response to the same survey, 33 state regulators noted that it would be helpful or very helpful for the NAIC to develop information and tools to assist in reviewing rate filings based on GLMs, and 34 noted that it would be helpful to develop similar items to assist in reviewing "Other Advanced Modeling Techniques." One outgrowth of this need was the development of a white paper, @naic_white_paper, on best practices for regulatory review of predictive models.
 The white paper focuses on review of GLMs, particularly with respect to private passenger automobile and homeowners' insurance.
@@
 * determine that these input characteristics are not unfairly discriminatory, and
-* determine the extent of premium disruption among policyholders and be able to explain the source of premium disruptions. [@naic_white_paper]
+* determine the extent of premium disruption among policyholders and be able to explain the source of premium disruptions.
 

### 2019-10-05 manuscript/manuscript.Rmd

@@
 @bailey_simon_1960 introduced minimum bias methods, which were expanded throughout the 60s, 70s, and 80s.
-As computing power developed, minimum bias began to give away to generalized linear models, with papers such as @brown_1988 and @mildenhall_1999 bridging the gap between the methods.
+As computing power developed, minimum bias began to give away to GLMs, with papers such as @brown_1988 and @mildenhall_1999 bridging the gap between the methods.
 
-Arguably, generalized linear models predate minimum bias procedures by a significant margin.
-The term "Generalized Linear Model" was coined by @nelder_wedderburn_1972, but generalizations of least squares linear regression date back at least to the 1930s.
+Arguably, GLMs predate minimum bias procedures by a significant margin.
+The term was coined by @nelder_wedderburn_1972, but generalizations of least squares linear regression date back at least to the 1930s.
 Like minimum bias methods, GLMs did not become mainstream in actuarial science for some time.
@@
 
-Paralelling the development of generalized linear models was the development of machine learning algorithms throughout the middle part of the 20th century.
+Paralelling the development of GLM was the development of machine learning algorithms throughout the middle part of the 20th century.
 Detailed histories of machine learning may be found in sources such as @Nilsson_2009 and @wang_raj_2017.
@@
 Outside of telematics, @yang_2018 uses a gradient tree-boosting approach to capture non-linearities that would be a challenge for GLMs.
-@henckaerts_2018 makes use of "generalized additive models" to improve predictions of GLMs.
+@henckaerts_2018 makes use of generalized additive models (GAM) to improve predictions of GLMs.
 Many researchers, in an apparent effort to demonstrate the range of possibilities and advantages of machine learning, have approached the topic by comparing many different machine learning algorithms within a single study, such as in @dugas_2003, @noll_2018, @spedicato_2018.
@@
 This last item raises a final interesting issue.
-Generalized linear models and their ilk are often fitted using one of a handful of standard and well-understood approaches (e.g., maximum likelihood estimation).
+GLMs and their ilk are often fitted using one of a handful of standard and well-understood approaches (e.g., maximum likelihood estimation).
 However, this is not possible in general with machine learning models, as machine learning algorithms often use loss surfaces that are very complex such that it may not be feasible to calculate the global minimum of the surface.

### 2019-10-05 manuscript/manuscript.Rmd
cite actual text from naic whitepaper

@@
 For example, as previously noted, p-values are not a concept that translates well to deterministic machine learning algorithms.
-However, among the guidance applicable to machine learning algorithms are the following (paraphrasing):
+However, among the guidance applicable to machine learning algorithms are the following:
 
-* understand the relationship between the inputs and the expected loss or expense differences in risk,
-* determine that these input characteristics are not unfairly discriminatory, and
-* determine the extent of premium disruption among policyholders and be able to explain the source of premium disruptions.
+* Determine the extent the model causes premium disruption for individual policyholders, and how the insurer will
+explain the disruption to individual consumers that inquire about it.
+* Determine that individual input characteristics to a predictive model are related to the expected loss or expense
+differences in risk. Each input characteristic should have an intuitive or demonstrable actual relationship to expected loss or expense.
+* Determine that individual outputs from a predictive model and their associated selected relativities are not unfairly discriminatory.
 

### 2019-10-05 manuscript/manuscript.Rmd
glue sections

@@
 
-Note that this list is not exhaustive of the guidance offered by any means, but that these three items are those that may present greater challenge for machine learning algorithms compared to GLMs.
-
-Depending on the model, machine learning algorithms can be non-linear in the inputs.
-In areas where data are sparse, the model could generate unnatural response curves that could go undetected by a very high-level view of the model results.
-This leads to a situation where lift charts or tests on holdout data may indicate that the model is performing adequately, but it may be difficult to explain the reasons for a particular premium indicaton or a change in premium when policy characteristics change.
-
--For this, we recommend ____ (grid search?)
--Because the model form may be difficult to represent in the form of a single equation (and harder still to evaluate), one method of evaluating the reasonableness of model results is local approximations of the impact of different rating variables (LIME?)
+These items are by no means exhaustive, but they pertain to the concept of model interpretability for ratemaking that we develop next.
 

### 2019-10-06 manuscript/manuscript.Rmd
wrap up interp section

@@
 We motivate our discussion by considering several aspects of interpretability.
-For the time being, one can take any of the definitions cited above.
 As we proceed through the points below, we aim to arrive at a more scoped and relevant definition of what it means to attain interpretability in a pricing model.
-(TODO: add a sentence mapping out what we'll talk about in this section)
+We first clarify a couple concepts regarding interpretable *classes* of models and and the computational transparency of ML models.
+We then emphasize findings on the social aspects of ML explanations most relevant to our use case.
+With the provided background, we discuss a potential framework for implementing ML interpretability in practice.
 
@@
 
-An occasional misconception is that we have no visibility into how some ML models *compute* predictions, which render them uninterpretable.
+Another occasional misconception is that we have no visibility into how some ML models *compute* predictions, which render them uninterpretable.
 Outside of proprietary algorithms, all common ML models, including neural networks, gradient boosted trees, and random forests, are well studied and have large bodies of literature documenting their inner workings.
@@
 As in Equation \ref{eq:badglm}, one can write down the mathematical equation for calculating the prediction given some inputs, but it may be difficult for a human to reason about it.
-This is a somewhat philosophical point, as we show later in the paper that we can still provide explanations of completely "black box" models, but is important to note given that pricing models undergo regulatory scrutiny.
+We show later that we can still provide explanations of completely "black box" models, but is important to note that ML model predictions are still governed by mathematical rules, given that pricing models undergo regulatory scrutiny.
 
-## Explanations are Social
+## Explanations are Contextual
 
@@
 While some of the individuals in this audience may be technical, they are likely less familiar with predictive modeling techniques than the actuaries and data scientists who build the models.
-(TODO: what's a representative question here?) Next, we have the regulators, who may have limited resources to review the models, and will focus on a specific list of questions.
+Next, we have the regulators, who may have limited resources to review the models, and will focus on a specific list of questions.
+Finally, we have the public, who has an interest (perhaps more so than the other parties) because their lives are directly affected by the premiums charged.
+
+It is interesting to note that, the modelers, who are most familiar with the models, tend to be same people designing and communicating the explanations.
+This poses a challenge that @millerExplainableAI2017 calls "inmates running the asylum", where the modelers design explanations for *themselves* rather than the intended audience.
+
+Another point outlined in Miller's survey [@millerExplanationArtificial2017] is that explanations are *contrastive*.
+In other words, people are often interested in not why something happened, but rather why it happened instead of something else.
+For example, policyholders might not care exactly how their auto premiums are computed, but would like to know why they are being charged more than their coworkers who drive similar vehicles.
+As an extension, policyholders may want to know what they can change in order to obtain lower premiums.
+
+## Asking and Answering the Right Questions
+
+With the above considerations in mind, we propose a potential framework for interpreting ML models for insurance pricing.
+The actuarial profession, in collaboration with regulators, define a set of questions to be answered by explanations accompanying ML models, along with acceptance criteria and examples of successful explanations.
+In other words, interpretability for our purposes is defined as the ability of a model's explanations to answer the posed questions.
+
+These questions should relate to existing guidelines, such as those described in @naic_white_paper and outlined in Section \ref{need}, standards of practice, and regulation, and in fact should not be specific only to ML models.
+By standardizing on a set of questions, we reduce the burden of both companies and regulators; this is especially important for the latter, who are already resource constrained facing increasing variety of models being filed.
+This format should also be familiar to actuaries who are accustomed to adhering to specific guidelines in, for example, ASOPs.
+Like the ASOPs, We envision that these questions and guidelines will be continually updated to reflect feedback obtained and advances in research.
+
+While the realization of a set of such guidelines is an ambitious undertaking beyond the scope of this paper, we present in the next section a sample set of questions and techniques one can leverage to answer them.
+The goal of these case studies is twofold: to more concretely illustruate the proposed framework, and to expose the actuarial audience to modern ML interpretation techniques.
 

### 2019-10-23 manuscript/manuscript.Rmd
additions

@@
 author:
-  - name: Author One
+  - name: Kevin Kuo
     affil: a
-    email: foo@bar.foo
-  - name: Author Two
+    email: kevin@kasa.ai
+  - name: Daniel Lupton
     affil: b
-    email: bar@foo.bar
+    email: dlupton@taylorandmulder.com
 affiliation:
@@
     address: |
-      Foo
+      Kasa AI,
+      3040 78th Ave SE #1271,
+      Mercer Island, WA 98040, USA
   - num: b
     address: |
-      Bar
+      Taylor & Mulder,
+      10508 Rivers Bend Lane,
+      Potomac, MD 20854, USA
 bibliography: explain-pricing.bib
 appendix: appendix.tex
+citation_package: bibtex
 abstract: |
@@
   \def\tightlist{}
-output: rticles::tf_article
+output: 
+  bookdown::pdf_book:
+    base_format: rticles::tf_article
+    number_sections: true
 ---
@@
 
-## Not All Linear Models are Interpretable
+## Not All Linear Models are Interpretable {#linear-models}
 
@@
 While some of the individuals in this audience may be technical, they are likely less familiar with predictive modeling techniques than the actuaries and data scientists who build the models.
-Next, we have the regulators, who may have limited resources to review the models, and will focus on a specific list of questions.
+Next, we have the regulators, who may have limited resources to review the models, and will focus on a specific list of questions motivated by statute.
 Finally, we have the public, who has an interest (perhaps more so than the other parties) because their lives are directly affected by the premiums charged.
@@
 This poses a challenge that @millerExplainableAI2017 calls "inmates running the asylum", where the modelers design explanations for *themselves* rather than the intended audience.
+For example, they may be interested in technical questions like extrapolation behavior and shape the explanations accordingly, which may be lost on the prospective policyholder.
 
@@
 
-With the above considerations in mind, we propose a potential framework for interpreting ML models for insurance pricing.
-The actuarial profession, in collaboration with regulators, define a set of questions to be answered by explanations accompanying ML models, along with acceptance criteria and examples of successful explanations.
+With the above considerations in mind, we propose a potential framework for interpreting ML models for insurance pricing: the actuarial profession, in collaboration with regulators, define a set of questions to be answered by explanations accompanying ML models, along with acceptance criteria and examples of successful explanations.
 In other words, interpretability for our purposes is defined as the ability of a model's explanations to answer the posed questions.
@@
 While the realization of a set of such guidelines is an ambitious undertaking beyond the scope of this paper, we present in the next section a sample set of questions and techniques one can leverage to answer them.
-The goal of these case studies is twofold: to more concretely illustruate the proposed framework, and to expose the actuarial audience to modern ML interpretation techniques.
+The goal of these case studies is twofold: to more concretely illustrate the proposed framework, and to expose the actuarial audience to modern ML interpretation techniques.
 
@@
 
-(some definitions, e.g. global/model vs. local/instance, categorize questions accordingly)
+Now that we have established a framework for model interpretation in the form of asking and answering relevant questions, we demonstrate examples of such exchanges via an illustrative case study.
+Analytically, our starting point is a fitted deep neural network model for predicting loss costs.
+As the modeling details are of secondary importance, they are available in the appendix (TODO: add appendix and link to it.) The questions that we ask of the model are as follows:
 
-## (answer each question)
+1. What are the most important predictors in the model?
+2. How does the predicted loss cost change, on average, as we change an input?
+3. For a particular policyholder, how does each characteristic contribute to the loss cost prediction?
 
-(for each question, propose a technique, mention alternative techniques, pros/cons, implement, interpret results)
+## A Simplified View of Interpretation Techniques
+
+(TODO: add citation to interpretable ml book)
+
+Before we dive into the answering questions, we provide some background on model interpretation techniques.
+As the literature is quite sizeable and fast moving, we do not attempt a full taxomony of techniques.
+The goal is to orient ourselves among broad categories of techniques, so we can map them to tasks indicated by the questions being asked.
+We classify model interpretation techniques across two dimensions: intrinsic vs.
+post-hoc and global vs.
+local.
+
+### Intrinsic vs. Post-hoc
+
+Intrinsic model interpretation draws conclusions from the structure of the fitted model and are what we typically associate with "interpretable" classes of models.
+This is only viable with models with simple structures, such as the sparse linear model and shallow decision tree we see in Section \ref{linear-models}, where we arrive at explanations by reading off parameter estimates or a few decision rules.
+For algorithms that produce models with complex structure that do not lend themselves easily to intrinsic exploration, we can appeal to post-hoc techniques.
+This class of techniques interrogate the model by presenting it with data for and observing the prediction behavior of the model.
+These techniques are concerned with only the inputs and outputs, and hence are *agnostic* of the model itself, which means they can be applied also to simple models.
+Since most useful ML models have a level of complexity beyond the threshold of intrinsic interpretability, we focus on model-agnostic techniques in our case study.
+As we will see later on, the data that we present to the models are usually some perturbed variations of test data.
+
+### Global vs. Local
+
+Along the other dimension, we categorize model interpretations as global, or model-level, and local, or instance-level.
+The former class provides insights with respect to the model as a whole.
+Some examples of these eplanations include variable importances and sensitivities, on average, of the predicted response with respect to individual predictors.
+In our case study, questions 1 and 2 are associated with this category.
+On the other hand, question 3 pertains to an individual prediction, which would fall in the local interpretation category.
+In addition to individual variable attribution, we can also inquire about what would happen to the current predicted response if we were to perturb specific predictor variables.
+
+## Answering the Questions
+
+Having aligned the questions with the categories of interpretation techniques, we now introduce a sample of appropriate techniques to answer them.
+
+### Variable Importance
+
+> "What are the most important predictors in the model?"
+
+For linear models and their generalizations, and some ML models, measures of variable importance can be obtained from the fitted model structure.
+In the case of GLMs, one can observe the magnitudes of the $t$-statistics, whereas for random forests, one can use out-of-bag errors [@breimanRandomForests2001].
+For more complex models, such as the neural network in our case study, we need to devise another approach.
+
+We follow the methodology of permutation feature importance as described in @fisherAllModels2018, and utilize the notation introduced by @biecekPMVEE.
+The gist of the technique is as follows: to see how important a variable is, we make predictions without it and see how much worse off we are in terms of accuracy.
+One way to achieve this would be to re-fit the model many times (as many times as the number of variables.) However, this may be intractable with lengthy model training times or large number of variables, so a more popular approach is to instead keep the same fitted model but permute the values of each predictor.
+
+More formally, let $y$ denote the vector of responses, $X$ denote the matrix of predictor variables, $\widehat{f}$ denote the fitted model, and $L = \mathcal{L}(\widehat{f}(X), y)$ denote the value of the loss function, which is mean squared error in the case of regression.
+Now, if $\widetilde{X}^{j}$ denotes the predictor matrix where the $j$th variable has been permutated, then we can compute the loss with the permuted dataset as $L^{-j} = \mathcal{L}(\widehat{f}(\widetilde{X}^{j}), y)$.
+With this, we define the absolute variable importance $VI^{j}$ as $L^{-j} - L$ or the relative variable importance as $VI_{R}^{j} = \frac{L^{-j}}{L}$.
+
+In Figure \@ref(fig:fi), we show a plot of variable importances.
+We see that FOO and BAR are the most important variables in terms of contributing accuracy to our model.
+Note that these measures do not provide information regarding the directional sensitivity of the predictors on the response.
+In our particular example, we see that the "make" variable contributes most to the accuracy of the model with "XXX" contributing the least.
+This provides a way for the audience to quickly glance at the most relevant variables, and ask further question as necessary.
+For example, if a variable that can potentially be a proxy for a discriminatory variable appears as very important, more investigation would be warranted.
+
+```{r fi, out.width='60%', fig.align = "center", fig.cap="Permutation feature importances for ML model.", echo = FALSE}
+knitr::include_graphics("figures/fi-plot.png")
+```
+
+### Partial Dependence Plots
+
+> "How does the predicted loss cost change, on average, as we change an input?"
+
+For this question, we again consider first how it would be answered in the GLM setting.
+When the input predictor in question is continuous, we can answer the question by looking at the estimated coefficient, which provides the change in the response per unit change in the predictor (on the scale of the linear predictor).
+For non-parametric models and neural networks, where no coefficients are available, we can appeal to partial dependence plots (PDP), first proposed by @friedmanGreedyFunction2001 for gradient boosting machines (GBM).
+
+To describe PDP, we first introduce some notation. Let $x^j$ denote the input variable of interest. Then we define the partial dependence function as
+
+\begin{equation}
+h(z) = E_{X^{-j}}[\widehat{f}(x|x^j = z)],
+\end{equation}
+
+where the expectation is taken over the distribution of the other predictor variables.
+In other words, we marginalize them out so we can focus on the relationship between the predicted response and the variable of interest.
+Empirically, we estimate $h$ by
+
+\begin{equation}
+\widehat{h}(z) = \frac{1}{N}\sum_{i = 1}^{N}\widehat{f}(x_i|x_i^{j} = z).
+\end{equation}
+
+In Figure \@ref(fig:pdp), we exhibit the PDP for the "vehicle age" variable.
+We see that the average predicted loss cost decreases with vehicle age until the latter is around 18.
+Note that the ticks on the axis tell us that the data is quite thin for vehicle age > 18, so the apparent upward trend to the right is driven by just a few data points.
+With this in mind, we can conclude that as cars get older, we expect the loss cost predictions generated by our model to decrease.
+
+```{r pdp, out.width='60%', fig.align = "center", fig.cap="Partial dependence plot for ML model.", echo = FALSE}
+knitr::include_graphics("figures/pdp-plot.png")
+```
+
+### Variable Attribution
+
+> "For a particular policyholder, how does each characteristic contribute to the loss cost prediction?"
+
+In the previous two examples, we look at model-level explanations; now we move on to one where we investigate one particular prediction instance.
+As before, we consider how we would approach the question for linear models.
+For a GLM with a log link common in ratemaking applications, for example, we simply start with the base rate, then the exponentiated coefficients would have multiplicative effects on the final rate.
+Similar to the previous examples, for ML models in general we do not have directly interpretable weights.
+Instead, one way to arrive at variable contributions is calculating the change in the expected model prediction for each predictor, conditioned on other predictors.
+
+Formally, for a fitted model $\widehat{f}$ and a specific instance $x_*$, we would like to decompose the model prediction $\widehat{f}(x_*)$ into
+
+\begin{equation}
+\widehat{f}(x_* ) = v_0 + \sum_{j=1}^p v(j, x_*),
+\end{equation}
+
+where $v(j, x_*)$ denotes the contribution of the $j$th variable in instance $x_*$, defined as
+
+\begin{equation}\label{eq:var-contrib}
+v(j, x_* ) = E_X[\widehat{f}(X) | X^1 = x_* ^1, \dots, X^j = x_* ^j] - E_X[\widehat{f}(X^1 = x_* ^1, \dots, X^{j-1} = x_*^{j - 1})].
+\end{equation}
+
+Hence, the contribution of the $j$th variable to the prediction is the incremental change in the expected model prediction when we set $X^j = x_*^ j$ assuming the other variables take their values in $x_*$.
+Note here that this definition implies that the order in which we consider the variables affects the results.
+Empirically, the expectations in \@ref(eq:var-contrib) are calculated by sampling the test dataset.
+
+In Figure \@ref(fig:breakdown) we exhibit a waterfall plot of variable contributions.
+The "intercept" value denotes the average model prediction.
+The predicted loss cost for this particular policyholder is slightly less than average; the characteristics that makes this policyholder more risky are the fact that he is a male between the ages of 18 and 25; counteracting the risky driver characteristics are the vehicle properties: it is a GM vehicle built domestically and is seven years old.
+
+```{r breakdown, out.width='80%', fig.align = "center", fig.cap="Variable contribution plot for ML model.", echo = FALSE}
+knitr::include_graphics("figures/breakdown-plot.png")
+```
 
 # Conclusion {#conclusion}
-
-(conclude)

### 2019-10-23 R/explain.R
plotting code

@@ -1,0 +1,127 @@
+library(tidyverse)
+library(DALEX)
+library(ggplot2)
+library(scales)
+
+
+pins::board_register_github(name = "cork", repo = "kasaai/cork")
+testing_data <- pins::pin_get("toy-model-testing-data", board = "cork")
+piggyback::pb_download(file = "model_artifacts/toy-model.tar.gz", repo = "kasaai/cork")
+untar("model_artifacts/toy-model.tar.gz", exdir = "model_artifacts")
+toy_model <- keras::load_model_tf("model_artifacts/toy-model")
+
+predictors <- c(
+  "sex", "age_range", "vehicle_age", "make", 
+  "vehicle_category", "region"
+)
+
+custom_predict <- function(model, newdata) {
+  predict(model, newdata, batch_size = 10000)
+}
+
+explainer_nn <- DALEX::explain(
+  model = toy_model,
+  data = testing_data,
+  y = testing_data$loss_per_exposure,
+  predict_function = custom_predict,
+  label = "neural_net"
+)
+
+pdp_vehicle_age <- ingredients::partial_dependency(
+  explainer_nn, 
+  "vehicle_age",
+  N = 1500
+)
+
+as.data.frame(pdp_vehicle_age) %>% 
+  ggplot(aes(x = `_x_`, y = `_yhat_`)) + 
+  geom_line() +
+  xlab("Vehicle Age") +
+  ylab("Average Predicted Loss Cost") +
+  theme_bw() +
+  geom_rug(sides = "b")
+
+fi <- ingredients::feature_importance(
+  explainer_nn,
+  loss_function = function(observed, predicted, weights) {
+    sqrt(
+      sum(((observed - predicted) ^ 2 * weights) /sum(weights))
+    )
+  },
+  weights = testing_data$exposure,
+  variables = predictors,
+  B = 10,
+  n_sample = 50000
+)
+
+fi %>% 
+  as.data.frame() %>% 
+  (function(df) {
+    full_model_loss <- df %>% 
+      filter(variable == "_full_model_") %>% 
+      pull(dropout_loss)
+    df %>% 
+      filter(!variable %in% c("_full_model_", "_baseline_")) %>% 
+      ggplot(aes(x = reorder(variable, dropout_loss), y = dropout_loss)) +
+      geom_bar(stat = "identity") +
+      geom_hline(yintercept = full_model_loss, col = "red", linetype = "dashed")+
[…]

### 2019-10-23 R/explain.R
code for saving plots

@@ -3,7 +3,6 @@ library(DALEX)
 library(ggplot2)
 library(scales)
 
-
 pins::board_register_github(name = "cork", repo = "kasaai/cork")
 testing_data <- pins::pin_get("toy-model-testing-data", board = "cork")
 piggyback::pb_download(file = "model_artifacts/toy-model.tar.gz", repo = "kasaai/cork")
@@ -33,13 +32,15 @@ pdp_vehicle_age <- ingredients::partial_dependency(
   N = 1500
 )
 
-as.data.frame(pdp_vehicle_age) %>% 
+pdp_plot <- as.data.frame(pdp_vehicle_age) %>% 
   ggplot(aes(x = `_x_`, y = `_yhat_`)) + 
   geom_line() +
   xlab("Vehicle Age") +
   ylab("Average Predicted Loss Cost") +
   theme_bw() +
   geom_rug(sides = "b")
+
+ggsave("manuscript/figures/pdp-plot.png", plot = pdp_plot)
 
 fi <- ingredients::feature_importance(
   explainer_nn,
@@ -54,7 +55,7 @@ fi <- ingredients::feature_importance(
   n_sample = 50000
 )
 
-fi %>% 
+fi_plot <- fi %>% 
   as.data.frame() %>% 
   (function(df) {
     full_model_loss <- df %>% 
@@ -75,6 +76,8 @@ fi %>%
       NULL
   })
 
+ggsave("manuscript/figures/fi-plot.png", plot = fi_plot)
+
 sample_row <- testing_data[1,] %>% 
   select(predictors)
 breakdown <- iBreakDown::break_down(explainer_nn, sample_row)
@@ -88,7 +91,7 @@ df <- breakdown %>%
                      paste0("+", .x),
                      .x))
 
-p <- df %>% 
+breakdown_plot <- df %>% 
   ggplot(aes(reorder(variable, position), fill = sign,
              xmin = position - 0.40, 
              xmax = position + 0.40, 
@@ -112,16 +115,15 @@ p <- df %>%
   coord_flip() +
   theme_bw() +
   theme(legend.position = "none") + 
-  NULL
-p + geom_text(
-  aes(label = label, 
-      y = pmax(df$cummulative,  df$cummulative - df$contribution)), 
-  nudge_y = 10,
-  hjust = "inward", 
-  color = "black"
-) +
+  geom_text(
+    aes(label = label, 
[…]

### 2019-10-23 manuscript/manuscript.Rmd
decision tree plots

@@
 While we will not provide a comprehensive survey of the prolific and fast evolving ML interpretability literature, we draw from it as appropriate in setting the stage for our discussion.
-There is not a standard definition of interpretability with consensus from all researchers; here are a few from frequently cited papers:
+Even among researchers in ML, there is not a consensus on the definition of interpretability; here are a few from frequently cited papers:
 
@@
 
-We look at some examples of models that are considered by many to be interpretable.
+Let us look at some examples of models that are considered by many to be interpretable.
 The GLM is probably the most oft-cited example of such a model (or rather, modeling framework) in the actuarial literature.
 Given a set of inputs, we can easily reason about what the output of the model is.
-As an illustrative example, we may have a claims severity model with age, amount of insurance (AOI), and gender as features; assuming a log link function, we have
+As an illustrative example, we may have a claims severity model with driver age, amount of insurance (AOI), and sex as features; assuming a log link function, we have
 
 \begin{equation}
-  \log(E(\text{severity})) = \beta_0 + \beta_1 \cdot \text{age} + \beta_2 \cdot \text{AOI} + \beta_3 \cdot \text{gender}_{\text{male}}.
+  \log(E(\text{severity})) = \beta_0 + \beta_1 \cdot \text{age} + \beta_2 \cdot \text{AOI} + \beta_3 \cdot \text{sex}_{\text{male}}.
 \end{equation}
@@
 Another commonly cited example of an interpretable model is a decision tree.
-An illustrative example is shown in (TODO: FIGURE).
+An illustrative example is shown in Figure \@ref(fig:tree-plot1).
 Here, the prediction is arrived at by following a sequence of if-else decisions.
+
+```{r tree-plot1, echo = FALSE, fig.cap="A simple decision tree for loss cost prediction.", out.width="70%", fig.align='center'}
+knitr::include_graphics("figures/tree-plot1.png")
+```
 
@@
 
-Similarly, one can see that in (TODO: FIGURE), large and deep trees are tough to reason about.
+Similarly, one can see that in Figure \@ref(fig:tree-plot2), large and deep trees are tough to reason about.
 In other words, even when working within the framework of an "interpretable" class of models, we may still end up with something that many would consider "black box."
+
+```{r tree-plot2, echo = FALSE, fig.cap="A more complex decision tree for loss cost prediction.", out.width="90%", fig.align='center'}
+knitr::include_graphics("figures/tree-plot2.png")
+```
 
@@
 Note that these measures do not provide information regarding the directional sensitivity of the predictors on the response.
-In our particular example, we see that the "make" variable contributes most to the accuracy of the model with "XXX" contributing the least.
+In our particular example, we see that the "make" variable contributes most to the accuracy of the model with "sex" contributing the least.
 This provides a way for the audience to quickly glance at the most relevant variables, and ask further question as necessary.

### 2019-10-24 manuscript/manuscript.Rmd
use elsevier template

@@
 ---
-title: |
- Towards Explainability of Machine Learning Models in Insurance Pricing
-type: WIP DRAFT 
+title: Towards Explainability of Machine Learning Models in Insurance Pricing
 author:
   - name: Kevin Kuo
-    affil: a
     email: kevin@kasa.ai
+    affiliation: a
+    footnote: Corresponding Author
   - name: Daniel Lupton
-    affil: b
     email: dlupton@taylorandmulder.com
-affiliation:
-  - num: a
-    address: |
-      Kasa AI,
-      3040 78th Ave SE #1271,
-      Mercer Island, WA 98040, USA
-  - num: b
-    address: |
-      Taylor & Mulder,
-      10508 Rivers Bend Lane,
-      Potomac, MD 20854, USA
+    affiliation: b
+address:
+  - code: a
+    address: Kasa AI, 3040 78th Ave SE \#1271, Mercer Island, WA 98040, USA
+  - code: b
+    address: Taylor & Mulder, 10508 Rivers Bend Lane, Potomac, MD 20854, USA
+abstract: |
+  Machine learning methods have garnered increasing interest among actuaries in recent years.
+However, their adoption by practitioners has been limited, partly due to the lack of transparency of these methods, as compared to generalized linear models.
+In this paper, we discuss the need for model interpretability in property & casualty insurance ratemaking, propose a framework for explaining models, and present a case study to illustrate the framework.
+
+journal: "[journal]"
+date: "`r Sys.Date()`"
 bibliography: explain-pricing.bib
-appendix: appendix.tex
-citation_package: bibtex
-abstract: |
-  Abstract tbd
-keywords: |
-  actuarial science; general insurance
-header-includes: |
-  \usepackage{hyperref}
-  \usepackage[utf8]{inputenc}
-  \def\tightlist{}
+#linenumbers: true
+numbersections: true
+# csl: elsevier-harvard.csl
 output: 
   bookdown::pdf_book:
-    base_format: rticles::tf_article
-    number_sections: true
+    base_format: rticles::elsevier_article
+layout: preprint, 3p, twocolumn, letterpaper, 12pt
 ---
@@
 Paralelling the development of GLM was the development of machine learning algorithms throughout the middle part of the 20th century.
-Detailed histories of machine learning may be found in sources such as @Nilsson_2009 and @wang_raj_2017.
+Detailed histories of machine learning may be found in sources such as @nilsson_2009 and @wang_raj_2017.
 Consistent with GLMs, machine learning was relatively unpopular in actuarial science until the last ten years as computing power has become cheaper and more easily available and as machine learning software packages have obviated the need for developing analyses from scratch each time an analysis is performed.
@@
 
-```{r tree-plot1, echo = FALSE, fig.cap="A simple decision tree for loss cost prediction.", out.width="70%", fig.align='center'}
+```{r tree-plot1, echo = FALSE, fig.cap="A simple decision tree for loss cost prediction.", out.width="100%", fig.align='center'}
 knitr::include_graphics("figures/tree-plot1.png")
@@
 
-```{r tree-plot2, echo = FALSE, fig.cap="A more complex decision tree for loss cost prediction.", out.width="90%", fig.align='center'}
+```{r tree-plot2, echo = FALSE, fig.cap="A more complex decision tree for loss cost prediction.", out.width="60%", fig.align='center',  fig.env='figure*'}
 knitr::include_graphics("figures/tree-plot2.png")
@@
 
-```{r fi, out.width='60%', fig.align = "center", fig.cap="Permutation feature importances for ML model.", echo = FALSE}
+```{r fi, out.width='100%', fig.align = "center", fig.cap="Permutation feature importances for ML model.", echo = FALSE}
 knitr::include_graphics("figures/fi-plot.png")
@@
 For non-parametric models and neural networks, where no coefficients are available, we can appeal to partial dependence plots (PDP), first proposed by @friedmanGreedyFunction2001 for gradient boosting machines (GBM).
+
+```{r pdp, out.width='70%', fig.align = "center", fig.cap="Partial dependence plot for ML model.", echo = FALSE,  fig.env='figure*', fig.pos='h'}
+knitr::include_graphics("figures/pdp-plot.png")
+```
+
+```{r breakdown, out.width='80%', fig.align = "center", fig.cap="Variable contribution plot for ML model.", echo = FALSE, fig.env = 'figure*', fig.pos = 'h'}
+knitr::include_graphics("figures/breakdown-plot.png")
+```
 
@@
 With this in mind, we can conclude that as cars get older, we expect the loss cost predictions generated by our model to decrease.
-
-```{r pdp, out.width='60%', fig.align = "center", fig.cap="Partial dependence plot for ML model.", echo = FALSE}
-knitr::include_graphics("figures/pdp-plot.png")
-```
 
@@
 
-```{r breakdown, out.width='80%', fig.align = "center", fig.cap="Variable contribution plot for ML model.", echo = FALSE}
-knitr::include_graphics("figures/breakdown-plot.png")
-```
+# Conclusion {#conclusion}
 
-# Conclusion {#conclusion}
+References {#references .unnumbered}
+==========
+<div id="refs"></div>
+
+\appendix
+
+Appendix {.unnumbered}
+=============
+
+# ML Model Development
+
+In this appendix, we describe the ML model and the data used to train it.
+Note that, for our paper, the ultimate goal of the modeling procedure is to develop something that can produce predictions.
+As a result, we do not follow standard practices for supervised learning model development.
+However, for the sake of completeness and reproducibility, we include the details of the process here.
+
+## Data
+
+We use data from the AUTOSEG ("Automobile Statistics System") of Brazil's Superintendence of Private Insurance (SUSEP).
+The organization maintains policy-characteristics-level data, including claim counts and amounts, for all insured automobiles in Brazil.
+The data contains variables from policyholder characteristics to losses by peril.
+
+## Model
+
+Model

### 2019-10-24 manuscript/manuscript.Rmd
fix equation alignment

@@
 Given a set of inputs, we can easily reason about what the output of the model is.
-As an illustrative example, we may have a claims severity model with driver age, amount of insurance (AOI), and sex as features; assuming a log link function, we have
+As an illustrative example, we may have a claims severity model with driver age, sex, and vehicle age as features; assuming a log link function and $Y$ denotes the response, we have
 
-\begin{equation}
-  \log(E(\text{severity})) = \beta_0 + \beta_1 \cdot \text{age} + \beta_2 \cdot \text{AOI} + \beta_3 \cdot \text{sex}_{\text{male}}.
-\end{equation}
+\begin{align}
+& \log(E[Y]) = \beta_0 + \beta_1 \cdot \text{age} \nonumber \\
+& \quad + \beta_2 \cdot \text{vehicle\_age} + \beta_3 \cdot \text{sex}_{\text{male}}.
+\end{align}
 
@@
 
-\begin{equation}\label{eq:badglm}
-  \log(E(\text{severity})) = \beta_0 + \beta_1 \cdot \text{AOI}  + \beta_2\cdot\text{age} + \beta_3\cdot\text{age}^2 + \beta_4 \cdot \text{age:AOI} + \beta_5 \cdot \text{gender}_{\text{male}}.
-\end{equation}
+\begin{align}\label{eq:badglm}
+& \log(E[Y]) = \beta_0 + \beta_1 \cdot \text{age} + \beta_2 \cdot \text{vehicle\_age} \nonumber\\
+& \quad + \beta_3\cdot\text{vehicle\_age}^2 \nonumber \\
+& \quad + \beta_4 \cdot \text{age} \cdot \text{vehicle\_age} \nonumber \\
+& \quad + \beta_5 \cdot \text{sex}_{\text{male}} + \beta_6 \cdot \text{sex}_{\text{male}} \cdot \text{age}. 
+\end{align}
 
@@
 
-\begin{equation}\label{eq:var-contrib}
-v(j, x_* ) = E_X[\widehat{f}(X) | X^1 = x_* ^1, \dots, X^j = x_* ^j] - E_X[\widehat{f}(X^1 = x_* ^1, \dots, X^{j-1} = x_*^{j - 1})].
-\end{equation}
+\begin{align}\label{eq:var-contrib}
+& v(j, x_* ) = E_X[\widehat{f}(X) | X^1 = x_* ^1, \dots, X^j = x_* ^j] \nonumber\\
+& \quad - E_X[\widehat{f}(X^1 = x_* ^1, \dots, X^{j-1} = x_*^{j - 1})].
+\end{align}
 

### 2019-10-24 manuscript/manuscript.Rmd

@@
 As we proceed through the points below, we aim to arrive at a more scoped and relevant definition of what it means to attain interpretability in a pricing model.
-We first clarify a couple concepts regarding interpretable *classes* of models and and the computational transparency of ML models.
-We then emphasize findings on the social aspects of ML explanations most relevant to our use case.
-With the provided background, we discuss a potential framework for implementing ML interpretability in practice.
+In the remainder of this section, we clarify a couple concepts regarding interpretable *classes* of models and and the computational transparency of ML models, outline findings on the social aspects of ML explanations most relevant to our use case, then discuss a potential framework for implementing ML interpretability in practice.
 
@@
 Given a set of inputs, we can easily reason about what the output of the model is.
-As an illustrative example, we may have a claims severity model with driver age, sex, and vehicle age as features; assuming a log link function and $Y$ denotes the response, we have
+As an illustrative example, consider a claim severity model with driver age, sex, and vehicle age as features; assuming a log link function and letting $Y$ denotes the response, we have
 
@@
 
-Here, we can tell, for example, what the model would predict for the expected severity if we were to increase age by a certain amount, *all else being equal*, because the relationship between the feature and the response is simply the coefficient $\beta_1$ and the link function.
+Here, we can tell, for example, what the model would predict for the expected severity if we were to increase age by a certain amount, *all else being equal*, because the relationship between the feature and the response is simply multiplication by the coefficient $\beta_1$ and applying the inverse link function.
 Another commonly cited example of an interpretable model is a decision tree.
@@
 
-Now, it is worth pointing out that, when declaring that GLM or decision trees are interpretable models, we are implicitly qualifying that we are considering a reasonable number of linear features.
-In fact, the ease at which we can reason about a model rapidly declines as the number of features and interactions increase, as in the following (somewhat pathological) example:
+Now, it is worth pointing out that, when declaring that GLM or decision trees are interpretable models, we are implicitly assuming that we are considering only a handful  of features.
+In fact, the ease at which we can reason about a model rapidly declines as the number of features, transformations, and interactions increase, as in the following (somewhat pathological) example:
 
@@
 
-```{r tree-plot2, echo = FALSE, fig.cap="A more complex decision tree for loss cost prediction.", out.width="60%", fig.align='center',  fig.env='figure*'}
+```{r tree-plot2, echo = FALSE, fig.cap="A more complex decision tree. This is still much simpler than typical realistic examples.", out.width="60%", fig.align='center',  fig.env='figure*'}
 knitr::include_graphics("figures/tree-plot2.png")
@@
 As in Equation \ref{eq:badglm}, one can write down the mathematical equation for calculating the prediction given some inputs, but it may be difficult for a human to reason about it.
-We show later that we can still provide explanations of completely "black box" models, but is important to note that ML model predictions are still governed by mathematical rules, given that pricing models undergo regulatory scrutiny.
+We show later that we can still provide explanations of completely "black box" models, but is important to note that ML model predictions are still governed by mathematical rules.
 
@@
 
-With the above considerations in mind, we propose a potential framework for interpreting ML models for insurance pricing: the actuarial profession, in collaboration with regulators, define a set of questions to be answered by explanations accompanying ML models, along with acceptance criteria and examples of successful explanations.
+With the above considerations in mind, we propose a potential framework for interpreting ML models for insurance pricing: the actuarial profession, in collaboration with regulators and representatives of the public, define a set of questions to be answered by explanations accompanying ML models, along with acceptance criteria and examples of successful explanations.
 In other words, interpretability for our purposes is defined as the ability of a model's explanations to answer the posed questions.
@@
 Analytically, our starting point is a fitted deep neural network model for predicting loss costs.
-As the modeling details are of secondary importance, they are available in the appendix (TODO: add appendix and link to it.) The questions that we ask of the model are as follows:
+As the modeling details are of secondary importance, they are available in Appendix \@ref{model-dev}.
+The questions that we ask of the model are as follows:
 
@@
 
+In our discussion, we adopt the organization of techniques and some notation presented in @molnar2018interpretable and @biecekPMVEE, which are comprehensive references on the most established ML interpretation techniques.
+
 ## A Simplified View of Interpretation Techniques
-
-(TODO: add citation to interpretable ml book)
 
@@
 The goal is to orient ourselves among broad categories of techniques, so we can map them to tasks indicated by the questions being asked.
-We classify model interpretation techniques across two dimensions: intrinsic vs.
+Model interpretation techniques can be categorized across two dimensions: intrinsic vs.
 post-hoc and global vs.
@@
 This class of techniques interrogate the model by presenting it with data for and observing the prediction behavior of the model.
-These techniques are concerned with only the inputs and outputs, and hence are *agnostic* of the model itself, which means they can be applied also to simple models.
+These techniques are concerned with only the inputs and outputs, and hence are *agnostic* of the model itself, which means they can also be applied to simple models.
 Since most useful ML models have a level of complexity beyond the threshold of intrinsic interpretability, we focus on model-agnostic techniques in our case study.
@@
 
-Having aligned the questions with the categories of interpretation techniques, we now introduce a sample of appropriate techniques to answer them.
+Having aligned the questions with the categories of interpretation techniques, we now introduce a selection of appropriate techniques to answer them.
 
@@
 
-> "What are the most important predictors in the model?"
+"What are the most important predictors in the model?"
 
@@
 In Figure \@ref(fig:fi), we show a plot of variable importances.
-We see that FOO and BAR are the most important variables in terms of contributing accuracy to our model.
+In our particular example, we see that the "make" variable contributes most to the accuracy of the model with "sex" contributing the least.
 Note that these measures do not provide information regarding the directional sensitivity of the predictors on the response.
-In our particular example, we see that the "make" variable contributes most to the accuracy of the model with "sex" contributing the least.
 This provides a way for the audience to quickly glance at the most relevant variables, and ask further question as necessary.
@@
 
-> "How does the predicted loss cost change, on average, as we change an input?"
+"How does the predicted loss cost change, on average, as we change an input?"
 
@@
 
-To describe PDP, we first introduce some notation. Let $x^j$ denote the input variable of interest. Then we define the partial dependence function as
+To describe PDP, we need to introduce some additional notation. Let $x^j$ denote the input variable of interest. Then we define the partial dependence function as
 
@@
 \begin{equation}
-\widehat{h}(z) = \frac{1}{N}\sum_{i = 1}^{N}\widehat{f}(x_i|x_i^{j} = z).
+\widehat{h}(z) = \frac{1}{N}\sum_{i = 1}^{N}\widehat{f}(x_i|x_i^{j} = z),
 \end{equation}
+
+where $N$ is the number of records in our dataset.
 
@@
 We see that the average predicted loss cost decreases with vehicle age until the latter is around 18.
-Note that the ticks on the axis tell us that the data is quite thin for vehicle age > 18, so the apparent upward trend to the right is driven by just a few data points.
+Note that the ticks on the $x$-axis tell us that the data is quite thin for vehicle age greater than 18, so the apparent upward trend to the right is driven by just a few data points.
 With this in mind, we can conclude that as cars get older, we expect the loss cost predictions generated by our model to decrease.
+
+The question posed here is particularly important for regulators, who would like to know whether each variable affects the prediction in the direction that is expected, based on intuition, experience, and existing models.
+During the model development stage, PDP can also be used as a reasonableness test for candidate models by identifying unexpected relationships for the analyst to investigate.
 
@@
 
-> "For a particular policyholder, how does each characteristic contribute to the loss cost prediction?"
+"For a particular policyholder, how does each characteristic contribute to the loss cost prediction?"
 
@@
 
+Instance-level explanations are useful for investigating specific problematic predictions generated by the model.
+Regulators and model reviewers may be interested in variable contributions for the safest and riskiest policyholders to see if they conform to intuition.
+A policyholder with a particularly high premium may wish to find out what of their characteristics contribute to it, and may follow up with a question about how he can lower it, which would require another type of explanation.
+
+## Other Techniques
+
+(TODO: briefly discuss other stuff no covered)
+
 # Conclusion {#conclusion}
+
+(TODO)
 
@@
 
-# ML Model Development
+# ML Model Development {#model-dev}
 
@@
 The data contains variables from policyholder characteristics to losses by peril.
+(TODO: finish)
 
@@
 
-Model
+(TODO)

### 2019-10-24 manuscript/manuscript.Rmd
10pt font

@@
     base_format: rticles::elsevier_article
-layout: preprint, 3p, twocolumn, letterpaper, 12pt
+layout: preprint, 3p, twocolumn, letterpaper, 10pt
 ---

### 2019-10-24 manuscript/manuscript.Rmd
add appendix

@@
 
+Acknowledgments {.unnumbered}
+==========
+
+This work is supported by the Casualty Actuarial Society.
+
 References {#references .unnumbered}
@@
 
-# ML Model Development {#model-dev}
+```{r variables, results='asis', echo = FALSE}
+library(magrittr)
+variables <- tibble::tribble(
+  ~ Variable, ~ Type, ~ Transformation,
+  "Age range", "Categorical", "One-hot encode",
+  "Sex", "Categorical", "One-hot encode",
+  "Vehicle category", "Categorical", "One-hot encode",
+  "Make", "Categorical", "Embed in $\\mathbb{R}^2$",
+  "Vehicle age", "Numeric", "Center and scale",
+  "Region", "Categorical", "Embed in $\\mathbb{R}^2$"
+)
+knitr::kable(
+  variables,
+  caption = 'Input variables and their transformations.',
+  booktabs = TRUE, 
+  table.env='table',
+  escape = FALSE,
+  linesep = ""
+) %>% 
+  kableExtra::kable_styling(latex_options = "hold_position")
+```
+
+# Model Development {#model-dev}
 
@@
 As a result, we do not follow standard practices for supervised learning model development.
-However, for the sake of completeness and reproducibility, we include the details of the process here.
+However, for the sake of completeness and reproducibility, we include an overview of the process here.
+Implementation is done using the R [@rlang] interface to TensorFlow [@tensorflow2015-whitepaper].
+The model explanation visualizations utilize the implemention by @biecekPMVEE, and the code to reproduce them are available on GitHub[^1].
+
+[^1]: https://github.com/kasaai/explain-ml-pricing
 
@@
 The data contains variables from policyholder characteristics to losses by peril.
-(TODO: finish)
+We use the records from the first half of 2012, which contains 1,707,651 records.
+One-fifth of the data is reserved for testing; the remainder is further split into 3/4 of analysis and 1/4 into assessment for determining early stopping.
 
@@
 
-(TODO)
+Table \@ref(tab:variables) shows the input variables to our model and their associated transformations.
+For "make" and "region", we map each level to a point in $\mathbb{R}^2$ through embedding layers.
+The model is a feedforward neural network with two hidden layers with 64 units each.
+The activations for the hidden layers are ReLU while for the output layer it is softplus.
+Exposures for each record are used as sample weights during training.
+We fit the model via ADAM with an initial learning rate of 0.1, a mini-batch size of 10,000, and trigger early stopping when the mean squared error on the asssessment set does not improve for five epochs.

### 2019-10-25 manuscript/manuscript.Rmd
some edits

@@
 
-# The Need to Understand the Black Box {#need}
+# The Need to See Inside the Black Box {#need}
 
@@
 While we will not provide a comprehensive survey of the prolific and fast evolving ML interpretability literature, we draw from it as appropriate in setting the stage for our discussion.
-Even among researchers in ML, there is not a consensus on the definition of interpretability; here are a few from frequently cited papers:
+Even among researchers in the subject, there is not a consensus on the definition of interpretability; here are a few from frequently cited papers:
 
@@
 
-Let us look at some examples of models that are considered by many to be interpretable.
-The GLM is probably the most oft-cited example of such a model (or rather, modeling framework) in the actuarial literature.
+In the actuarial science literature, the GLM is probably the most oft-cited example of an easily interpretable model.
 Given a set of inputs, we can easily reason about what the output of the model is.
@@
 Here, we can tell, for example, what the model would predict for the expected severity if we were to increase age by a certain amount, *all else being equal*, because the relationship between the feature and the response is simply multiplication by the coefficient $\beta_1$ and applying the inverse link function.
+
 Another commonly cited example of an interpretable model is a decision tree.
@@
 
-Now, it is worth pointing out that, when declaring that GLM or decision trees are interpretable models, we are implicitly assuming that we are considering only a handful  of features.
-In fact, the ease at which we can reason about a model rapidly declines as the number of features, transformations, and interactions increase, as in the following (somewhat pathological) example:
+Now, it is worth pointing out that, when declaring that GLMs or decision trees are interpretable models, we are implicitly assuming that we are considering only a handful  of features.
+In fact, the ease at which we can reason about a model declines as the number of features, transformations of them, and interactions increase, as in the following (somewhat pathological) example:
 
@@
 
-Similarly, one can see that in Figure \@ref(fig:tree-plot2), large and deep trees are tough to reason about.
+Similarly, one can see that in Figure \@ref(fig:tree-plot2), larger trees are tough to reason about.
 In other words, even when working within the framework of an "interpretable" class of models, we may still end up with something that many would consider "black box."
@@
 
-Another occasional misconception is that we have no visibility into how some ML models *compute* predictions, which render them uninterpretable.
+Another occasional misconception is that we have no visibility into how some ML models *compute* predictions, which renders them uninterpretable.
 Outside of proprietary algorithms, all common ML models, including neural networks, gradient boosted trees, and random forests, are well studied and have large bodies of literature documenting their inner workings.
-As an example, a fitted neural network model on tabular data is simply a composition of linear transformations followed with nonlinear activation functions.
+As an example, a fitted feedforward neural network is simply a composition of linear transformations followed by nonlinear activation functions.
 As in Equation \ref{eq:badglm}, one can write down the mathematical equation for calculating the prediction given some inputs, but it may be difficult for a human to reason about it.
-We show later that we can still provide explanations of completely "black box" models, but is important to note that ML model predictions are still governed by mathematical rules.
+We show later that we can still provide explanations of completely "black box" models, but is important to note that ML model predictions are still governed by mathematical rules, and are deterministic in most cases.
 

### 2019-10-25 manuscript/manuscript.Rmd
copyedits

@@
 While some of the individuals in this audience may be technical, they are likely less familiar with predictive modeling techniques than the actuaries and data scientists who build the models.
-Next, we have the regulators, who may have limited resources to review the models, and will focus on a specific list of questions motivated by statute.
+Next, we have the regulators, who may have limited resources to review the models, and will focus on a specific list of questions motivated by statute and public policy.
 Finally, we have the public, who has an interest (perhaps more so than the other parties) because their lives are directly affected by the premiums charged.
@@
 This poses a challenge that @millerExplainableAI2017 calls "inmates running the asylum", where the modelers design explanations for *themselves* rather than the intended audience.
-For example, they may be interested in technical questions like extrapolation behavior and shape the explanations accordingly, which may be lost on the prospective policyholder.
+For example, they may be interested in technical questions such as extrapolation behavior and shape the explanations accordingly, which may be lost on the prospective policyholder.
 
@@
 Analytically, our starting point is a fitted deep neural network model for predicting loss costs.
-As the modeling details are of secondary importance, they are available in Appendix \@ref{model-dev}.
+As the modeling details are of secondary importance, they are available in Appendix \ref{model-dev}.
 The questions that we ask of the model are as follows:
@@
 
-Before we dive into the answering questions, we provide some background on model interpretation techniques.
-As the literature is quite sizeable and fast moving, we do not attempt a full taxomony of techniques.
-The goal is to orient ourselves among broad categories of techniques, so we can map them to tasks indicated by the questions being asked.
-Model interpretation techniques can be categorized across two dimensions: intrinsic vs.
+Before we dive into the answering questions, we present a brief taxonomy of ML interpretation techniques.
+Rather than attempting an exhaustive classification, the goal is to orient ourselves among broad categories of techniques, so we can map them to tasks indicated by the questions being asked.
+For our purposes, model interpretation techniques can be categorized across two dimensions: intrinsic vs.
 post-hoc and global vs.
@@
 For algorithms that produce models with complex structure that do not lend themselves easily to intrinsic exploration, we can appeal to post-hoc techniques.
-This class of techniques interrogate the model by presenting it with data for and observing the prediction behavior of the model.
+This class of techniques interrogate the model by presenting it with data for scoring and observing the prediction behavior of the model.
 These techniques are concerned with only the inputs and outputs, and hence are *agnostic* of the model itself, which means they can also be applied to simple models.
@@
 In our case study, questions 1 and 2 are associated with this category.
-On the other hand, question 3 pertains to an individual prediction, which would fall in the local interpretation category.
+On the other hand, question 3 pertains to an individual prediction, which would fall in the local, or instance-level, category.
 In addition to individual variable attribution, we can also inquire about what would happen to the current predicted response if we were to perturb specific predictor variables.
@@
 The gist of the technique is as follows: to see how important a variable is, we make predictions without it and see how much worse off we are in terms of accuracy.
-One way to achieve this would be to re-fit the model many times (as many times as the number of variables.) However, this may be intractable with lengthy model training times or large number of variables, so a more popular approach is to instead keep the same fitted model but permute the values of each predictor.
+One way to achieve this would be to re-fit the model many times (as many times as the number of variables.) However, this may be intractable with lengthy model training times or large numbers of variables, so a more popular approach is to instead keep the same fitted model but permute the values of each predictor.
 
-More formally, let $y$ denote the vector of responses, $X$ denote the matrix of predictor variables, $\widehat{f}$ denote the fitted model, and $L = \mathcal{L}(\widehat{f}(X), y)$ denote the value of the loss function, which is mean squared error in the case of regression.
+More formally, let $y$ denote the vector of responses, $X$ denote the matrix of predictor variables, $\widehat{f}$ denote the fitted model, and $L = \mathcal{L}(\widehat{f}(X), y)$, where $\widehat{f}$ applies to $X$ rowwise, denote the value of the loss function, which is mean squared error in the case of regression.
 Now, if $\widetilde{X}^{j}$ denotes the predictor matrix where the $j$th variable has been permutated, then we can compute the loss with the permuted dataset as $L^{-j} = \mathcal{L}(\widehat{f}(\widetilde{X}^{j}), y)$.
-With this, we define the absolute variable importance $VI^{j}$ as $L^{-j} - L$ or the relative variable importance as $VI_{R}^{j} = \frac{L^{-j}}{L}$.
+With this, we define the  variable importance $VI^{j}$ as $L^{-j} - L$.
 
@@
 
-where $N$ is the number of records in our dataset.
+where $N$ is the number of records in the dataset.
 
@@
 We see that the average predicted loss cost decreases with vehicle age until the latter is around 18.
-Note that the ticks on the $x$-axis tell us that the data is quite thin for vehicle age greater than 18, so the apparent upward trend to the right is driven by just a few data points.
+Note that the rug on the $x$-axis tell us that the data is quite thin for vehicle age greater than 18, so the apparent upward trend to the right is driven by just a few data points.
 With this in mind, we can conclude that as cars get older, we expect the loss cost predictions generated by our model to decrease.
@@
 As before, we consider how we would approach the question for linear models.
-For a GLM with a log link common in ratemaking applications, for example, we simply start with the base rate, then the exponentiated coefficients would have multiplicative effects on the final rate.
+For a GLM with a log link common in ratemaking applications, for example, we start with the base rate, then the exponentiated coefficients would have multiplicative effects on the final rate.
 Similar to the previous examples, for ML models in general we do not have directly interpretable weights.
@@
 
-\begin{equation}
+\begin{equation}\label{eq:contrib-decompose}
 \widehat{f}(x_* ) = v_0 + \sum_{j=1}^p v(j, x_*),
@@
 
-where $v(j, x_*)$ denotes the contribution of the $j$th variable in instance $x_*$, defined as
+where $v_0$ denotes the average model response, and $v(j, x_*)$ denotes the contribution of the $j$th variable in instance $x_*$, defined as
 
@@
 & v(j, x_* ) = E_X[\widehat{f}(X) | X^1 = x_* ^1, \dots, X^j = x_* ^j] \nonumber\\
-& \quad - E_X[\widehat{f}(X^1 = x_* ^1, \dots, X^{j-1} = x_*^{j - 1})].
+& \quad - E_X[\widehat{f}(X) | X^1 = x_* ^1, \dots, X^{j-1} = x_*^{j - 1})].
 \end{align}
@@
 
-In Figure \@ref(fig:breakdown) we exhibit a waterfall plot of variable contributions.
-The "intercept" value denotes the average model prediction.
+In Figure \@ref(fig:breakdown), we exhibit a waterfall plot of variable contributions.
+The "intercept" value denotes the average model prediction and represents the $v_0$ term in Equation \@ref(eq:contrib-decompose).
 The predicted loss cost for this particular policyholder is slightly less than average; the characteristics that makes this policyholder more risky are the fact that he is a male between the ages of 18 and 25; counteracting the risky driver characteristics are the vehicle properties: it is a GM vehicle built domestically and is seven years old.

### 2019-10-25 manuscript/manuscript.Rmd
add things to watch out for

@@
 In our particular example, we see that the "make" variable contributes most to the accuracy of the model with "sex" contributing the least.
-Note that these measures do not provide information regarding the directional sensitivity of the predictors on the response.
 This provides a way for the audience to quickly glance at the most relevant variables, and ask further question as necessary.
@@
 
-```{r fi, out.width='100%', fig.align = "center", fig.cap="Permutation feature importances for ML model.", echo = FALSE}
+Note that these measures do not provide information regarding the directional sensitivity of the predictors on the response.
+Also, when there are correlated variables, one should be careful about interpretation, as the result may be biased by unrealistic records in the permuted dataset.
+Another ramification of a group correlated variables is that their inclusion may cause each to appear less important than if only one is included in the model.
+
+```{r fi, out.width='100%', fig.align = "center", fig.cap="Permutation feature importances for the neural network model.", echo = FALSE}
 knitr::include_graphics("figures/fi-plot.png")
@@
 
-```{r pdp, out.width='70%', fig.align = "center", fig.cap="Partial dependence plot for ML model.", echo = FALSE,  fig.env='figure*', fig.pos='h'}
+```{r pdp, out.width='70%', fig.align = "center", fig.cap="Partial dependence plot for the neural network model.", echo = FALSE,  fig.env='figure*', fig.pos='h'}
 knitr::include_graphics("figures/pdp-plot.png")
@@
 
-```{r breakdown, out.width='80%', fig.align = "center", fig.cap="Variable contribution plot for ML model.", echo = FALSE, fig.env = 'figure*', fig.pos = 'h'}
+```{r breakdown, out.width='80%', fig.align = "center", fig.cap="Variable contribution plot for the neural network model.", echo = FALSE, fig.env = 'figure*', fig.pos = 'h'}
 knitr::include_graphics("figures/breakdown-plot.png")
@@
 During the model development stage, PDP can also be used as a reasonableness test for candidate models by identifying unexpected relationships for the analyst to investigate.
+
+As with permutation feature importance, one should be careful when interpreting PDP when there are strongly correlated variables.
+Since we average over the marginal distribution of the rest of the variables, we may take into account unrealistic data (e.g.
+high vehicle age for a model that is brand new).
 
@@
 
+As noted earlier, the ordering of variables has an impact on the contributions calculated, especially for models that are non-additive, which could cause inconsistent explanations.
+There are several approaches to ameliorate this phenomenon, including selecting variables with the largest contributions first, including interactions terms, and averaging over possible orderings; these are discussed further in @biecekPMVEE and its references.
+
 ## Other Techniques
 
-(TODO: briefly discuss other stuff no covered)
+(TODO: briefly discuss other stuff not covered)
 

### 2019-10-26 manuscript/manuscript.Rmd

@@
 As noted earlier, the ordering of variables has an impact on the contributions calculated, especially for models that are non-additive, which could cause inconsistent explanations.
-There are several approaches to ameliorate this phenomenon, including selecting variables with the largest contributions first, including interactions terms, and averaging over possible orderings; these are discussed further in @biecekPMVEE and its references.
+There are several approaches to ameliorate this phenomenon, including selecting variables with the largest contributions first, including interactions terms, and averaging over possible orderings.
+The last of these ideas is implemented by @NIPS2017_7062 using Shapley values from cooperative game theory, and is referred to as Shapley additive explanations (SHAP).
+These approaches are discussed further in @biecekPMVEE and its references.
 
@@
 
-(TODO: briefly discuss other stuff not covered)
+In this paper, we demonstrate just a few model-agnostic ML interpretation techniques.
+These represent a small subset of existing techniques, each of which have additional variations.
+In the remainder of this section, we point out a few common techniques not covered in our case study.
+
+Individual conditional expectation (ICE) plots disaggregate PDPs into their instance-level components for a more granular view into predictor sensitivities [@goldstein2015peeking].
+To accommodate correlated variables in PDP, accumulated local effect (ALE) plots computes expected changes in model response over the conditional, rather than marginal, distribution of the other variables [@apley2016visualizing].
+
+Local interpretable model-agnostic explanations (LIME) [@ribeiro2016should] builds simple surrogate models using model predictions, with higher training weights given to the point of interest, in effect replacing the complex ML model with an easily interpretable linear regressions or decision trees in neighborhoods of specific points for the purpose of explanation.
+Taking the concept further, one can also train a global surrogate model across the entire domain of interest.
 

### 2019-10-28 manuscript/manuscript.Rmd

@@
 abstract: |
-  Machine learning methods have garnered increasing interest among actuaries in recent years.
-However, their adoption by practitioners has been limited, partly due to the lack of transparency of these methods, as compared to generalized linear models.
-In this paper, we discuss the need for model interpretability in property & casualty insurance ratemaking, propose a framework for explaining models, and present a case study to illustrate the framework.
+  Machine learning methods have garnered increasing interest among actuaries in
+  recent years. However, their adoption by practitioners has been limited, partly
+  due to the lack of transparency of these methods, as compared to generalized 
+  linear models. In this paper, we discuss the need for model interpretability in
+  property & casualty insurance ratemaking, propose a framework for explaining 
+  models, and present a case study to illustrate the framework.
 
@@
 
-Risk classification for property & casualty (P&C) insurance rating has traditionally been done with one-way, or univariate, analysis techniques.
-In recent years, many insurers have moved towards using generalized linear models (GLM), a multivariate predictive modeling technique, which addresses many shortcomings of univariate approaches, and is currently considered the gold standard in insurance risk classification.
-At the same time, machine learning (ML) techniques such as deep neural networks have gained popularity in many industries due to their superior predictive performance over linear models [@lecunDeepLearning2015].
-In fact, there is a fast growing body of literature on applying ML to P&C reserving [@kuoDeepTriangleDeep2018; @wuthrichMachineLearning2018; @gabrielliNeuralNetwork2019a; @gabrielliNeuralNetwork2019].
-However, these ML techniques, often considered to be completely “black box”, have been less successful in gaining adoption in pricing, which is a regulated discipline and requires a certain amount of transparency in models.
+Risk classification for property & casualty (P&C) insurance rating has traditionally
+been done with one-way, or univariate, analysis techniques. In recent years, many
+insurers have moved towards using generalized linear models (GLM), a multivariate
+predictive modeling technique, which addresses many shortcomings of univariate
+approaches, and is currently considered the gold standard in insurance risk
+classification. At the same time, machine learning (ML) techniques such as deep 
+neural networks have gained popularity in many industries due to their superior
+predictive performance over linear models [@lecunDeepLearning2015]. In fact, there
+is a fast growing body of literature on applying ML to P&C reserving
+[@kuoDeepTriangleDeep2018; @wuthrichMachineLearning2018; @gabrielliNeuralNetwork2019a;
+@gabrielliNeuralNetwork2019]. However, these ML techniques, often considered to be
+completely “black box”, have been less successful in gaining adoption in pricing, 
+which is a regulated discipline and requires a certain amount of transparency in models.
 
-If insurers can gain more insight into how ML models behave in risk classification contexts, it would increase their ability to reassure regulators and the public that accepted ratemaking principles are met.
-Being able to charge more accurate premiums would, in turn, make the risk transfer system more efficient and contribute to the betterment of society.
-In this paper, we aim to take a step towards liberating actuaries from the confines of linear models in pricing projects, by proposing a framework for explaining ML models for ratemaking that regulators, practitioners, and researchers in actuarial science can build upon.
+If insurers can gain more insight into how ML models behave in risk classification
+contexts, it would increase their ability to reassure regulators and the public 
+that accepted ratemaking principles are met. Being able to charge more accurate 
+premiums would, in turn, make the risk transfer system more efficient and contribute
+to the betterment of society. In this paper, we aim to take a step towards liberating 
+actuaries from the confines of linear models in pricing projects, by proposing a 
+framework for explaining ML models for ratemaking that regulators, practitioners,
+and researchers in actuarial science can build upon.
 
-The rest of this paper is organized as follows: Section \ref{ratemaking} provides an overview of P&C ratemaking, Section \ref{need} discusses the importance of interpretation, Section \ref{interpretability} discusses model interpretability in the context of ratemaking and proposes specific tasks for model explanation, Section \ref{application} describes current model interpretation techniques and applies them to the tasks defined in the previous section, and Section \ref{conclusion} concludes.
+The rest of this paper is organized as follows: Section \ref{ratemaking} provides
+an overview of P&C ratemaking, Section \ref{need} discusses the importance of 
+interpretation, Section \ref{interpretability} discusses model interpretability 
+in the context of ratemaking and proposes specific tasks for model explanation,
+Section \ref{application} describes current model interpretation techniques and 
+applies them to the tasks defined in the previous section, and Section 
+\ref{conclusion} concludes.
 
@@
 Early classification ratemaking procedures were typically univariate in nature.
-For example, @lange_1966 notes that (at that time) most major lines of insurance used univariate methods based around the same principle: distributing an overall indication to territorial relativities or classification relativities based on the extent to which they deviated from the average experience.
+For example, @lange_1966 notes that (at that time) most major lines of insurance 
+used univariate methods based around the same principle: distributing an overall 
+indication to territorial relativities or classification relativities based on 
+the extent to which they deviated from the average experience.
 
-@bailey_simon_1960 introduced minimum bias methods, which were expanded throughout the 60s, 70s, and 80s.
-As computing power developed, minimum bias began to give away to GLMs, with papers such as @brown_1988 and @mildenhall_1999 bridging the gap between the methods.
+@bailey_simon_1960 introduced minimum bias methods, which were expanded 
+throughout the 60s, 70s, and 80s. As computing power developed, minimum bias 
+began to give away to GLMs, with papers such as @brown_1988 and @mildenhall_1999 
+bridging the gap between the methods.
 
-Arguably, GLMs predate minimum bias procedures by a significant margin.
-The term was coined by @nelder_wedderburn_1972, but generalizations of least squares linear regression date back at least to the 1930s.
-Like minimum bias methods, GLMs did not become mainstream in actuarial science for some time.
-For example, the syllabus of basic education does not seem to include any mention of GLMs prior to @brown_1988 in the 1990 syllabus for basic education.
-From there, GLMs seem to have received only passing mention until 2006 with the introduction of @anderson_2005 to the syllabus.
+Arguably, GLMs predate minimum bias procedures by a significant margin. The term 
+was coined by @nelder_wedderburn_1972, but generalizations of least squares 
+linear regression date back at least to the 1930s. Like minimum bias methods, 
+GLMs did not become mainstream in actuarial science for some time. For example, 
+the syllabus of basic education does not seem to include any mention of GLMs 
+prior to @brown_1988 in the 1990 syllabus for basic education. From there, GLMs 
+seem to have received only passing mention until 2006 with the introduction of 
+@anderson_2005 to the syllabus.
 
@@
 
-Paralelling the development of GLM was the development of machine learning algorithms throughout the middle part of the 20th century.
-Detailed histories of machine learning may be found in sources such as @nilsson_2009 and @wang_raj_2017.
-Consistent with GLMs, machine learning was relatively unpopular in actuarial science until the last ten years as computing power has become cheaper and more easily available and as machine learning software packages have obviated the need for developing analyses from scratch each time an analysis is performed.
-Due to the breadth of machine learning as a field, it is difficult to identify the first time it entered the CAS syllabus; however, cluster analysis (in the form of k-means) seems to have been first included in 2011 with @robertson_2009.
-More recently, the MAS-I and MAS-II exams introduced in 2018 have included machine learning explicitly.
+Paralelling the development of GLM was the development of machine learning 
+algorithms throughout the middle part of the 20th century. Detailed histories 
+of machine learning may be found in sources such as @nilsson_2009 and 
+@wang_raj_2017. Consistent with GLMs, machine learning was relatively unpopular 
+in actuarial science until the last ten years as computing power has become cheaper
+and more easily available and as machine learning software packages have obviated
+the need for developing analyses from scratch each time an analysis is performed.
+Due to the breadth of machine learning as a field, it is difficult to identify 
+the first time it entered the CAS syllabus; however, cluster analysis (in the 
+form of k-means) seems to have been first included in 2011 with @robertson_2009.
+More recently, the MAS-I and MAS-II exams introduced in 2018 have included 
+machine learning explicitly.
 
-Within the area of ratemaking, machine learning is still in its infancy.
-A significant portion of machine learning applications to ratemaking has been in the context of automobile telematics, such as @gao_2018, @gao_2018_2, @gao_2019, @roel_2018, or @wuthrich_2017.
-Presumably this focus has been a result of the high-dimensionality and complexity of telematics data, making it a field in which the unique abilities of machine learning techniques give a clear advantage over traditional approaches.
+Within the area of ratemaking, machine learning is still in its infancy. A 
+significant portion of machine learning applications to ratemaking has been in 
+the context of automobile telematics, such as @gao_2018, @gao_2018_2, @gao_2019,
+@roel_2018, or @wuthrich_2017. Presumably this focus has been a result of the 
+high-dimensionality and complexity of telematics data, making it a field in 
+which the unique abilities of machine learning techniques give a clear 
+advantage over traditional approaches.
 
-Outside of telematics, @yang_2018 uses a gradient tree-boosting approach to capture non-linearities that would be a challenge for GLMs.
-@henckaerts_2018 makes use of generalized additive models (GAM) to improve predictions of GLMs.
-Many researchers, in an apparent effort to demonstrate the range of possibilities and advantages of machine learning, have approached the topic by comparing many different machine learning algorithms within a single study, such as in @dugas_2003, @noll_2018, @spedicato_2018.
-These studies make use of such varied techniques as regression trees, boosting machines, support vector machines, and neural networks.
+Outside of telematics, @yang_2018 uses a gradient tree-boosting approach to 
+capture non-linearities that would be a challenge for GLMs. @henckaerts_2018 
+makes use of generalized additive models (GAM) to improve predictions of GLMs.
+Many researchers, in an apparent effort to demonstrate the range of possibilities
+and advantages of machine learning, have approached the topic by comparing many
+different machine learning algorithms within a single study, such as in
+@dugas_2003, @noll_2018, @spedicato_2018. These studies make use of such varied
+techniques as regression trees, boosting machines, support vector machines, and 
+neural networks.
 
@@
 
-Regardless of the method employed for determining this risk of various classifications, the actual process of setting rate relativities typically involves some variation of the following steps:
+Regardless of the method employed for determining this risk of various 
+classifications, the actual process of setting rate relativities typically 
+involves some variation of the following steps:
 
@@
 2. Prepare data for analysis
-3. Perform analysis on the data, employing desired method or methods to estimate needed rate relativites
+3. Perform analysis on the data, employing desired method or methods to estimate
+needed rate relativites
 4. Select final rate relativities based on rate indications
-5. Present rates to the reglator, including explanation of the steps followed to derive the rates
+5. Present rates to the reglator, including explanation of the steps followed to 
+derive the rates
 6. Answer questions from regulators regarding the method employed
 
-The focus of this paper is on steps 5 and 6.
-In particular, rate regulators are concerned with whether rates are inadequate, excessive, or unfairly discriminatory.
-In many states, rate filings that exceed certain thresholds for magnitude of rate changes or filings that make use of new or sophisticated predictive models may be subject to intense regulatory scrutiny.
-In these cases, it is necessary to be able to explain the results of the modeling process in a way that is understandable without sacrificing statistical rigor.
+The focus of this paper is on steps 5 and 6. In particular, rate regulators are 
+concerned with whether rates are inadequate, excessive, or unfairly 
+discriminatory. In many states, rate filings that exceed certain thresholds for
+magnitude of rate changes or filings that make use of new or sophisticated 
+predictive models may be subject to intense regulatory scrutiny. In these cases,
+it is necessary to be able to explain the results of the modeling process in a 
+way that is understandable without sacrificing statistical rigor.
 
-It should be noted that communicating results is not simply a method of passing regulatory muster.
-Generating interpretable modeling output is an important - even essential - facet of model checking.
-Therefore, the techniques discussed in this paper may be viewed from the lens of providing useful information to regulators, but they should also be considered as part of a thorough vetting of any rating model.
+It should be noted that communicating results is not simply a method of passing 
+regulatory muster. Generating interpretable modeling output is an important - 
+even essential - facet of model checking. Therefore, the techniques discussed 
+in this paper may be viewed from the lens of providing useful information to 
+regulators, but they should also be considered as part of a thorough vetting of 
+any rating model.
 
@@
 
-Within the actuarial profession, Actuarial Standard of Practice 41 ("Actuarial Communications") notes that "...another actuary qualified in the same practice area [should be able to] make an objective appraisal of the reasonableness of the actuary's work as presented in the actuarial report." [@asop_41] Underlying this requirement is an assumption that the hypothetical other actuary qualified in the same practice area is adequately familiar with the relevant techniques employed.
-Although the syllabus of basic education is constantly changing, there has at times been an assumption that all techniques and assumptions that have ever been a part of the syllabus of basic education needn't be explained from first principles in general actuarial communications, and that an actuary practicing in the same field should be able to make an objective appraisal of the results from the methods found in the syllabus.
-[This can be supported by reference to the edits to ASOP 38 over time - originally it was designed to be about all models, but they revised it to be about only models that incorporate specialized knowledge outside of the actuary's area of expertise...
-do we need this citation, though?] This is notable because, beginning with the introduction of the MAS-I and MAS-II examination in July of 2018, several machine learning models were formally included in the syllabus of basic education.
-These exams cover a wide range of topics, such as splines, clustering algorithms, decision trees, boosting, and principle components analysis [@cas_syllabus_2018].
+Within the actuarial profession, Actuarial Standard of Practice 41 ("Actuarial 
+Communications") notes that "...another actuary qualified in the same practice 
+area [should be able to] make an objective appraisal of the reasonableness of 
+the actuary's work as presented in the actuarial report." [@asop_41] Underlying 
+this requirement is an assumption that the hypothetical other actuary qualified 
+in the same practice area is adequately familiar with the relevant techniques 
+employed. Although the syllabus of basic education is constantly changing, there
+has at times been an assumption that all techniques and assumptions that have 
+ever been a part of the syllabus of basic education needn't be explained from 
+first principles in general actuarial communications, and that an actuary 
+practicing in the same field should be able to make an objective appraisal of 
+the results from the methods found in the syllabus. [This can be supported by 
+reference to the edits to ASOP 38 over time - originally it was designed to be 
+about all models, but they revised it to be about only models that incorporate 
+specialized knowledge outside of the actuary's area of expertise... do we need 
+this citation, though?] This is notable because, beginning with the introduction
+of the MAS-I and MAS-II examination in July of 2018, several machine learning 
+models were formally included in the syllabus of basic education. These exams 
+cover a wide range of topics, such as splines, clustering algorithms, decision 
+trees, boosting, and principle components analysis [@cas_syllabus_2018].
 
-Nevertheless, machine learning poses something of a special challenge for ASOP 41 for several reasons.
-Machine learning models can be very ad hoc compared to traditional statistical models.
-Because many machine learning models are deterministic, they may not admit of standard metrics for model comparison (e.g., it's not straightforward to calculate an AIC over a neural network).
-In addition, machine learning methods are often combined into ensembles that may not be easily separated and that may, as a collection, cease to resemble a single standard version of a model.
-Complicating matters still further, machine learning models can be "black boxes" insofar as the final form of response curve cannot be easily predicted and may depend heavily on the available data (which may not, in turn, be available to the reviewer).
+Nevertheless, machine learning poses something of a special challenge for ASOP 
+41 for several reasons. Machine learning models can be very ad hoc compared to 
+traditional statistical models. Because many machine learning models are 
+deterministic, they may not admit of standard metrics for model comparison 
+(e.g., it's not straightforward to calculate an AIC over a neural network). 
+In addition, machine learning methods are often combined into ensembles that 
+may not be easily separated and that may, as a collection, cease to resemble 
+a single standard version of a model. Complicating matters still further, 
+machine learning models can be "black boxes" insofar as the final form of 
+response curve cannot be easily predicted and may depend heavily on the 
+available data (which may not, in turn, be available to the reviewer).
 
-This last item raises a final interesting issue.
-GLMs and their ilk are often fitted using one of a handful of standard and well-understood approaches (e.g., maximum likelihood estimation).
-However, this is not possible in general with machine learning models, as machine learning algorithms often use loss surfaces that are very complex such that it may not be feasible to calculate the global minimum of the surface.
-Certainly, closed form representations of the loss surfaces are not generally available.
-For this reason, the training phase of a machine learning model is, in many ways, just as important to one's understanding as the model form and the data on which the model is fitted.
-Because the final model result is inseparable from these three components (training method, model form, and data), it is not generally adequate to just know the method employed to make an objective appraisal of the reasonableness of the model.
-More information is necessary.
+This last item raises a final interesting issue. GLMs and their ilk are often 
+fitted using one of a handful of standard and well-understood approaches (e.g.,
+maximum likelihood estimation). However, this is not possible in general with 
+machine learning models, as machine learning algorithms often use loss surfaces 
+that are very complex such that it may not be feasible to calculate the global 
+minimum of the surface. Certainly, closed form representations of the loss 
+surfaces are not generally available. For this reason, the training phase of a 
+machine learning model is, in many ways, just as important to one's 
+understanding as the model form and the data on which the model is fitted. 
+Because the final model result is inseparable from these three components 
+(training method, model form, and data), it is not generally adequate to just
+know the method employed to make an objective appraisal of the reasonableness 
+of the model. More information is necessary.
 
-Of course, these comments only apply to the actuarial profession.
-Outside of the actuarial profession, communication of results may be more challenging.
-A 2017 survey conducted by the Casualty Actuarial and Statistical Task Force of the National Association of Insurance Commissioners found that the plurality of responding regulators identified "Filing complexity and/or a lack of resources or expertise" as a key challenge that impedes their ability to review GLMs or other predictive models [@naic_summer_2017].
-Given that machine learning algorithms are generally regarded as more complex than GLMs, this implies that the challenge of communicating machine learning model results is significant.
+Of course, these comments only apply to the actuarial profession. Outside of 
+the actuarial profession, communication of results may be more challenging. A 
+2017 survey conducted by the Casualty Actuarial and Statistical Task Force of 
+the National Association of Insurance Commissioners found that the plurality of
+responding regulators identified "Filing complexity and/or a lack of resources 
+or expertise" as a key challenge that impedes their ability to review GLMs or 
+other predictive models [@naic_summer_2017]. Given that machine learning 
+algorithms are generally regarded as more complex than GLMs, this implies that
+the challenge of communicating machine learning model results is significant.
 
-In response to the same survey, 33 state regulators noted that it would be helpful or very helpful for the NAIC to develop information and tools to assist in reviewing rate filings based on GLMs, and 34 noted that it would be helpful to develop similar items to assist in reviewing "Other Advanced Modeling Techniques." One outgrowth of this need was the development of a white paper, @naic_white_paper, on best practices for regulatory review of predictive models.
-The white paper focuses on review of GLMs, particularly with respect to private passenger automobile and homeowners' insurance.
-Some of the guidance offered in this regard is therefore not strictly applicable to the review of machine learning models.
-For example, as previously noted, p-values are not a concept that translates well to deterministic machine learning algorithms.
-However, among the guidance applicable to machine learning algorithms are the following:
+In response to the same survey, 33 state regulators noted that it would be 
+helpful or very helpful for the NAIC to develop information and tools to assist
+in reviewing rate filings based on GLMs, and 34 noted that it would be helpful 
+to develop similar items to assist in reviewing "Other Advanced Modeling 
+Techniques." One outgrowth of this need was the development of a white paper,
+@naic_white_paper, on best practices for regulatory review of predictive models.
+The white paper focuses on review of GLMs, particularly with respect to private
+passenger automobile and homeowners' insurance. Some of the guidance offered in
+this regard is therefore not strictly applicable to the review of machine 
+learning models. For example, as previously noted, p-values are not a concept
+that translates well to deterministic machine learning algorithms. However,
+among the guidance applicable to machine learning algorithms are the following:
 
-* Determine the extent the model causes premium disruption for individual policyholders, and how the insurer will
+* Determine the extent the model causes premium disruption for individual 
+policyholders, and how the insurer will
 explain the disruption to individual consumers that inquire about it.
-* Determine that individual input characteristics to a predictive model are related to the expected loss or expense
-differences in risk. Each input characteristic should have an intuitive or demonstrable actual relationship to expected loss or expense.
-* Determine that individual outputs from a predictive model and their associated selected relativities are not unfairly discriminatory.
+* Determine that individual input characteristics to a predictive model are 
+related to the expected loss or expense
+differences in risk. Each input characteristic should have an intuitive or 
+demonstrable actual relationship to expected loss or expense.
+* Determine that individual outputs from a predictive model and their associated
+selected relativities are not unfairly discriminatory.
 
-These items are by no means exhaustive, but they pertain to the concept of model interpretability for ratemaking that we develop next.
+These items are by no means exhaustive, but they pertain to the concept of model
+interpretability for ratemaking that we develop next.
 
@@
 
-In this section, we attempt to develop a working definition of interpretability for ratemaking applications.
-While we will not provide a comprehensive survey of the prolific and fast evolving ML interpretability literature, we draw from it as appropriate in setting the stage for our discussion.
-Even among researchers in the subject, there is not a consensus on the definition of interpretability; here are a few from frequently cited papers:
+In this section, we attempt to develop a working definition of interpretability
+for ratemaking applications. While we will not provide a comprehensive survey of
+the prolific and fast evolving ML interpretability literature, we draw from it 
+as appropriate in setting the stage for our discussion. Even among researchers 
+in the subject, there is not a consensus on the definition of interpretability;
+here are a few from frequently cited papers:
 
-- Ability to explain or to present in understandable terms to a human [@doshi-velezRigorousScience2017];
-- The degree to which an observer can understand the cause of a decision [@biranExplanationJustification2017]; and
-- A method is interpretable if a user can correctly and efficiently predict the method’s results [@kimExamplesAre2016].
+- Ability to explain or to present in understandable terms to a human 
+[@doshi-velezRigorousScience2017];
+- The degree to which an observer can understand the cause of a decision 
+[@biranExplanationJustification2017]; and
+- A method is interpretable if a user can correctly and efficiently predict the 
+method’s results [@kimExamplesAre2016].
 
 We motivate our discussion by considering several aspects of interpretability. 
-As we proceed through the points below, we aim to arrive at a more scoped and relevant definition of what it means to attain interpretability in a pricing model.
-In the remainder of this section, we clarify a couple concepts regarding interpretable *classes* of models and and the computational transparency of ML models, outline findings on the social aspects of ML explanations most relevant to our use case, then discuss a potential framework for implementing ML interpretability in practice.
+As we proceed through the points below, we aim to arrive at a more scoped and 
+relevant definition of what it means to attain interpretability in a pricing
+model. In the remainder of this section, we clarify a couple concepts regarding 
+interpretable *classes* of models and and the computational transparency of ML
+models, outline findings on the social aspects of ML explanations most relevant
+to our use case, then discuss a potential framework for implementing ML 
+interpretability in practice.
 
@@
 
-In the actuarial science literature, the GLM is probably the most oft-cited example of an easily interpretable model.
-Given a set of inputs, we can easily reason about what the output of the model is.
-As an illustrative example, consider a claim severity model with driver age, sex, and vehicle age as features; assuming a log link function and letting $Y$ denotes the response, we have
+In the actuarial science literature, the GLM is probably the most oft-cited
+example of an easily interpretable model. Given a set of inputs, we can easily
+reason about what the output of the model is. As an illustrative example, 
+consider a claim severity model with driver age, sex, and vehicle age as 
+features; assuming a log link function and letting $Y$ denotes the response, we
+have
 
@@
 
-Here, we can tell, for example, what the model would predict for the expected severity if we were to increase age by a certain amount, *all else being equal*, because the relationship between the feature and the response is simply multiplication by the coefficient $\beta_1$ and applying the inverse link function.
+Here, we can tell, for example, what the model would predict for the expected 
+severity if we were to increase age by a certain amount, *all else being equal*,
+because the relationship between the feature and the response is simply 
+multiplication by the coefficient $\beta_1$ and applying the inverse link
+function.
 
-Another commonly cited example of an interpretable model is a decision tree.
-An illustrative example is shown in Figure \@ref(fig:tree-plot1).
-Here, the prediction is arrived at by following a sequence of if-else decisions.
+Another commonly cited example of an interpretable model is a decision tree. An
+illustrative example is shown in Figure \@ref(fig:tree-plot1). Here, the 
+prediction is arrived at by following a sequence of if-else decisions.
 
@@
 
-Now, it is worth pointing out that, when declaring that GLMs or decision trees are interpretable models, we are implicitly assuming that we are considering only a handful  of features.
-In fact, the ease at which we can reason about a model declines as the number of features, transformations of them, and interactions increase, as in the following (somewhat pathological) example:
+Now, it is worth pointing out that, when declaring that GLMs or decision trees 
+are interpretable models, we are implicitly assuming that we are considering 
+only a handful  of features. In fact, the ease at which we can reason about a
+model declines as the number of features, transformations of them, and
+interactions increase, as in the following (somewhat pathological) example:
 
@@
 
-Similarly, one can see that in Figure \@ref(fig:tree-plot2), larger trees are tough to reason about.
-In other words, even when working within the framework of an "interpretable" class of models, we may still end up with something that many would consider "black box."
+Similarly, one can see that in Figure \@ref(fig:tree-plot2), larger trees are
+tough to reason about. In other words, even when working within the framework
+of an "interpretable" class of models, we may still end up with something that
+many would consider "black box."
 
@@
 
-Another occasional misconception is that we have no visibility into how some ML models *compute* predictions, which renders them uninterpretable.
-Outside of proprietary algorithms, all common ML models, including neural networks, gradient boosted trees, and random forests, are well studied and have large bodies of literature documenting their inner workings.
-As an example, a fitted feedforward neural network is simply a composition of linear transformations followed by nonlinear activation functions.
-As in Equation \ref{eq:badglm}, one can write down the mathematical equation for calculating the prediction given some inputs, but it may be difficult for a human to reason about it.
-We show later that we can still provide explanations of completely "black box" models, but is important to note that ML model predictions are still governed by mathematical rules, and are deterministic in most cases.
+Another occasional misconception is that we have no visibility into how some ML
+models *compute* predictions, which renders them uninterpretable. Outside of 
+proprietary algorithms, all common ML models, including neural networks, 
+gradient boosted trees, and random forests, are well studied and have large
+bodies of literature documenting their inner workings. As an example, a fitted
+feedforward neural network is simply a composition of linear transformations
+followed by nonlinear activation functions. As in Equation \ref{eq:badglm}, one
+can write down the mathematical equation for calculating the prediction given
+some inputs, but it may be difficult for a human to reason about it. We show
+later that we can still provide explanations of completely "black box" models,
+but is important to note that ML model predictions are still governed by
+mathematical rules, and are deterministic in most cases.
 
@@
 
-Proposed by @hiltonConversationalProcesses1990 and interpreted by @millerExplanationArtificial2017 in the context of ML, model explanations are conversations or social interactions.
-One consequence of this identification is that explanations need to be *relevant* to the *audience*.
-In developing, filing, and operationalizing a pricing model, one needs to accommodate a variety of stakeholders, each of whom has a different set of questions, assumptions, and technical capacity.
-First, there are internal stakeholders at the company, which includes management and underwriters.
-While some of the individuals in this audience may be technical, they are likely less familiar with predictive modeling techniques than the actuaries and data scientists who build the models.
-Next, we have the regulators, who may have limited resources to review the models, and will focus on a specific list of questions motivated by statute and public policy.
-Finally, we have the public, who has an interest (perhaps more so than the other parties) because their lives are directly affected by the premiums charged.
+Proposed by @hiltonConversationalProcesses1990 and interpreted by
+@millerExplanationArtificial2017 in the context of ML, model explanations
+are conversations or social interactions. One consequence of this identification
+is that explanations need to be *relevant* to the *audience*. In developing,
+filing, and operationalizing a pricing model, one needs to accommodate a variety
+of stakeholders, each of whom has a different set of questions, assumptions,
+and technical capacity. First, there are internal stakeholders at the company,
+which includes management and underwriters. While some of the individuals in
+this audience may be technical, they are likely less familiar with predictive
+modeling techniques than the actuaries and data scientists who build the models.
+Next, we have the regulators, who may have limited resources to review the
+models, and will focus on a specific list of questions motivated by statute
+and public policy. Finally, we have the public, who has an interest (perhaps
+more so than the other parties) because their lives are directly affected by
+the premiums charged.
 
-It is interesting to note that, the modelers, who are most familiar with the models, tend to be same people designing and communicating the explanations.
-This poses a challenge that @millerExplainableAI2017 calls "inmates running the asylum", where the modelers design explanations for *themselves* rather than the intended audience.
-For example, they may be interested in technical questions such as extrapolation behavior and shape the explanations accordingly, which may be lost on the prospective policyholder.
+It is interesting to note that, the modelers, who are most familiar with the
+models, tend to be same people designing and communicating the explanations.
+This poses a challenge that @millerExplainableAI2017 calls "inmates running
+the asylum", where the modelers design explanations for *themselves* rather
+than the intended audience. For example, they may be interested in technical
+questions such as extrapolation behavior and shape the explanations accordingly,
+which may be lost on the prospective policyholder.
 
-Another point outlined in Miller's survey [@millerExplanationArtificial2017] is that explanations are *contrastive*.
-In other words, people are often interested in not why something happened, but rather why it happened instead of something else.
-For example, policyholders might not care exactly how their auto premiums are computed, but would like to know why they are being charged more than their coworkers who drive similar vehicles.
-As an extension, policyholders may want to know what they can change in order to obtain lower premiums.
+Another point outlined in Miller's survey [@millerExplanationArtificial2017] is
+that explanations are *contrastive*. In other words, people are often interested
+in not why something happened, but rather why it happened instead of something
+else. For example, policyholders might not care exactly how their auto premiums
+are computed, but would like to know why they are being charged more than their
+coworkers who drive similar vehicles. As an extension, policyholders may want
+to know what they can change in order to obtain lower premiums.
 
@@
 
-With the above considerations in mind, we propose a potential framework for interpreting ML models for insurance pricing: the actuarial profession, in collaboration with regulators and representatives of the public, define a set of questions to be answered by explanations accompanying ML models, along with acceptance criteria and examples of successful explanations.
-In other words, interpretability for our purposes is defined as the ability of a model's explanations to answer the posed questions.
+With the above considerations in mind, we propose a potential framework for
+interpreting ML models for insurance pricing: the actuarial profession, in
+collaboration with regulators and representatives of the public, define a set
+of questions to be answered by explanations accompanying ML models, along
+with acceptance criteria and examples of successful explanations. In other
+words, interpretability for our purposes is defined as the ability of a model's
+explanations to answer the posed questions.
 
-These questions should relate to existing guidelines, such as those described in @naic_white_paper and outlined in Section \ref{need}, standards of practice, and regulation, and in fact should not be specific only to ML models.
-By standardizing on a set of questions, we reduce the burden of both companies and regulators; this is especially important for the latter, who are already resource constrained facing increasing variety of models being filed.
-This format should also be familiar to actuaries who are accustomed to adhering to specific guidelines in, for example, ASOPs.
-Like the ASOPs, We envision that these questions and guidelines will be continually updated to reflect feedback obtained and advances in research.
+These questions should relate to existing guidelines, such as those described in
+@naic_white_paper and outlined in Section \ref{need}, standards of practice, and
+regulation, and in fact should not be specific only to ML models. By
+standardizing on a set of questions, we reduce the burden of both companies and
+regulators; this is especially important for the latter, who are already
+resource constrained facing increasing variety of models being filed. This
+format should also be familiar to actuaries who are accustomed to adhering to
+specific guidelines in, for example, ASOPs. Like the ASOPs, We envision that
+these questions and guidelines will be continually updated to reflect feedback
+obtained and advances in research.
 
-While the realization of a set of such guidelines is an ambitious undertaking beyond the scope of this paper, we present in the next section a sample set of questions and techniques one can leverage to answer them.
-The goal of these case studies is twofold: to more concretely illustrate the proposed framework, and to expose the actuarial audience to modern ML interpretation techniques.
+While the realization of a set of such guidelines is an ambitious undertaking
+beyond the scope of this paper, we present in the next section a sample set of
+questions and techniques one can leverage to answer them. The goal of these case
+studies is twofold: to more concretely illustrate the proposed framework, and
+to expose the actuarial audience to modern ML interpretation techniques.
 
@@
 
-Now that we have established a framework for model interpretation in the form of asking and answering relevant questions, we demonstrate examples of such exchanges via an illustrative case study.
-Analytically, our starting point is a fitted deep neural network model for predicting loss costs.
-As the modeling details are of secondary importance, they are available in Appendix \ref{model-dev}.
-The questions that we ask of the model are as follows:
+Now that we have established a framework for model interpretation in the form of
+asking and answering relevant questions, we demonstrate examples of such
+exchanges via an illustrative case study. Analytically, our starting point is a
+fitted deep neural network model for predicting loss costs. As the modeling
+details are of secondary importance, they are available in Appendix
+\ref{model-dev}. The questions that we ask of the model are as follows:
 
@@
 2. How does the predicted loss cost change, on average, as we change an input?
-3. For a particular policyholder, how does each characteristic contribute to the loss cost prediction?
+3. For a particular policyholder, how does each characteristic contribute to
+the loss cost prediction?
 
-In our discussion, we adopt the organization of techniques and some notation presented in @molnar2018interpretable and @biecekPMVEE, which are comprehensive references on the most established ML interpretation techniques.
+In our discussion, we adopt the organization of techniques and some notation
+presented in @molnar2018interpretable and @biecekPMVEE, which are comprehensive
+references on the most established ML interpretation techniques.
 
@@
 
-Before we dive into the answering questions, we present a brief taxonomy of ML interpretation techniques.
-Rather than attempting an exhaustive classification, the goal is to orient ourselves among broad categories of techniques, so we can map them to tasks indicated by the questions being asked.
-For our purposes, model interpretation techniques can be categorized across two dimensions: intrinsic vs.
-post-hoc and global vs.
-local.
+Before we dive into the answering questions, we present a brief taxonomy of ML
+interpretation techniques. Rather than attempting an exhaustive classification,
+the goal is to orient ourselves among broad categories of techniques, so we can
+map them to tasks indicated by the questions being asked. For our purposes,
+model interpretation techniques can be categorized across two dimensions:
+intrinsic vs. post-hoc and global vs. local.
 
@@
 
-Intrinsic model interpretation draws conclusions from the structure of the fitted model and are what we typically associate with "interpretable" classes of models.
-This is only viable with models with simple structures, such as the sparse linear model and shallow decision tree we see in Section \ref{linear-models}, where we arrive at explanations by reading off parameter estimates or a few decision rules.
-For algorithms that produce models with complex structure that do not lend themselves easily to intrinsic exploration, we can appeal to post-hoc techniques.
-This class of techniques interrogate the model by presenting it with data for scoring and observing the prediction behavior of the model.
-These techniques are concerned with only the inputs and outputs, and hence are *agnostic* of the model itself, which means they can also be applied to simple models.
-Since most useful ML models have a level of complexity beyond the threshold of intrinsic interpretability, we focus on model-agnostic techniques in our case study.
-As we will see later on, the data that we present to the models are usually some perturbed variations of test data.
+Intrinsic model interpretation draws conclusions from the structure of the
+fitted model and are what we typically associate with "interpretable" classes
+of models. This is only viable with models with simple structures, such as the
+sparse linear model and shallow decision tree we see in Section
+\ref{linear-models}, where we arrive at explanations by reading off parameter
+estimates or a few decision rules. For algorithms that produce models
+with complex structure that do not lend themselves easily to intrinsic
+exploration, we can appeal to post-hoc techniques. This class of techniques
+interrogate the model by presenting it with data for scoring and observing the
+prediction behavior of the model. These techniques are concerned with only the
+inputs and outputs, and hence are *agnostic* of the model itself, which means
+they can also be applied to simple models. Since most useful ML models have a
+level of complexity beyond the threshold of intrinsic interpretability, we focus
+on model-agnostic techniques in our case study. As we will see later on, the
+data that we present to the models are usually some perturbed variations of test
+data.
 
@@
 
-Along the other dimension, we categorize model interpretations as global, or model-level, and local, or instance-level.
-The former class provides insights with respect to the model as a whole.
-Some examples of these eplanations include variable importances and sensitivities, on average, of the predicted response with respect to individual predictors.
-In our case study, questions 1 and 2 are associated with this category.
-On the other hand, question 3 pertains to an individual prediction, which would fall in the local, or instance-level, category.
-In addition to individual variable attribution, we can also inquire about what would happen to the current predicted response if we were to perturb specific predictor variables.
+Along the other dimension, we categorize model interpretations as global, or
+model-level, and local, or instance-level. The former class provides insights
+with respect to the model as a whole. Some examples of these eplanations include
+variable importances and sensitivities, on average, of the predicted response
+with respect to individual predictors. In our case study, questions 1 and 2 are
+associated with this category. On the other hand, question 3 pertains to an
+individual prediction, which would fall in the local, or instance-level,
+category. In addition to individual variable attribution, we can also inquire
+about what would happen to the current predicted response if we were to perturb
+specific predictor variables.
 
@@
 
-Having aligned the questions with the categories of interpretation techniques, we now introduce a selection of appropriate techniques to answer them.
+Having aligned the questions with the categories of interpretation techniques,
+we now introduce a selection of appropriate techniques to answer them.
 
@@
 
-For linear models and their generalizations, and some ML models, measures of variable importance can be obtained from the fitted model structure.
-In the case of GLMs, one can observe the magnitudes of the $t$-statistics, whereas for random forests, one can use out-of-bag errors [@breimanRandomForests2001].
-For more complex models, such as the neural network in our case study, we need to devise another approach.
+For linear models and their generalizations, and some ML models, measures of
+variable importance can be obtained from the fitted model structure. In the
+case of GLMs, one can observe the magnitudes of the $t$-statistics, whereas
+for random forests, one can use out-of-bag errors [@breimanRandomForests2001].
+For more complex models, such as the neural network in our case study, we need
+to devise another approach.
 
-We follow the methodology of permutation feature importance as described in @fisherAllModels2018, and utilize the notation introduced by @biecekPMVEE.
-The gist of the technique is as follows: to see how important a variable is, we make predictions without it and see how much worse off we are in terms of accuracy.
-One way to achieve this would be to re-fit the model many times (as many times as the number of variables.) However, this may be intractable with lengthy model training times or large numbers of variables, so a more popular approach is to instead keep the same fitted model but permute the values of each predictor.
+We follow the methodology of permutation feature importance as described in
+@fisherAllModels2018, and utilize the notation introduced by @biecekPMVEE.
+The gist of the technique is as follows: to see how important a variable is,
+we make predictions without it and see how much worse off we are in terms of
+accuracy. One way to achieve this would be to re-fit the model many times (as
+many times as the number of variables.) However, this may be intractable with
+lengthy model training times or large numbers of variables, so a more popular
+approach is to instead keep the same fitted model but permute the values of each
+predictor.
 
-More formally, let $y$ denote the vector of responses, $X$ denote the matrix of predictor variables, $\widehat{f}$ denote the fitted model, and $L = \mathcal{L}(\widehat{f}(X), y)$, where $\widehat{f}$ applies to $X$ rowwise, denote the value of the loss function, which is mean squared error in the case of regression.
-Now, if $\widetilde{X}^{j}$ denotes the predictor matrix where the $j$th variable has been permutated, then we can compute the loss with the permuted dataset as $L^{-j} = \mathcal{L}(\widehat{f}(\widetilde{X}^{j}), y)$.
-With this, we define the  variable importance $VI^{j}$ as $L^{-j} - L$.
+More formally, let $y$ denote the vector of responses, $X$ denote the matrix of
+predictor variables, $\widehat{f}$ denote the fitted model, and
+$L = \mathcal{L}(\widehat{f}(X), y)$, where $\widehat{f}$ applies to $X$
+rowwise, denote the value of the loss function, which is mean squared error
+in the case of regression. Now, if $\widetilde{X}^{j}$ denotes the predictor
+matrix where the $j$th variable has been permutated, then we can compute the
+loss with the permuted dataset as
+$L^{-j} = \mathcal{L}(\widehat{f}(\widetilde{X}^{j}), y)$. With this, we define
+the  variable importance $VI^{j}$ as $L^{-j} - L$.
 
-In Figure \@ref(fig:fi), we show a plot of variable importances.
-In our particular example, we see that the "make" variable contributes most to the accuracy of the model with "sex" contributing the least.
-This provides a way for the audience to quickly glance at the most relevant variables, and ask further question as necessary.
-For example, if a variable that can potentially be a proxy for a discriminatory variable appears as very important, more investigation would be warranted.
+In Figure \@ref(fig:fi), we show a plot of variable importances. In our
+particular example, we see that the "make" variable contributes most to the
+accuracy of the model with "sex" contributing the least. This provides a way
+for the audience to quickly glance at the most relevant variables, and ask
+further question as necessary. For example, if a variable that can potentially
+be a proxy for a discriminatory variable appears as very important, more
+investigation would be warranted.
 
-Note that these measures do not provide information regarding the directional sensitivity of the predictors on the response.
-Also, when there are correlated variables, one should be careful about interpretation, as the result may be biased by unrealistic records in the permuted dataset.
-Another ramification of a group correlated variables is that their inclusion may cause each to appear less important than if only one is included in the model.
+Note that these measures do not provide information regarding the directional
+sensitivity of the predictors on the response. Also, when there are correlated
+variables, one should be careful about interpretation, as the result may be
+biased by unrealistic records in the permuted dataset. Another ramification
+of a group correlated variables is that their inclusion may cause each to
+appear less important than if only one is included in the model.
 
@@
 
-For this question, we again consider first how it would be answered in the GLM setting.
-When the input predictor in question is continuous, we can answer the question by looking at the estimated coefficient, which provides the change in the response per unit change in the predictor (on the scale of the linear predictor).
-For non-parametric models and neural networks, where no coefficients are available, we can appeal to partial dependence plots (PDP), first proposed by @friedmanGreedyFunction2001 for gradient boosting machines (GBM).
+For this question, we again consider first how it would be answered in the GLM
+setting. When the input predictor in question is continuous, we can answer the
+question by looking at the estimated coefficient, which provides the change in
+the response per unit change in the predictor (on the scale of the linear
+predictor). For non-parametric models and neural networks, where no coefficients
+are available, we can appeal to partial dependence plots (PDP), first proposed
+by @friedmanGreedyFunction2001 for gradient boosting machines (GBM).
 
@@
 
-To describe PDP, we need to introduce some additional notation. Let $x^j$ denote the input variable of interest. Then we define the partial dependence function as
+To describe PDP, we need to introduce some additional notation. Let $x^j$ denote
+the input variable of interest. Then we define the partial dependence function as
 
@@
 
-where the expectation is taken over the distribution of the other predictor variables.
-In other words, we marginalize them out so we can focus on the relationship between the predicted response and the variable of interest.
+where the expectation is taken over the distribution of the other predictor
+variables. In other words, we marginalize them out so we can focus on the
+relationship between the predicted response and the variable of interest.
 Empirically, we estimate $h$ by
@@
 In Figure \@ref(fig:pdp), we exhibit the PDP for the "vehicle age" variable.
-We see that the average predicted loss cost decreases with vehicle age until the latter is around 18.
-Note that the rug on the $x$-axis tell us that the data is quite thin for vehicle age greater than 18, so the apparent upward trend to the right is driven by just a few data points.
-With this in mind, we can conclude that as cars get older, we expect the loss cost predictions generated by our model to decrease.
+We see that the average predicted loss cost decreases with vehicle age until
+the latter is around 18. Note that the rug on the $x$-axis tell us that the
+data is quite thin for vehicle age greater than 18, so the apparent upward
+trend to the right is driven by just a few data points. With this in mind,
+we can conclude that as cars get older, we expect the loss cost predictions
+generated by our model to decrease.
 
-The question posed here is particularly important for regulators, who would like to know whether each variable affects the prediction in the direction that is expected, based on intuition, experience, and existing models.
-During the model development stage, PDP can also be used as a reasonableness test for candidate models by identifying unexpected relationships for the analyst to investigate.
+The question posed here is particularly important for regulators, who would
+like to know whether each variable affects the prediction in the direction that
+is expected, based on intuition, experience, and existing models. During the
+model development stage, PDP can also be used as a reasonableness test for
+candidate models by identifying unexpected relationships for the analyst to
+investigate.
 
-As with permutation feature importance, one should be careful when interpreting PDP when there are strongly correlated variables.
-Since we average over the marginal distribution of the rest of the variables, we may take into account unrealistic data (e.g.
-high vehicle age for a model that is brand new).
+As with permutation feature importance, one should be careful when interpreting
+PDP when there are strongly correlated variables. Since we average over the
+marginal distribution of the rest of the variables, we may take into account
+unrealistic data (e.g. high vehicle age for a model that is brand new).
 
@@
 
-"For a particular policyholder, how does each characteristic contribute to the loss cost prediction?"
+"For a particular policyholder, how does each characteristic contribute to the
+loss cost prediction?"
 
-In the previous two examples, we look at model-level explanations; now we move on to one where we investigate one particular prediction instance.
-As before, we consider how we would approach the question for linear models.
-For a GLM with a log link common in ratemaking applications, for example, we start with the base rate, then the exponentiated coefficients would have multiplicative effects on the final rate.
-Similar to the previous examples, for ML models in general we do not have directly interpretable weights.
-Instead, one way to arrive at variable contributions is calculating the change in the expected model prediction for each predictor, conditioned on other predictors.
+In the previous two examples, we look at model-level explanations; now we 
+move on to one where we investigate one particular prediction instance. As 
+before, we consider how we would approach the question for linear models. For
+a GLM with a log link common in ratemaking applications, for example, we start
+with the base rate, then the exponentiated coefficients would have 
+multiplicative effects on the final rate. Similar to the previous examples,
+for ML models in general we do not have directly interpretable weights.
+Instead, one way to arrive at variable contributions is calculating the change
+in the expected model prediction for each predictor, conditioned on other
+predictors.
 
-Formally, for a fitted model $\widehat{f}$ and a specific instance $x_*$, we would like to decompose the model prediction $\widehat{f}(x_*)$ into
+Formally, for a fitted model $\widehat{f}$ and a specific instance $x_*$, we
+would like to decompose the model prediction $\widehat{f}(x_*)$ into
 
@@
 
-where $v_0$ denotes the average model response, and $v(j, x_*)$ denotes the contribution of the $j$th variable in instance $x_*$, defined as
+where $v_0$ denotes the average model response, and $v(j, x_*)$ denotes the 
+contribution of the $j$th variable in instance $x_*$, defined as
 
@@
 
-Hence, the contribution of the $j$th variable to the prediction is the incremental change in the expected model prediction when we set $X^j = x_*^ j$ assuming the other variables take their values in $x_*$.
-Note here that this definition implies that the order in which we consider the variables affects the results.
-Empirically, the expectations in \@ref(eq:var-contrib) are calculated by sampling the test dataset.
+Hence, the contribution of the $j$th variable to the prediction is the 
+incremental change in the expected model prediction when we set $X^j = x_*^ j$
+assuming the other variables take their values in $x_*$. Note here that this
+definition implies that the order in which we consider the variables affects 
+the results. Empirically, the expectations in \@ref(eq:var-contrib) are 
+calculated by sampling the test dataset.
 
-In Figure \@ref(fig:breakdown), we exhibit a waterfall plot of variable contributions.
-The "intercept" value denotes the average model prediction and represents the $v_0$ term in Equation \@ref(eq:contrib-decompose).
-The predicted loss cost for this particular policyholder is slightly less than average; the characteristics that makes this policyholder more risky are the fact that he is a male between the ages of 18 and 25; counteracting the risky driver characteristics are the vehicle properties: it is a GM vehicle built domestically and is seven years old.
+In Figure \@ref(fig:breakdown), we exhibit a waterfall plot of variable
+contributions. The "intercept" value denotes the average model prediction and 
+represents the $v_0$ term in Equation \@ref(eq:contrib-decompose). The predicted
+loss cost for this particular policyholder is slightly less than average; the 
+characteristics that makes this policyholder more risky are the fact that he is
+a male between the ages of 18 and 25; counteracting the risky driver
+characteristics are the vehicle properties: it is a GM vehicle built
+domestically and is seven years old.
 
-Instance-level explanations are useful for investigating specific problematic predictions generated by the model.
-Regulators and model reviewers may be interested in variable contributions for the safest and riskiest policyholders to see if they conform to intuition.
-A policyholder with a particularly high premium may wish to find out what of their characteristics contribute to it, and may follow up with a question about how he can lower it, which would require another type of explanation.
+Instance-level explanations are useful for investigating specific problematic
+predictions generated by the model. Regulators and model reviewers may be 
+interested in variable contributions for the safest and riskiest policyholders
+to see if they conform to intuition. A policyholder with a particularly high
+premium may wish to find out what of their characteristics contribute to it,
+and may follow up with a question about how he can lower it, which would 
+require another type of explanation.
 
-As noted earlier, the ordering of variables has an impact on the contributions calculated, especially for models that are non-additive, which could cause inconsistent explanations.
-There are several approaches to ameliorate this phenomenon, including selecting variables with the largest contributions first, including interactions terms, and averaging over possible orderings.
-The last of these ideas is implemented by @NIPS2017_7062 using Shapley values from cooperative game theory, and is referred to as Shapley additive explanations (SHAP).
-These approaches are discussed further in @biecekPMVEE and its references.
+As noted earlier, the ordering of variables has an impact on the contributions
+calculated, especially for models that are non-additive, which could cause 
+inconsistent explanations. There are several approaches to ameliorate this 
+phenomenon, including selecting variables with the largest contributions first,
+including interactions terms, and averaging over possible orderings. The last
+of these ideas is implemented by @NIPS2017_7062 using Shapley values from 
+cooperative game theory, and is referred to as Shapley additive explanations 
+(SHAP). These approaches are discussed further in @biecekPMVEE and its references.
 
@@
 
-In this paper, we demonstrate just a few model-agnostic ML interpretation techniques.
-These represent a small subset of existing techniques, each of which have additional variations.
-In the remainder of this section, we point out a few common techniques not covered in our case study.
+In this paper, we demonstrate just a few model-agnostic ML interpretation 
+techniques. These represent a small subset of existing techniques, each of 
+which have additional variations. In the remainder of this section, we point
+out a few common techniques not covered in our case study.
 
-Individual conditional expectation (ICE) plots disaggregate PDPs into their instance-level components for a more granular view into predictor sensitivities [@goldstein2015peeking].
-To accommodate correlated variables in PDP, accumulated local effect (ALE) plots computes expected changes in model response over the conditional, rather than marginal, distribution of the other variables [@apley2016visualizing].
+Individual conditional expectation (ICE) plots disaggregate PDPs into their 
+instance-level components for a more granular view into predictor sensitivities 
+[@goldstein2015peeking]. To accommodate correlated variables in PDP, accumulated 
+local effect (ALE) plots computes expected changes in model response over the 
+conditional, rather than marginal, distribution of the other variables
+[@apley2016visualizing].
 
-Local interpretable model-agnostic explanations (LIME) [@ribeiro2016should] builds simple surrogate models using model predictions, with higher training weights given to the point of interest, in effect replacing the complex ML model with an easily interpretable linear regressions or decision trees in neighborhoods of specific points for the purpose of explanation.
-Taking the concept further, one can also train a global surrogate model across the entire domain of interest.
+Local interpretable model-agnostic explanations (LIME) [@ribeiro2016should] 
+builds simple surrogate models using model predictions, with higher training 
+weights given to the point of interest, in effect replacing the complex ML model 
+with an easily interpretable linear regressions or decision trees in 
+neighborhoods of specific points for the purpose of explanation. Taking the 
+concept further, one can also train a global surrogate model across the entire 
+domain of interest.
 
@@
 variables <- tibble::tribble(
-  ~ Variable, ~ Type, ~ Transformation,
+  ~Variable, ~Type, ~Transformation,
   "Age range", "Categorical", "One-hot encode",
@@
   variables,
-  caption = 'Input variables and their transformations.',
+  caption = "Input variables and their transformations.",
   booktabs = TRUE,
-  table.env='table',
+  table.env = "table",
   escape = FALSE,
@@
 
-In this appendix, we describe the ML model and the data used to train it.
-Note that, for our paper, the ultimate goal of the modeling procedure is to develop something that can produce predictions.
-As a result, we do not follow standard practices for supervised learning model development.
-However, for the sake of completeness and reproducibility, we include an overview of the process here.
-Implementation is done using the R [@rlang] interface to TensorFlow [@tensorflow2015-whitepaper].
-The model explanation visualizations utilize the implemention by @biecekPMVEE, and the code to reproduce them are available on GitHub[^1].
+In this appendix, we describe the ML model and the data used to train it. Note 
+that, for our paper, the ultimate goal of the modeling procedure is to develop 
+something that can produce predictions. As a result, we do not follow standard
+practices for supervised learning model development. However, for the sake of 
+completeness and reproducibility, we include an overview of the process here. 
+Implementation is done using the R [@rlang] interface to TensorFlow 
+[@tensorflow2015-whitepaper]. The model explanation visualizations utilize the 
+implemention by @biecekPMVEE, and the code to reproduce them are available on
+GitHub[^1].
 
@@
 
-We use data from the AUTOSEG ("Automobile Statistics System") of Brazil's Superintendence of Private Insurance (SUSEP).
-The organization maintains policy-characteristics-level data, including claim counts and amounts, for all insured automobiles in Brazil.
-The data contains variables from policyholder characteristics to losses by peril.
-We use the records from the first half of 2012, which contains 1,707,651 records.
-One-fifth of the data is reserved for testing; the remainder is further split into 3/4 of analysis and 1/4 into assessment for determining early stopping.
+We use data from the AUTOSEG ("Automobile Statistics System") of Brazil's 
+Superintendence of Private Insurance (SUSEP). The organization maintains 
+policy-characteristics-level data, including claim counts and amounts, for all 
+insured automobiles in Brazil. The data contains variables from policyholder 
+characteristics to losses by peril. We use the records from the first half of 
+2012, which contains 1,707,651 records. One-fifth of the data is reserved for 
+testing; the remainder is further split into 3/4 of analysis and 1/4 into 
+assessment for determining early stopping.
 
@@
 
-Table \@ref(tab:variables) shows the input variables to our model and their associated transformations.
-For "make" and "region", we map each level to a point in $\mathbb{R}^2$ through embedding layers.
-The model is a feedforward neural network with two hidden layers with 64 units each.
-The activations for the hidden layers are ReLU while for the output layer it is softplus.
-Exposures for each record are used as sample weights during training.
-We fit the model via ADAM with an initial learning rate of 0.1, a mini-batch size of 10,000, and trigger early stopping when the mean squared error on the asssessment set does not improve for five epochs.
+Table \@ref(tab:variables) shows the input variables to our model and their 
+associated transformations. For "make" and "region", we map each level to a 
+point in $\mathbb{R}^2$ through embedding layers. The model is a feedforward
+neural network with two hidden layers with 64 units each. The activations for
+the hidden layers are ReLU while for the output layer it is softplus. Exposures
+for each record are used as sample weights during training. We fit the model
+via ADAM with an initial learning rate of 0.1, a mini-batch size of 10,000,
+and trigger early stopping when the mean squared error on the asssessment set 
+does not improve for five epochs.

### 2019-10-30 manuscript/manuscript.Rmd
Remove pdf and tex

@@
 GLMs did not become mainstream in actuarial science for some time. For example, 
-the syllabus of basic education does not seem to include any mention of GLMs 
-prior to @brown_1988 in the 1990 syllabus for basic education. From there, GLMs 
+the syllabus of basic education of the Casualty Actuarial Society does not seem to include any mention of GLMs 
+prior to @brown_1988 in the 1990 syllabus. From there, GLMs 
 seem to have received only passing mention until 2006 with the introduction of 
@@
 2. Prepare data for analysis
-3. Perform analysis on the data, employing desired method or methods to estimate
-needed rate relativites
-4. Select final rate relativities based on rate indications
+3. Perform analysis on the data, employing desired method(s) to estimate
+needed rates
+4. Select final rates based on rate indications
 5. Present rates to the reglator, including explanation of the steps followed to 
@@
 magnitude of rate changes or filings that make use of new or sophisticated 
-predictive models may be subject to intense regulatory scrutiny. In these cases,
+predictive models may be subject to particular regulatory scrutiny. In these cases,
 it is necessary to be able to explain the results of the modeling process in a 
@@
 practicing in the same field should be able to make an objective appraisal of 
-the results from the methods found in the syllabus. [This can be supported by 
-reference to the edits to ASOP 38 over time - originally it was designed to be 
-about all models, but they revised it to be about only models that incorporate 
-specialized knowledge outside of the actuary's area of expertise... do we need 
-this citation, though?] This is notable because, beginning with the introduction
+the results from the methods found in the syllabus.
+This is notable because, beginning with the introduction
 of the MAS-I and MAS-II examination in July of 2018, several machine learning 
@@
 Nevertheless, machine learning poses something of a special challenge for ASOP 
-41 for several reasons. Machine learning models can be very ad hoc compared to 
-traditional statistical models. Because many machine learning models are 
-deterministic, they may not admit of standard metrics for model comparison 
+41 for several reasons:
+
+1. Machine learning models can be very ad hoc compared to 
+traditional statistical models.
+2. Because many machine learning models do not assume an underlying probability
+distribution or stochastic process, they may not admit of standard metrics for model comparison 
 (e.g., it's not straightforward to calculate an AIC over a neural network). 
-In addition, machine learning methods are often combined into ensembles that 
+3. Machine learning methods are often combined into ensembles that 
 may not be easily separated and that may, as a collection, cease to resemble 
-a single standard version of a model. Complicating matters still further, 
-machine learning models can be "black boxes" insofar as the final form of 
+a single standard version of a model.
+4. Machine learning models can be "black boxes" insofar as the final form of 
 response curve cannot be easily predicted and may depend heavily on the 
@@
 know the method employed to make an objective appraisal of the reasonableness 
-of the model. More information is necessary.
+of the result. More information is necessary.
 
-Of course, these comments only apply to the actuarial profession. Outside of 
+Of course, these comments only apply within the actuarial profession. Outside of 
 the actuarial profession, communication of results may be more challenging. A 
@@
 
-* Determine the extent the model causes premium disruption for individual 
-policyholders, and how the insurer will
+* Determine the extent to which the model causes premium disruption for individual 
+policyholders and how the insurer will
 explain the disruption to individual consumers that inquire about it.
@@
 As we proceed through the points below, we aim to arrive at a more scoped and 
-relevant definition of what it means to attain interpretability in a pricing
-model. In the remainder of this section, we clarify a couple concepts regarding 
-interpretable *classes* of models and and the computational transparency of ML
-models, outline findings on the social aspects of ML explanations most relevant
-to our use case, then discuss a potential framework for implementing ML 
+relevant definition of what it means for a pricing
+model to be interpretable. In the remainder of this section, we clarify a couple concepts regarding 
+interpretable *classes* of models and the computational transparency of ML
+models, outline frameworks for understanding the communication goals of interpretability,
+then discuss a potential framework for implementing ML 
 interpretability in practice.
@@
 consider a claim severity model with driver age, sex, and vehicle age as 
-features; assuming a log link function and letting $Y$ denotes the response, we
+predictors; assuming a log link function and letting $Y$ denote the response, we
 have
@@
 severity if we were to increase age by a certain amount, *all else being equal*,
-because the relationship between the feature and the response is simply 
+because the relationship between the predictor and the response is simply 
 multiplication by the coefficient $\beta_1$ and applying the inverse link
@@
 are interpretable models, we are implicitly assuming that we are considering 
-only a handful  of features. In fact, the ease at which we can reason about a
-model declines as the number of features, transformations of them, and
+only a handful  of predictors. In fact, the ease with which we can reason about a
+model declines as the number of predictors, transformations of them, and
 interactions increase, as in the following (somewhat pathological) example:
@@
 
-Proposed by @hiltonConversationalProcesses1990 and interpreted by
-@millerExplanationArtificial2017 in the context of ML, model explanations
-are conversations or social interactions. One consequence of this identification
-is that explanations need to be *relevant* to the *audience*. In developing,
+@hiltonConversationalProcesses1990 proposed a framework, later interpreted by
+@millerExplanationArtificial2017 in the context of ML, for understanding model explanations
+as *conversations* or *social interactions.* One consequence of this identification
+is that explanations need to be *relevant* to the *audience*. This framework is
+consistent with ASOP 41, which formulates a similar requirement in terms of an
+*intended user* of the actuarial communication. In developing,
 filing, and operationalizing a pricing model, one needs to accommodate a variety
@@
 models, and will focus on a specific list of questions motivated by statute
-and public policy. Finally, we have the public, who has an interest (perhaps
-more so than the other parties) because their lives are directly affected by
-the premiums charged.
+and public policy. Finally, we have potential policyholders, who have an interest (perhaps
+more so than the other parties) as they are responsible for paying the premiums.
 
-It is interesting to note that, the modelers, who are most familiar with the
+It is interesting to note that the modelers, who are most familiar with the
 models, tend to be same people designing and communicating the explanations.
@@
 than the intended audience. For example, they may be interested in technical
-questions such as extrapolation behavior and shape the explanations accordingly,
-which may be lost on the prospective policyholder.
+questions, such as extrapolation behavior, and shape the explanations accordingly,
+which may be irrelevant to a prospective policyholder.
 
@@
 
+It should be noted that no ideal set of questions exists that would encompass
+all potential models. Rather, the actuary must consider what aspects of the
+model would raise questions from the perspective of the model's intended users.
+We propose that relevant stakeholders, by providing example questions and
+answers, would inherently provide guidance by which actuaries can reasonably
+anticipate the kinds of specific questions most important to those stakeholders
+and address them proactively.
+
 These questions should relate to existing guidelines, such as those described in
@@
 regulation, and in fact should not be specific only to ML models. By
-standardizing on a set of questions, we reduce the burden of both companies and
+conceptualizing a set of questions, we reduce the burden of both companies and
 regulators; this is especially important for the latter, who are already
@@
 
-1. What are the most important predictors in the model?
+1. What are the most important predictors in the model? Put another way, to
+what extent do the predictors improve the accuracy of the model?
 2. How does the predicted loss cost change, on average, as we change an input?
@@
 variable importance can be obtained from the fitted model structure. In the
-case of GLMs, one can observe the magnitudes of the $t$-statistics, whereas
-for random forests, one can use out-of-bag errors [@breimanRandomForests2001].
+case of GLMs, one might observe the magnitudes of the $t$-statistics, step-wise
+reduction in information criteria, or out-of-sample performance, whereas
+for random forests, one might use out-of-bag errors [@breimanRandomForests2001].
 For more complex models, such as the neural network in our case study, we need
@@
 in the case of regression. Now, if $\widetilde{X}^{j}$ denotes the predictor
-matrix where the $j$th variable has been permutated, then we can compute the
+matrix where the $j$th variable has been permuted, then we can compute the
 loss with the permuted dataset as
@@
 for the audience to quickly glance at the most relevant variables, and ask
-further question as necessary. For example, if a variable that can potentially
-be a proxy for a discriminatory variable appears as very important, more
-investigation would be warranted.
+further question as necessary.
 
@@
 
+Note that this proposed plot can be calculated over the dataset or over a
+hypothetical dataset in which the value of $x_i^{j}$ is set equal to $z$ for
+each record.
+
 In Figure \@ref(fig:pdp), we exhibit the PDP for the "vehicle age" variable.
@@
 data is quite thin for vehicle age greater than 18, so the apparent upward
-trend to the right is driven by just a few data points. With this in mind,
-we can conclude that as cars get older, we expect the loss cost predictions
-generated by our model to decrease.
+trend to the right is driven by just a few data points.
+
+This information allows the modeler and stakeholders to consider whether it
+is reasonable for the anticipated loss cost to follow this shape.
 
@@
 
-(TODO)
+Actuarial standards of practice, most notably ASOP 41, places responsibility on
+the actuary to clearly communicate actuarial work products, including insurance
+pricing models. These responsibilities create special challenges for
+communicating machine learning models, which are often seen "black boxes" due
+in part to their complexity, nonlinearity, flexible construction, and ad hoc
+nature.
+
+In this paper, we discussed particular questions of model validation that are
+of key importance in communicating a model that may present particular
+difficulty for machine learning models compared to GLMs or traditional pricing
+models. Specifically,
+
+* How does the model impact individual insurance consumers?
+* How are the predictor variables related to expected losses?
+* Are the variables unfairly discriminatory?
+
+We contextualized these questions in terms of different frameworks for defining
+interpretability. We conceptualize interpretability in terms of the ability of
+a model (or modeler) to answer a set of idealized questions that would be
+refined. We then offered potential (families of) model-agnostic techniques for
+providing answers to these questions.
+
+Much work remains to be done in terms of defining the role of machine learning
+algorithms in actuarial practice. Lack of interpretability has been a key
+barrier preventing wider adoption and exploration of these techniques. The
+methods proposed in this paper could therefore represent important strides in
+nlocking the potential of machine learning within the insurance industry.
 
@@
 something that can produce predictions. As a result, we do not follow standard
-practices for supervised learning model development. However, for the sake of 
+practices for tuning and validation. However, for the sake of 
 completeness and reproducibility, we include an overview of the process here. 
@@
 associated transformations. For "make" and "region", we map each level to a 
-point in $\mathbb{R}^2$ through embedding layers. The model is a feedforward
+point in $\mathbb{R}^2$ through embedding layers. The model predicts expected
+loss cost for all perils combined. The architecture is a feedforward
 neural network with two hidden layers with 64 units each. The activations for

### 2019-10-30 manuscript/manuscript.tex
Copy edits and conclusion

% interactcadsample.tex
% v1.03 - April 2017

\documentclass[]{interact}

\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
\usepackage{subfigure}% Support for small, `sub' figures and tables
%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required

\usepackage{natbib}% Citation support using natbib.sty
\bibpunct[, ]{(}{)}{;}{a}{}{,}% Citation support using natbib.sty
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty

\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}

% see https://stackoverflow.com/a/47122900

\usepackage{hyperref} \usepackage[utf8]{inputenc} \def\tightlist{}

\begin{document}

\articletype{WIP DRAFT}

\title{Towards Explainability of Machine Learning Models in Insurance Pricing}

\author{\name{Author One$^{a}$, Author Two$^{b}$}
\affil{$^{a}$Foo; $^{b}$Bar}
}

\thanks{CONTACT Author One. Email: \href{mailto:foo@bar.foo}{\nolinkurl{foo@bar.foo}}, Author Two. Email: \href{mailto:bar@foo.bar}{\nolinkurl{bar@foo.bar}}}

\maketitle

\begin{abstract}
Abstract tbd
\end{abstract}

\begin{keywords}
actuarial science; general insurance
\end{keywords}

\section{Introduction}\label{introduction}

Risk classification for property \& casualty (P\&C) insurance rating has
traditionally been done with one-way, or univariate, analysis
techniques. In recent years, many insurers have moved towards using
generalized linear models (GLM), a multivariate predictive modeling
technique, which addresses many shortcomings of univariate approaches,
and is currently considered the gold standard in insurance risk
classification. At the same time, machine learning (ML) techniques such
as deep neural networks have gained popularity in many industries due to
their superior predictive performance over linear models
\citep{lecunDeepLearning2015}. In fact, there is a fast growing body of
literature on applying ML to P\&C reserving
\citep{kuoDeepTriangleDeep2018, wuthrichMachineLearning2018, gabrielliNeuralNetwork2019a, gabrielliNeuralNetwork2019}.
However, these ML techniques, often considered to be completely ``black
box'', have been less successful in gaining adoption in pricing, which
is a regulated discipline and requires a certain amount of transparency
in models.

If insurers can gain more insight into how ML models behave in risk
classification contexts, it would increase their ability to reassure
regulators and the public that accepted ratemaking principles are met.
Being able to charge more accurate premiums would, in turn, make the
risk transfer system more efficient and contribute to the betterment of
society. In this paper, we aim to take a step towards liberating
actuaries from the confines of linear models in pricing projects, by
proposing a framework for explaining ML models for ratemaking that
regulators, practitioners, and researchers in actuarial science can
build upon.

The rest of this paper is organized as follows: Section \ref{ratemaking}
provides an overview of P\&C ratemaking, Section \ref{interpretability}
discusses model interpretability in the context of ratemaking and
proposes specific tasks for model explanation, Section \ref{application}
describes current model interpretation techniques and applies them to
the tasks defined in the previous section, and Section \ref{conclusion}
concludes.

\section{Property and Casualty Ratemaking}\label{ratemaking}

\subsection{History of Ratemaking}\label{history-of-ratemaking}

Early classification ratemaking procedures were typically univariate in
nature. For example, \citep{lange_1966} notes that (at that time) most
major lines of insurance used univariate methods based around the same
principle: distributing an overall indication to territorial
relativities or classification relativities based on the extent to which
they deviated from the average experience.

\citep{bailey_simon_1960} introduced minimum bias methods, which were
expanded throughout the 60s, 70s, and 80s. As computing power developed,
minimum bias began to give away to generalized linear models, with
papers such as \citep{brown_1988} and \citep{mildenhall_1999} bridging
the gap between the methods.

Arguably, generalized linear models predate minimum bias procedures by a
significant margin. The term ``Generalized Linear Model'' was coined by
\citep{nelder_wedderburn_1972}, but generalizations of least squares
linear regression date back at least to the 1930s. Like minimum bias
methods, GLMs did not become mainstream in actuarial science for some
time. For example, the syllabus of basic education does not seem to
include any mention of GLMs prior to \citep{brown_1988} in the 1990
syllabus for basic education. From there, GLMs seem to have received
only passing mention until 2006 with the introduction of
\citep{anderson_2005} to the syllabus.

\subsection{Machine Learning in
Ratemaking}\label{machine-learning-in-ratemaking}

Paralelling the development of generalized linear models was the
development of machine learning algorithms throughout the middle part of
the 20th century. Detailed histories of machine learning may be found in
sources such as \citep{Nilsson_2009} and \citep{wang_raj_2017}.
Consistent with GLMs, machine learning was relatively unpopular in
actuarial science until the last ten years as computing power has become
cheaper and more easily available and as machine learning software
packages have obviated the need for developing analyses from scratch
each time an analysis is performed. Due to the breadth of machine
learning as a field, it is difficult to identify the first time it
entered the CAS syllabus; however, cluster analysis (in the form of
k-means) seems to have been first included in 2011 with
\citep{robertson_2009}. More recently, the MAS-I and MAS-II exams
introduced in 2018 have included machine learning explicitly.

Within the area of ratemaking, machine learning is still in its infancy.
A significant portion of machine learning applications to ratemaking has
been in the context of automobile telematics, such as \citep{gao_2018},
\citep{gao_2018_2}, \citep{gao_2019}, \citep{roel_2018}, or
\citep{wuthrich_2017}. Presumably this focus has been a result of the
high-dimensionality and complexity of telematics data, making it a field
in which the unique abilities of machine learning techniques give a
clear advantage over traditional approaches.

Outside of telematics, \citep{yang_2018} uses a gradient tree-boosting
approach to capture non-linearities that would be a challenge for GLMs.
\citep{henckaerts_2018} makes use of ``generalized additive models'' to
improve predictions of GLMs. Many researchers, in an apparent effort to
demonstrate the range of possibilities and advantages of machine
learning, have approached the topic by comparing many different machine
learning algorithms within a single study, such as in
\citep{dugas_2003}, \citep{noll_2018}, \citep{spedicato_2018}. These
studies make use of such varied techniques as regression trees, boosting
machines, support vector machines, and neural networks.

\subsection{Ratemaking Process}\label{ratemaking-process}

Regardless of the method employed for determining this risk of various
classifications, the actual process of setting rate relativities
typically involves some variation of the following steps:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Obtain relevant policy-level data
\item
  Prepare data for analysis
\item
  Perform analysis on the data, employing desired method or methods to
  estimate needed rate relativites
\item
  Select final rate relativities based on rate indications
\item
  Present rates to the reglator, including explanation of the steps
  followed to derive the rates
\item
  Answer questions from regulators regarding the method employed
\end{enumerate}

The focus of this paper is on steps 5 and 6. In particular, rate
regulators are concerned with whether rates are inadequate, excessive,
or unfairly discriminatory. In many states, rate filings that exceed
certain thresholds for magnitude of rate changes or filings that make
use of new or sophisticated predictive models may be subject to intense
regulatory scrutiny. In these cases, it is necessary to be able to
explain the results of the modeling process in a way that is
understandable without sacrificing statistical rigor.

It should be noted that communicating results is not simply a method of
passing regulatory muster. Generating interpretable modeling output is
an important - even essential - facet of model checking. Therefore, the
techniques discussed in this paper may be viewed from the lens of
providing useful information to regulators, but they should also be
considered as part of a thorough vetting of any rating model.

\section{Interpretability in the Ratemaking
Context}\label{interpretability}

(literature review, settle on a definition to work with, e.g.
\citet{doshi-velezRigorousScience2017}, identify maybe 3 questions about
a model to answer, tie to principles on p/c ratemaking)

Within the actuarial profession, Actuarial Standard of Practice 41
(``Actuarial Communications'') notes that ``\ldots{}another actuary
qualified in the same practice area {[}should be able to{]} make an
objective appraisal of the reasonableness of the actuary's work as
presented in the actuarial report.'' \citep{asop_41} Underlying this
requirement is an assumption that the hypothetical other actuary
qualified in the same practice area is adequately familiar with the
relevant techniques employed. Although the syllabus of basic education
is constantly changing, there has at times been an assumption that all
techniques and assumptions that have ever been a part of the syllabus of
basic education needn't be explained from first principles in general
actuarial communications, and that an actuary practicing in the same
field should be able to make an objective appraisal of the results from
the methods found in the syllabus. {[}This can be supported by reference
to the edits to ASOP 38 over time - originally it was designed to be
about all models, but they revised it to be about only models that
incorporate specialized knowledge outside of the actuary's area of
expertise\ldots{} do we need this citation, though?{]} This is notable
because, beginning with the introduction of the MAS-I and MAS-II
examination in July of 2018, several machine learning models were
formally included in the syllabus of basic education. These exams cover
a wide range of topics, such as splines, clustering algorithms, decision
trees, boosting, and principle components analysis.
\citep{cas_syllabus_2018}

Nevertheless, machine learning poses something of a special challenge
for ASOP 41 for several reasons. Machine learning models can be very ad
hoc compared to traditional statistical models. Because many machine
learning models are deterministic, they may not admit of standard
metrics for model comparison (e.g., it's not straightforward to
calculate an AIC over a neural network). In addition, machine learning
methods are often combined into ensembles that may not be easily
separated and that may, as a collection, cease to resemble a single
standard version of a model. Complicating matters still further, machine
learning models can be ``black boxes'' insofar as the final form of
response curve cannot be easily predicted and may depend heavily on the
available data (which may not, in turn, be available to the reviewer).

This last item raises a final interesting issue. Generalized linear
models and their ilk are often fitted using one of a handful of standard
and well-understood approaches (e.g., maximum likelihood estimation).
However, this is not possible in general with machine learning models,
as machine learning algorithms often use loss surfaces that are very
complex such that it may not be feasible to calculate the global minimum
of the surface. Certainly, closed form representations of the loss
surfaces are not generally available. For this reason, the training
phase of a machine learning model is, in many ways, just as important to
one's understanding as the model form and the data on which the model is
fitted. Because the final model result is inseparable from these three
components (training method, model form, and data), it is not generally
adequate to just know the method employed to make an objective appraisal
of the reasonableness of the model. More information is necessary.

Of course, these comments only apply to the actuarial profession.
Outside of the actuarial profession, communication of results may be
more challenging. A 2017 survey conducted by the Casualty Actuarial and
Statistical Task Force of the National Association of Insurance
Commissioners found that the plurality of responding regulators
identified ``Filing complexity and/or a lack of resources or expertise''
as a key challenge that impedes their ability to review GLMs or other
predictive models. \citep{naic_summer_2017} Given that machine learning
algorithms are generally regarded as more complex than GLMs, this
implies that the challenge of communicating machine learning model
results is significant.

In response to the same survey, 33 state regulators noted that it would
be helpful or very helpful for the NAIC to develop information and tools
to assist in reviewing rate filings based on GLMs, and 34 noted that it
would be helpful to develop similar items to assist in reviewing ``Other
Advanced Modeling Techniques.'' One outgrowth of this need was the
development of a white paper on best practices for regulatory review of
predictive models. The white paper focuses on review of GLMs,
particularly with respect to private passenger automobile and
homeowners' insurance. Some of the guidance offered in this regard is
therefore not strictly applicable to the review of machine learning
models. For example, as previously noted, p-values are not a concept
that translates well to deterministic machine learning algorithms.
However, among the guidance applicable to machine learning algorithms
are the following (paraphrasing):

\begin{itemize}
\tightlist
\item
  understand the relationship between the inputs and the expected loss
  or expense differences in risk,
\item
  determine that these input characteristics are not unfairly
  discriminatory, and
\item
  determine the extent of premium disruption among policyholders and be
  able to explain the source of premium disruptions.
  \citep{naic_white_paper}
\end{itemize}

Note that this list is not exhaustive of the guidance offered by any
means, but that these three items are those that may present greater
challenge for machine learning algorithms compared to GLMs.

Depending on the model, machine learning algorithms can be non-linear in
the inputs. In areas where data are sparse, the model could generate
unnatural response curves that could go undetected by a very high-level
view of the model results. This leads to a situation where lift charts
or tests on holdout data may indicate that the model is performing
adequately, but it may be difficult to explain the reasons for a
particular premium indicaton or a change in premium when policy
characteristics change.

-For this, we recommend \_\_\_\_ (grid search?) -Because the model form
may be difficult to represent in the form of a single equation (and
harder still to evaluate), one method of evaluating the reasonableness
of model results is local approximations of the impact of different
rating variables (LIME?)

\section{Applying Model Interpretation Techniques}\label{application}

(some definitions, e.g.~global/model vs.~local/instance, categorize
questions accordingly)

\subsection{(answer each question)}\label{answer-each-question}

(for each question, propose a technique, mention alternative techniques,
pros/cons, implement, interpret results)

\section{Conclusion}\label{conclusion}

(conclude)

\bibliographystyle{tfcad}
\bibliography{explain-pricing.bib}

\input{"appendix.tex"}

\end{document}

### 2019-10-31 R/explain.R
mess with pdp

@@ -2,6 +2,7 @@ library(tidyverse)
 library(DALEX)
 library(ggplot2)
 library(scales)
+library(patchwork)
 
 pins::board_register_github(name = "cork", repo = "kasaai/cork")
 testing_data <- pins::pin_get("toy-model-testing-data", board = "cork")
@@ -25,20 +26,31 @@ explainer_nn <- DALEX::explain(
   predict_function = custom_predict,
   label = "neural_net"
 )
-
+# debugonce(ingredients::partial_dependency)
 pdp_vehicle_age <- ingredients::partial_dependency(
   explainer_nn, 
   "vehicle_age",
-  N = 1500
+  N = 10000,
+  variable_splits = list(vehicle_age = seq(0, 35, by =0.1))
 )
+
 
 pdp_plot <- as.data.frame(pdp_vehicle_age) %>% 
   ggplot(aes(x = `_x_`, y = `_yhat_`)) + 
   geom_line() +
   xlab("Vehicle Age") +
   ylab("Average Predicted Loss Cost") +
-  theme_bw() +
-  geom_rug(sides = "b")
+  theme_bw()
+
+pdp_plot
+
+vehicle_age_histogram <- testing_data %>% 
+  ggplot(aes(x = vehicle_age)) + 
+  geom_histogram() +
+  theme_bw()
+
+pdp_plot / vehicle_age_histogram +
+  plot_layout(heights = c(2, 1))
 
 ggsave("manuscript/figures/pdp-plot.png", plot = pdp_plot)
 

### 2019-11-01 manuscript/manuscript.Rmd
conclusion to present tense

--- 2019-10-30 manuscript/manuscript.Rmd
@@
 
-In this paper, we discussed particular questions of model validation that are
+In this paper, we discuss particular questions of model validation that are
 of key importance in communicating a model that may present particular
@@
 
-We contextualized these questions in terms of different frameworks for defining
+We contextualize these questions in terms of different frameworks for defining
 interpretability. We conceptualize interpretability in terms of the ability of
 a model (or modeler) to answer a set of idealized questions that would be
-refined. We then offered potential (families of) model-agnostic techniques for
+refined. We then offer potential (families of) model-agnostic techniques for
 providing answers to these questions.
@@
 methods proposed in this paper could therefore represent important strides in
-nlocking the potential of machine learning within the insurance industry.
+unlocking the potential of machine learning within the insurance industry.
 

### 2019-11-01 manuscript/manuscript.Rmd
remove some pdp description
remove mention of stepwise

@@
 variable importance can be obtained from the fitted model structure. In the
-case of GLMs, one might observe the magnitudes of the $t$-statistics, step-wise
-reduction in information criteria, or out-of-sample performance, whereas
+case of GLMs, one might observe the magnitudes of the estimated coefficients or 
+$t$-statistics, whereas
 for random forests, one might use out-of-bag errors [@breimanRandomForests2001].
@@
 where $N$ is the number of records in the dataset.
-
-Note that this proposed plot can be calculated over the dataset or over a
-hypothetical dataset in which the value of $x_i^{j}$ is set equal to $z$ for
-each record.
 

### 2019-11-01 manuscript/manuscript.Rmd
qualify mas exams as being cas exams

@@
 GLMs did not become mainstream in actuarial science for some time. For example, 
-the syllabus of basic education of the Casualty Actuarial Society does not seem to include any mention of GLMs 
+the syllabus of basic education of the Casualty Actuarial Society (CAS) does not 
+seem to include any mention of GLMs 
 prior to @brown_1988 in the 1990 syllabus. From there, GLMs 
@@
 form of k-means) seems to have been first included in 2011 with @robertson_2009.
-More recently, the MAS-I and MAS-II exams introduced in 2018 have included 
+More recently, the CAS MAS-I and MAS-II exams introduced in 2018 have included 
 machine learning explicitly.
@@
 This is notable because, beginning with the introduction
-of the MAS-I and MAS-II examination in July of 2018, several machine learning 
+of the CAS MAS-I and MAS-II examination in July of 2018, several machine learning 
 models were formally included in the syllabus of basic education. These exams 

### 2019-11-01 manuscript/manuscript.Rmd
naic acronym

@@
 2017 survey conducted by the Casualty Actuarial and Statistical Task Force of 
-the National Association of Insurance Commissioners found that the plurality of
+the National Association of Insurance Commissioners (NAIC) found that the plurality of
 responding regulators identified "Filing complexity and/or a lack of resources 

### 2019-11-01 R/explain.R
update plots

@@ -26,30 +26,31 @@ explainer_nn <- DALEX::explain(
   predict_function = custom_predict,
   label = "neural_net"
 )
-# debugonce(ingredients::partial_dependency)
+
 pdp_vehicle_age <- ingredients::partial_dependency(
   explainer_nn, 
   "vehicle_age",
   N = 10000,
-  variable_splits = list(vehicle_age = seq(0, 35, by =0.1))
+  variable_splits = list(vehicle_age = seq(0, 35, by = 0.1))
 )
-
 
 pdp_plot <- as.data.frame(pdp_vehicle_age) %>% 
   ggplot(aes(x = `_x_`, y = `_yhat_`)) + 
   geom_line() +
-  xlab("Vehicle Age") +
   ylab("Average Predicted Loss Cost") +
-  theme_bw()
-
-pdp_plot
+  theme_bw()+
+  theme(axis.ticks.x = element_blank(),
+        axis.text.x=element_blank(),
+        axis.title.x=element_blank())
 
 vehicle_age_histogram <- testing_data %>% 
   ggplot(aes(x = vehicle_age)) + 
-  geom_histogram() +
-  theme_bw()
+  geom_histogram(alpha = 0.8) +
+  theme_bw() +
+  ylab("Count") +
+  xlab("Vehicle Age")
 
-pdp_plot / vehicle_age_histogram +
+pdp_plot <- pdp_plot / vehicle_age_histogram +
   plot_layout(heights = c(2, 1))
 
 ggsave("manuscript/figures/pdp-plot.png", plot = pdp_plot)
@@ -76,7 +77,7 @@ fi_plot <- fi %>%
     df %>% 
       filter(!variable %in% c("_full_model_", "_baseline_")) %>% 
       ggplot(aes(x = reorder(variable, dropout_loss), y = dropout_loss)) +
-      geom_bar(stat = "identity") +
+      geom_bar(stat = "identity", alpha = 0.8) +
       geom_hline(yintercept = full_model_loss, col = "red", linetype = "dashed")+
       scale_y_continuous(limits = c(full_model_loss, NA),
                          oob = rescale_none

### 2019-11-01 manuscript/manuscript.Rmd

@@
 We see that the average predicted loss cost decreases with vehicle age until
-the latter is around 18. Note that the rug on the $x$-axis tell us that the
+the latter is around 18. Note that the accompanying histogram shows that the
 data is quite thin for vehicle age greater than 18, so the apparent upward
@@
 
+We thank Navdeep Gill for his presentation on interpretable ML at SDSS and
+helpful follow-up discussions, and Daniel Falbel for his assistance with R and 
+TensorFlow.
 This work is supported by the Casualty Actuarial Society.

### 2020-03-02 manuscript/manuscript.Rmd

@@
 seem to have received only passing mention until 2006 with the introduction of 
-@anderson_2005 to the syllabus.
+@anderson_2005 to the syllabus. Beginning in 2016, the CAS introduced @goldburd_2016
+to the syllabus, which offers a comprehensive guide to GLMs.
 
@@
 
-The focus of this paper is on steps 5 and 6. In particular, rate regulators are 
-concerned with whether rates are inadequate, excessive, or unfairly 
-discriminatory. In many states, rate filings that exceed certain thresholds for
+The focus of this paper is on steps 5 and 6.
+In many states, rate filings that exceed certain thresholds for
 magnitude of rate changes or filings that make use of new or sophisticated 
@@
 regulatory muster. Generating interpretable modeling output is an important - 
-even essential - facet of model checking. Therefore, the techniques discussed 
+even essential - facet of model checking. Actuaries are bound by relevant standards
+to be able to exercise appropriate judgment in selecting risk characteristics as
+part of a risk classification system per Actuarial Standard of Practice 12
+("Risk Classification"). Therefore, the techniques discussed 
 in this paper may be viewed from the lens of providing useful information to 
@@
 any rating model.
+
+Although the focus of this paper is on communication to regulators, it should be
+said that selecting final rates based on indications (step 4 in the list above) may
+pose a unique challenge for black-box models. This, too, provides strong motivation
+for techniques that could add to the modeler's - or any stakeholder's -
+understanding of the model, such as the relative importance of variables or the
+shapes of response curves. Such techniques could be usefully employed in making
+decisions about how best to select rates.
+
+Similarly, although the focus of this paper is on communication in a pricing
+context, the techniques explored in this paper (and many of the concerns discussed)
+may also be relevant to other contexts, such as claim-level reserving / analytics
+or other applications of machine learning to the insurance industry.
 
@@
 
-Of course, these comments only apply within the actuarial profession. Outside of 
+These issues also pose particular challenges with respect to other standards.
+For instance, as discussed previously, ASOP 12 requires
+actuaries to be able to exercise appropriate judgment about risk classification
+systems. The recent ASOP 56 ("Models") speaks to more general concerns in all
+practice areas that might make use of models. ASOP 56 requires the actuary to
+"make reasonable efforts to confirm that the model structure, data, assumptions,
+governance and controls, and model testing and output validation are consistent
+with the intended purpose." [@asop_56] All of these efforts may be hampered if
+it is not possible to peer into the black box of the model.
+
+It should also be noted that these comments only apply within the actuarial profession. Outside of 
 the actuarial profession, communication of results may be more challenging. A 
@@
 
-These items are by no means exhaustive, but they pertain to the concept of model
+The last of these items is an entire topic unto itself. The methods and
+concepts introduced in this paper are useful for exploring the question of whether
+rates are appropriately related to risk of loss as defined by the variables used
+in the model, but there are many other aspects of discrimination-free that are
+outside the scope of this paper. The methods in this paper may help in understanding
+the model, which is a necessary precursor to addressing the question of unfair
+discrimination.
+
+The items in this list are by no means exhaustive, but they pertain to the concept of model
 interpretability for ratemaking that we develop next.
@@
 variable importances and sensitivities, on average, of the predicted response
-with respect to individual predictors. In our case study, questions 1 and 2 are
-associated with this category. On the other hand, question 3 pertains to an
+with respect to individual predictors. Note that these methods may be compared to
+the methods described in [@goldburd_2016], Chapter 7, which focus on global
+interpretation of GLMs.
+
+In our case study, questions 1 and 2 are
+associated with global interpretations. On the other hand, question 3 pertains to an
 individual prediction, which would fall in the local, or instance-level,
@@
 (SHAP). These approaches are discussed further in @biecekPMVEE and its references.
+Shapley values were also used in @mango_1998, which has appeared on the CAS syllabus
+starting in 2004, in the context of determining how to allocate catastrophe
+risk loads between multiple accounts.
 
@@
 * How are the predictor variables related to expected losses?
-* Are the variables unfairly discriminatory?
 

### 2020-03-02 manuscript/manuscript.Rmd
quick clarification on permuting a variable
mention ordering of variables in attribution definition

@@
 loss with the permuted dataset as
-$L^{-j} = \mathcal{L}(\widehat{f}(\widetilde{X}^{j}), y)$. With this, we define
+$L^{-j} = \mathcal{L}(\widehat{f}(\widetilde{X}^{j}), y)$. 
+Here, by *permuting* a variable, we mean that we randomly rearrange the values 
+in the column of data associated with the variable.
+With this, we define
 the variable importance $VI^{j}$ as $L^{-j} - L$.
@@
 
-Formally, for a fitted model $\widehat{f}$ and a specific instance $x_*$, we
+Formally, for a fitted model $\widehat{f}$, a given ordering of the variables
+$X^1,\dots,X^p$, where $p$ is the number of predictor variables,
+and a specific instance $x_*$, we
 would like to decompose the model prediction $\widehat{f}(x_*)$ into

### 2020-03-02 manuscript/manuscript.Rmd
singular/plural consistency for citations

@@
 
-Outside of telematics, @yang_2018 uses a gradient tree-boosting approach to 
+Outside of telematics, @yang_2018 use a gradient tree-boosting approach to 
 capture non-linearities that would be a challenge for GLMs. @henckaerts_2018 
-makes use of generalized additive models (GAM) to improve predictions of GLMs.
+make use of generalized additive models (GAM) to improve predictions of GLMs.
 Many researchers, in an apparent effort to demonstrate the range of possibilities
@@
 models, tend to be same people designing and communicating the explanations.
-This poses a challenge that @millerExplainableAI2017 calls "inmates running
+This poses a challenge that @millerExplainableAI2017 call "inmates running
 the asylum", where the modelers design explanations for *themselves* rather

### 2020-03-02 manuscript/manuscript.Rmd
a word

@@
 biased by unrealistic records in the permuted dataset. Another ramification
-of a group correlated variables is that their inclusion may cause each to
+of a group of correlated variables is that their inclusion may cause each to
 appear less important than if only one is included in the model.

### 2020-03-02 R/explain.R
update code to reflect current DALEX version

@@ -4,6 +4,9 @@ library(ggplot2)
 library(scales)
 library(patchwork)
 
+# Need to use the following branch of {ingredients}
+# remotes::install_github("kevinykuo/ingredients", ref = "weights")
+
 pins::board_register_github(name = "cork", repo = "kasaai/cork")
 testing_data <- pins::pin_get("toy-model-testing-data", board = "cork")
 piggyback::pb_download(file = "model_artifacts/toy-model.tar.gz", repo = "kasaai/cork")
@@ -23,6 +26,7 @@ explainer_nn <- DALEX::explain(
   model = toy_model,
   data = testing_data,
   y = testing_data$loss_per_exposure,
+  weights = testing_data$exposure,
   predict_function = custom_predict,
   label = "neural_net"
 )
@@ -62,7 +66,7 @@ fi <- ingredients::feature_importance(
       sum(((observed - predicted) ^ 2 * weights) /sum(weights))
     )
   },
-  weights = testing_data$exposure,
+  # weights = testing_data$exposure,
   variables = predictors,
   B = 10,
   n_sample = 50000
@@ -71,11 +75,14 @@ fi <- ingredients::feature_importance(
 fi_plot <- fi %>% 
   as.data.frame() %>% 
   (function(df) {
+    df <- df %>% 
+      group_by(variable) %>% 
+      summarize(dropout_loss = mean(dropout_loss))
     full_model_loss <- df %>% 
       filter(variable == "_full_model_") %>% 
       pull(dropout_loss)
     df %>% 
-      filter(!variable %in% c("_full_model_", "_baseline_")) %>% 
+      filter(!variable %in% c("_full_model_", "_baseline_")) %>%
       ggplot(aes(x = reorder(variable, dropout_loss), y = dropout_loss)) +
       geom_bar(stat = "identity", alpha = 0.8) +
       geom_hline(yintercept = full_model_loss, col = "red", linetype = "dashed")+
@@ -92,29 +99,34 @@ fi_plot <- fi %>%
 ggsave("manuscript/figures/fi-plot.png", plot = fi_plot)
 
 sample_row <- testing_data[1,] %>% 
-  select(predictors)
+  select(!!predictors)
 breakdown <- iBreakDown::break_down(explainer_nn, sample_row)
 
 df <- breakdown %>% 
   as.data.frame() %>% 
-  mutate(start = lag(cummulative, default = first(contribution)),
[…]

### 2020-03-03 manuscript/manuscript.Rmd
add reference to ALE
update thanks

@@
 marginal distribution of the rest of the variables, we may take into account
-unrealistic data (e.g. high vehicle age for a model that is brand new).
+unrealistic data (e.g. high vehicle age for a model that is brand new). To
+address this drawback, alternative visualization techniques have been proposed,
+such as accumulated local effect (ALE) plots, which take expectations over
+conditional, rather than marginal, distributions [@apley2016visualizing].
 
@@
 
-We thank Navdeep Gill for his presentation on interpretable ML at SDSS and
-helpful follow-up discussions, and Daniel Falbel for his assistance with R and 
-TensorFlow.
+We thank Navdeep Gill, Daniel Falbel, Morgan Bugbee, the volunteers of the CAS
+project oversight group, and two anonymous reviewers for helpful discussions
+and/or feedback.
 This work is supported by the Casualty Actuarial Society.

### 2020-03-03 manuscript/manuscript.Rmd

@@
 context, the techniques explored in this paper (and many of the concerns discussed)
-may also be relevant to other contexts, such as claim-level reserving / analytics
+may also be relevant to other contexts, such as claim-level reserving or analytics,
 or other applications of machine learning to the insurance industry.
@@
 know the method employed to make an objective appraisal of the reasonableness 
-of the result. More information is necessary.
+of the result.
 

### 2020-03-05 manuscript/manuscript.Rmd
journal name

@@
 
-journal: "[journal]"
+journal: "Variance"
 date: "`r Sys.Date()`"
@@
 
+The techniques we utilize to answer these questions are permutation variable
+importance, partial dependence plots, and additive variable attributions, 
+respectively.
 In our discussion, we adopt the organization of techniques and some notation

### 2020-03-05 manuscript.tex
Deploy kasaai/explain-ml-pricing to github.com/kasaai/explain-ml-pricing.git:gh-pages

--- 2020-03-05 manuscript/manuscript.Rmd
@@
----
-title: Towards Explainability of Machine Learning Models in Insurance Pricing
-author:
-  - name: Kevin Kuo
-    email: kevin@kasa.ai
-    affiliation: a
-    footnote: Corresponding Author
-  - name: Daniel Lupton
-    email: dlupton@taylorandmulder.com
-    affiliation: b
-address:
-  - code: a
-    address: Kasa AI, 3040 78th Ave SE \#1271, Mercer Island, WA 98040, USA
-  - code: b
-    address: Taylor & Mulder, 10508 Rivers Bend Lane, Potomac, MD 20854, USA
-abstract: |
+\documentclass[preprint, 3p, twocolumn, letterpaper, 10pt]{elsarticle} %review=doublespace preprint=single 5p=2 column
+%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
+\usepackage[hyphens]{url}
+
+  \journal{Variance} % Sets Journal name
+
+\usepackage{lineno} % add
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+\usepackage{graphicx}
+\usepackage{booktabs} % book-quality tables
+%%%%%%%%%%%%%%%% end my additions to header
+
+\usepackage[T1]{fontenc}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \usepackage{fontspec}
+  \ifxetex
+    \usepackage{xltxtra,xunicode}
+  \fi
+  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
+  \newcommand{\euro}{€}
+\fi
+% use microtype if available
+\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\bibliographystyle{elsarticle-harv}
+\usepackage{longtable}
+\ifxetex
+  \usepackage[setpagesize=false, % page size defined by xetex
+              unicode=false, % unicode breaks when used with xetex
+              xetex]{hyperref}
+\else
+  \usepackage[unicode=true]{hyperref}
+\fi
+\hypersetup{breaklinks=true,
+            bookmarks=true,
+            pdfauthor={},
+            pdftitle={Towards Explainability of Machine Learning Models in Insurance Pricing},
+            colorlinks=false,
+            urlcolor=blue,
+            linkcolor=magenta,
+            pdfborder={0 0 0}}
+\urlstyle{same}  % don't use monospace font for urls
+
+\setcounter{secnumdepth}{5}
+% Pandoc toggle for numbering sections (defaults to be off)
+
+% Pandoc header
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
+
+\begin{document}
+\begin{frontmatter}
+
+  \title{Towards Explainability of Machine Learning Models in Insurance Pricing}
+    \author[a]{Kevin Kuo\corref{Corresponding Author}}
+   \ead{kevin@kasa.ai} 
+    \author[b]{Daniel Lupton}
+   \ead{dlupton@taylorandmulder.com} 
+      \address[a]{Kasa AI, 3040 78th Ave SE \#1271, Mercer Island, WA 98040, USA}
+    \address[b]{Taylor \& Mulder, 10508 Rivers Bend Lane, Potomac, MD 20854, USA}
+
+  \begin{abstract}
   Machine learning methods have garnered increasing interest among actuaries in
@@
   linear models. In this paper, we discuss the need for model interpretability in
-  property & casualty insurance ratemaking, propose a framework for explaining 
+  property \& casualty insurance ratemaking, propose a framework for explaining
   models, and present a case study to illustrate the framework.
+  \end{abstract}
 
-journal: "Variance"
-date: "`r Sys.Date()`"
-bibliography: explain-pricing.bib
-#linenumbers: true
-numbersections: true
-# csl: elsevier-harvard.csl
-output: 
-  bookdown::pdf_book:
-    base_format: rticles::elsevier_article
-layout: preprint, 3p, twocolumn, letterpaper, 10pt
----
+ \end{frontmatter}
 
-# Introduction
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
-Risk classification for property & casualty (P&C) insurance rating has traditionally
+Risk classification for property \& casualty (P\&C) insurance rating has traditionally
 been done with one-way, or univariate, analysis techniques. In recent years, many
@@
 neural networks have gained popularity in many industries due to their superior
-predictive performance over linear models [@lecunDeepLearning2015]. In fact, there
-is a fast growing body of literature on applying ML to P&C reserving
-[@kuoDeepTriangleDeep2018; @wuthrichMachineLearning2018; @gabrielliNeuralNetwork2019a;
-@gabrielliNeuralNetwork2019]. However, these ML techniques, often considered to be
-completely “black box”, have been less successful in gaining adoption in pricing, 
+predictive performance over linear models (LeCun, Bengio, and Hinton 2015). In fact, there
+is a fast growing body of literature on applying ML to P\&C reserving
+(Kuo 2019; Wüthrich 2018; Gabrielli, Richman, and Wüthrich 2019; Gabrielli 2019). However, these ML techniques, often considered to be
+completely ``black box'', have been less successful in gaining adoption in pricing,
 which is a regulated discipline and requires a certain amount of transparency in models.
@@
 The rest of this paper is organized as follows: Section \ref{ratemaking} provides
-an overview of P&C ratemaking, Section \ref{need} discusses the importance of 
+an overview of P\&C ratemaking, Section \ref{need} discusses the importance of
 interpretation, Section \ref{interpretability} discusses model interpretability
@@
 
-# Property and Casualty Ratemaking {#ratemaking}
+\hypertarget{ratemaking}{%
+\section{Property and Casualty Ratemaking}\label{ratemaking}}
 
-## History of Ratemaking
+\hypertarget{history-of-ratemaking}{%
+\subsection{History of Ratemaking}\label{history-of-ratemaking}}
 
 Early classification ratemaking procedures were typically univariate in nature.
-For example, @lange_1966 notes that (at that time) most major lines of insurance 
+For example, Lange (1966) notes that (at that time) most major lines of insurance
 used univariate methods based around the same principle: distributing an overall
@@
 
-@bailey_simon_1960 introduced minimum bias methods, which were expanded 
+Bailey and Simon (1960) introduced minimum bias methods, which were expanded
 throughout the 60s, 70s, and 80s. As computing power developed, minimum bias
-began to give away to GLMs, with papers such as @brown_1988 and @mildenhall_1999 
+began to give away to GLMs, with papers such as Brown (1988) and Mildenhall (1999)
 bridging the gap between the methods.
@@
 Arguably, GLMs predate minimum bias procedures by a significant margin. The term
-was coined by @nelder_wedderburn_1972, but generalizations of least squares 
+was coined by Nelder and Wedderburn (1972), but generalizations of least squares
 linear regression date back at least to the 1930s. Like minimum bias methods,
@@
 seem to include any mention of GLMs
-prior to @brown_1988 in the 1990 syllabus. From there, GLMs 
+prior to Brown (1988) in the 1990 syllabus. From there, GLMs
 seem to have received only passing mention until 2006 with the introduction of
-@anderson_2005 to the syllabus. Beginning in 2016, the CAS introduced @goldburd_2016
+Anderson et al. (2005) to the syllabus. Beginning in 2016, the CAS introduced Goldburd, Khare, and Tevet (2016)
 to the syllabus, which offers a comprehensive guide to GLMs.
 
-## Machine Learning in Ratemaking
+\hypertarget{machine-learning-in-ratemaking}{%
+\subsection{Machine Learning in Ratemaking}\label{machine-learning-in-ratemaking}}
 
@@
 algorithms throughout the middle part of the 20th century. Detailed histories
-of machine learning may be found in sources such as @nilsson_2009 and 
-@wang_raj_2017. Consistent with GLMs, machine learning was relatively unpopular 
+of machine learning may be found in sources such as Nilsson (2009) and
+Wang and Raj (2017). Consistent with GLMs, machine learning was relatively unpopular
 in actuarial science until the last ten years as computing power has become cheaper
@@
 the first time it entered the CAS syllabus; however, cluster analysis (in the
-form of k-means) seems to have been first included in 2011 with @robertson_2009.
+form of k-means) seems to have been first included in 2011 with Robertson (2009).
 More recently, the CAS MAS-I and MAS-II exams introduced in 2018 have included
@@
 significant portion of machine learning applications to ratemaking has been in
-the context of automobile telematics, such as @gao_2018, @gao_2018_2, @gao_2019,
-@roel_2018, or @wuthrich_2017. Presumably this focus has been a result of the 
+the context of automobile telematics, such as Gao, Meng, and Wuthrich (2018), Gao and Wuthrich (2018), Gao and Wuthrich (2019),
+Roel, Antonio, and Claeskens (2018), or Wuthrich (2017). Presumably this focus has been a result of the
 high-dimensionality and complexity of telematics data, making it a field in
@@
 
-Outside of telematics, @yang_2018 use a gradient tree-boosting approach to 
-capture non-linearities that would be a challenge for GLMs. @henckaerts_2018 
+Outside of telematics, Yang, Qian, and Zou (2018) use a gradient tree-boosting approach to
+capture non-linearities that would be a challenge for GLMs. Henckaerts et al. (2018)
 make use of generalized additive models (GAM) to improve predictions of GLMs.
@@
 different machine learning algorithms within a single study, such as in
-@dugas_2003, @noll_2018, @spedicato_2018. These studies make use of such varied
+Dugas et al. (2003), Noll, Salzmann, and Wuthrich (2018), Spedicato, Dutang, and Petrini (2018). These studies make use of such varied
 techniques as regression trees, boosting machines, support vector machines, and
@@
 
-## Ratemaking Process
+\hypertarget{ratemaking-process}{%
+\subsection{Ratemaking Process}\label{ratemaking-process}}
 
@@
 
-1. Obtain relevant policy-level data
-2. Prepare data for analysis
-3. Perform analysis on the data, employing desired method(s) to estimate
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Obtain relevant policy-level data
+\item
+  Prepare data for analysis
+\item
+  Perform analysis on the data, employing desired method(s) to estimate
   needed rates
-4. Select final rates based on rate indications
-5. Present rates to the reglator, including explanation of the steps followed to 
+\item
+  Select final rates based on rate indications
+\item
+  Present rates to the reglator, including explanation of the steps followed to
   derive the rates
-6. Answer questions from regulators regarding the method employed
+\item
+  Answer questions from regulators regarding the method employed
+\end{enumerate}
 
@@
 part of a risk classification system per Actuarial Standard of Practice 12
-("Risk Classification"). Therefore, the techniques discussed 
+(``Risk Classification''). Therefore, the techniques discussed
 in this paper may be viewed from the lens of providing useful information to
@@
 
-# The Need to See Inside the Black Box {#need}
+\hypertarget{need}{%
+\section{The Need to See Inside the Black Box}\label{need}}
 
-Within the actuarial profession, Actuarial Standard of Practice 41 ("Actuarial 
-Communications") notes that "...another actuary qualified in the same practice 
-area [should be able to] make an objective appraisal of the reasonableness of 
-the actuary's work as presented in the actuarial report." [@asop_41] Underlying 
+Within the actuarial profession, Actuarial Standard of Practice 41 (``Actuarial
+Communications'') notes that ``\ldots{}another actuary qualified in the same practice
+area {[}should be able to{]} make an objective appraisal of the reasonableness of
+the actuary's work as presented in the actuarial report.'' (``Actuarial Standard of Practice No. 41 - Actuarial Communications'' 2010) Underlying
 this requirement is an assumption that the hypothetical other actuary qualified
@@
 cover a wide range of topics, such as splines, clustering algorithms, decision
-trees, boosting, and principle components analysis [@cas_syllabus_2018].
+trees, boosting, and principle components analysis (Casualty Actuarial Society 2018).
 
@@
 
-1. Machine learning models can be very ad hoc compared to 
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Machine learning models can be very ad hoc compared to
   traditional statistical models.
-2. Because many machine learning models do not assume an underlying probability
+\item
+  Because many machine learning models do not assume an underlying probability
   distribution or stochastic process, they may not admit of standard metrics for model comparison
   (e.g., it's not straightforward to calculate an AIC over a neural network).
-3. Machine learning methods are often combined into ensembles that 
+\item
+  Machine learning methods are often combined into ensembles that
   may not be easily separated and that may, as a collection, cease to resemble
   a single standard version of a model.
-4. Machine learning models can be "black boxes" insofar as the final form of 
+\item
+  Machine learning models can be ``black boxes'' insofar as the final form of
   response curve cannot be easily predicted and may depend heavily on the
   available data (which may not, in turn, be available to the reviewer).
+\end{enumerate}
 
@@
 actuaries to be able to exercise appropriate judgment about risk classification
-systems. The recent ASOP 56 ("Models") speaks to more general concerns in all
+systems. The recent ASOP 56 (``Models'') speaks to more general concerns in all
 practice areas that might make use of models. ASOP 56 requires the actuary to
-"make reasonable efforts to confirm that the model structure, data, assumptions,
+``make reasonable efforts to confirm that the model structure, data, assumptions,
 governance and controls, and model testing and output validation are consistent
-with the intended purpose." [@asop_56] All of these efforts may be hampered if
+with the intended purpose.'' (``Actuarial Standard of Practice No. 56 - Modeling'' 2019) All of these efforts may be hampered if
 it is not possible to peer into the black box of the model.
@@
 the National Association of Insurance Commissioners (NAIC) found that the plurality of
-responding regulators identified "Filing complexity and/or a lack of resources 
-or expertise" as a key challenge that impedes their ability to review GLMs or 
-other predictive models [@naic_summer_2017]. Given that machine learning 
+responding regulators identified ``Filing complexity and/or a lack of resources
+or expertise'' as a key challenge that impedes their ability to review GLMs or
+other predictive models (National Association of Insurance Commissioners 2017). Given that machine learning
 algorithms are generally regarded as more complex than GLMs, this implies that
@@
 in reviewing rate filings based on GLMs, and 34 noted that it would be helpful
-to develop similar items to assist in reviewing "Other Advanced Modeling 
-Techniques." One outgrowth of this need was the development of a white paper,
-@naic_white_paper, on best practices for regulatory review of predictive models.
+to develop similar items to assist in reviewing ``Other Advanced Modeling
+Techniques.'' One outgrowth of this need was the development of a white paper,
+National Association of Insurance Commissioners (2019), on best practices for regulatory review of predictive models.
 The white paper focuses on review of GLMs, particularly with respect to private
@@
 
-* Determine the extent to which the model causes premium disruption for individual 
+\begin{itemize}
+\tightlist
+\item
+  Determine the extent to which the model causes premium disruption for individual
   policyholders and how the insurer will
   explain the disruption to individual consumers that inquire about it.
-* Determine that individual input characteristics to a predictive model are 
+\item
+  Determine that individual input characteristics to a predictive model are
   related to the expected loss or expense
@@
   demonstrable actual relationship to expected loss or expense.
-* Determine that individual outputs from a predictive model and their associated
+\item
+  Determine that individual outputs from a predictive model and their associated
   selected relativities are not unfairly discriminatory.
+\end{itemize}
 
@@
 
-# Interpretability in the Ratemaking Context {#interpretability}
+\hypertarget{interpretability}{%
+\section{Interpretability in the Ratemaking Context}\label{interpretability}}
 
@@
 
-- Ability to explain or to present in understandable terms to a human 
-[@doshi-velezRigorousScience2017];
-- The degree to which an observer can understand the cause of a decision 
-[@biranExplanationJustification2017]; and
-- A method is interpretable if a user can correctly and efficiently predict the 
-method’s results [@kimExamplesAre2016].
+\begin{itemize}
+\tightlist
+\item
+  Ability to explain or to present in understandable terms to a human
+  (Doshi-Velez and Kim 2017);
+\item
+  The degree to which an observer can understand the cause of a decision
+  (Biran and Cotton 2017); and
+\item
+  A method is interpretable if a user can correctly and efficiently predict the
+  method's results (Kim, Khanna, and Koyejo 2016).
+\end{itemize}
 
@@
 model to be interpretable. In the remainder of this section, we clarify a couple concepts regarding
-interpretable *classes* of models and the computational transparency of ML
+interpretable \emph{classes} of models and the computational transparency of ML
 models, outline frameworks for understanding the communication goals of interpretability,
@@
 
-## Not All Linear Models are Interpretable {#linear-models}
+\hypertarget{linear-models}{%
+\subsection{Not All Linear Models are Interpretable}\label{linear-models}}
 
@@
 consider a claim severity model with driver age, sex, and vehicle age as
-predictors; assuming a log link function and letting $Y$ denote the response, we
+predictors; assuming a log link function and letting \(Y\) denote the response, we
 have
@@
 Here, we can tell, for example, what the model would predict for the expected
-severity if we were to increase age by a certain amount, *all else being equal*,
+severity if we were to increase age by a certain amount, \emph{all else being equal},
 because the relationship between the predictor and the response is simply
-multiplication by the coefficient $\beta_1$ and applying the inverse link
+multiplication by the coefficient \(\beta_1\) and applying the inverse link
 function.
@@
 Another commonly cited example of an interpretable model is a decision tree. An
-illustrative example is shown in Figure \@ref(fig:tree-plot1). Here, the 
+illustrative example is shown in Figure \ref{fig:tree-plot1}. Here, the
 prediction is arrived at by following a sequence of if-else decisions.
 
-```{r tree-plot1, echo = FALSE, fig.cap="A simple decision tree for loss cost prediction.", out.width="100%", fig.align='center'}
-knitr::include_graphics("figures/tree-plot1.png")
-```
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figures/tree-plot1}
+
+}
+
+\caption{A simple decision tree for loss cost prediction.}\label{fig:tree-plot1}
+\end{figure}
 
@@
 
-Similarly, one can see that in Figure \@ref(fig:tree-plot2), larger trees are
+Similarly, one can see that in Figure \ref{fig:tree-plot2}, larger trees are
 tough to reason about. In other words, even when working within the framework
-of an "interpretable" class of models, we may still end up with something that
-many would consider "black box."
+of an ``interpretable'' class of models, we may still end up with something that
+many would consider ``black box.''
 
-```{r tree-plot2, echo = FALSE, fig.cap="A more complex decision tree. This is still much simpler than typical realistic examples.", out.width="60%", fig.align='center',  fig.env='figure*'}
-knitr::include_graphics("figures/tree-plot2.png")
-```
+\begin{figure*}
 
-## The Machinery is Not a Secret
+{\centering \includegraphics[width=0.6\linewidth]{figures/tree-plot2}
+
+}
+
+\caption{A more complex decision tree. This is still much simpler than typical realistic examples.}\label{fig:tree-plot2}
+\end{figure*}
+
+\hypertarget{the-machinery-is-not-a-secret}{%
+\subsection{The Machinery is Not a Secret}\label{the-machinery-is-not-a-secret}}
 
 Another occasional misconception is that we have no visibility into how some ML
-models *compute* predictions, which renders them uninterpretable. Outside of 
+models \emph{compute} predictions, which renders them uninterpretable. Outside of
 proprietary algorithms, all common ML models, including neural networks,
@@
 some inputs, but it may be difficult for a human to reason about it. We show
-later that we can still provide explanations of completely "black box" models,
+later that we can still provide explanations of completely ``black box'' models,
 but is important to note that ML model predictions are still governed by
@@
 
-## Explanations are Contextual
+\hypertarget{explanations-are-contextual}{%
+\subsection{Explanations are Contextual}\label{explanations-are-contextual}}
 
-@hiltonConversationalProcesses1990 proposed a framework, later interpreted by
-@millerExplanationArtificial2017 in the context of ML, for understanding model explanations
-as *conversations* or *social interactions.* One consequence of this identification
-is that explanations need to be *relevant* to the *audience*. This framework is
+Hilton (1990) proposed a framework, later interpreted by
+Miller (2017) in the context of ML, for understanding model explanations
+as \emph{conversations} or \emph{social interactions.} One consequence of this identification
+is that explanations need to be \emph{relevant} to the \emph{audience}. This framework is
 consistent with ASOP 41, which formulates a similar requirement in terms of an
-*intended user* of the actuarial communication. In developing,
+\emph{intended user} of the actuarial communication. In developing,
 filing, and operationalizing a pricing model, one needs to accommodate a variety
@@
 models, tend to be same people designing and communicating the explanations.
-This poses a challenge that @millerExplainableAI2017 call "inmates running
-the asylum", where the modelers design explanations for *themselves* rather
+This poses a challenge that Miller, Howe, and Sonenberg (2017) call ``inmates running
+the asylum'', where the modelers design explanations for \emph{themselves} rather
 than the intended audience. For example, they may be interested in technical
@@
 
-Another point outlined in Miller's survey [@millerExplanationArtificial2017] is
-that explanations are *contrastive*. In other words, people are often interested
+Another point outlined in Miller's survey (Miller 2017) is
+that explanations are \emph{contrastive}. In other words, people are often interested
 in not why something happened, but rather why it happened instead of something
@@
 
-## Asking and Answering the Right Questions
+\hypertarget{asking-and-answering-the-right-questions}{%
+\subsection{Asking and Answering the Right Questions}\label{asking-and-answering-the-right-questions}}
 
@@
 These questions should relate to existing guidelines, such as those described in
-@naic_white_paper and outlined in Section \ref{need}, standards of practice, and
+National Association of Insurance Commissioners (2019) and outlined in Section \ref{need}, standards of practice, and
 regulation, and in fact should not be specific only to ML models. By
@@
 
-# Applying Model Interpretation Techniques {#application}
+\hypertarget{application}{%
+\section{Applying Model Interpretation Techniques}\label{application}}
 
@@
 
-1. What are the most important predictors in the model? Put another way, to
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  What are the most important predictors in the model? Put another way, to
   what extent do the predictors improve the accuracy of the model?
-2. How does the predicted loss cost change, on average, as we change an input?
-3. For a particular policyholder, how does each characteristic contribute to
+\item
+  How does the predicted loss cost change, on average, as we change an input?
+\item
+  For a particular policyholder, how does each characteristic contribute to
   the loss cost prediction?
+\end{enumerate}
 
@@
 In our discussion, we adopt the organization of techniques and some notation
-presented in @molnar2018interpretable and @biecekPMVEE, which are comprehensive
+presented in Molnar and others (2018) and Biecek and Burzykowski (2019), which are comprehensive
 references on the most established ML interpretation techniques.
 
-## A Simplified View of Interpretation Techniques
+\hypertarget{a-simplified-view-of-interpretation-techniques}{%
+\subsection{A Simplified View of Interpretation Techniques}\label{a-simplified-view-of-interpretation-techniques}}
 
@@
 model interpretation techniques can be categorized across two dimensions:
-intrinsic vs. post-hoc and global vs. local.
+intrinsic vs.~post-hoc and global vs.~local.
 
-### Intrinsic vs. Post-hoc
+\hypertarget{intrinsic-vs.post-hoc}{%
+\subsubsection{Intrinsic vs.~Post-hoc}\label{intrinsic-vs.post-hoc}}
 
 Intrinsic model interpretation draws conclusions from the structure of the
-fitted model and are what we typically associate with "interpretable" classes
+fitted model and are what we typically associate with ``interpretable'' classes
 of models. This is only viable with models with simple structures, such as the
@@
 prediction behavior of the model. These techniques are concerned with only the
-inputs and outputs, and hence are *agnostic* of the model itself, which means
+inputs and outputs, and hence are \emph{agnostic} of the model itself, which means
 they can also be applied to simple models. Since most useful ML models have a
@@
 
-### Global vs. Local
+\hypertarget{global-vs.local}{%
+\subsubsection{Global vs.~Local}\label{global-vs.local}}
 
@@
 with respect to individual predictors. Note that these methods may be compared to
-the methods described in [@goldburd_2016], Chapter 7, which focus on global
+the methods described in (Goldburd, Khare, and Tevet 2016), Chapter 7, which focus on global
 interpretation of GLMs.
@@
 
-## Answering the Questions
+\hypertarget{answering-the-questions}{%
+\subsection{Answering the Questions}\label{answering-the-questions}}
 
@@
 
-### Variable Importance
+\hypertarget{variable-importance}{%
+\subsubsection{Variable Importance}\label{variable-importance}}
 
-"What are the most important predictors in the model?"
+``What are the most important predictors in the model?''
 
@@
 case of GLMs, one might observe the magnitudes of the estimated coefficients or
-$t$-statistics, whereas
-for random forests, one might use out-of-bag errors [@breimanRandomForests2001].
+\(t\)-statistics, whereas
+for random forests, one might use out-of-bag errors (Breiman 2001).
 For more complex models, such as the neural network in our case study, we need
@@
 We follow the methodology of permutation feature importance as described in
-@fisherAllModels2018, and utilize the notation introduced by @biecekPMVEE.
+Fisher, Rudin, and Dominici (2018), and utilize the notation introduced by Biecek and Burzykowski (2019).
 The gist of the technique is as follows: to see how important a variable is,
@@
 
-More formally, let $y$ denote the vector of responses, $X$ denote the matrix of
-predictor variables, $\widehat{f}$ denote the fitted model, and
-$L = \mathcal{L}(\widehat{f}(X), y)$, where $\widehat{f}$ applies to $X$
+More formally, let \(y\) denote the vector of responses, \(X\) denote the matrix of
+predictor variables, \(\widehat{f}\) denote the fitted model, and
+\(L = \mathcal{L}(\widehat{f}(X), y)\), where \(\widehat{f}\) applies to \(X\)
 rowwise, denote the value of the loss function, which is mean squared error
-in the case of regression. Now, if $\widetilde{X}^{j}$ denotes the predictor
-matrix where the $j$th variable has been permuted, then we can compute the
+in the case of regression. Now, if \(\widetilde{X}^{j}\) denotes the predictor
+matrix where the \(j\)th variable has been permuted, then we can compute the
 loss with the permuted dataset as
-$L^{-j} = \mathcal{L}(\widehat{f}(\widetilde{X}^{j}), y)$. 
-Here, by *permuting* a variable, we mean that we randomly rearrange the values 
+\(L^{-j} = \mathcal{L}(\widehat{f}(\widetilde{X}^{j}), y)\).
+Here, by \emph{permuting} a variable, we mean that we randomly rearrange the values
 in the column of data associated with the variable.
 With this, we define
-the variable importance $VI^{j}$ as $L^{-j} - L$.
+the variable importance \(VI^{j}\) as \(L^{-j} - L\).
 
-In Figure \@ref(fig:fi), we show a plot of variable importances. In our
-particular example, we see that the "make" variable contributes most to the
-accuracy of the model with "sex" contributing the least. This provides a way
+In Figure \ref{fig:fi}, we show a plot of variable importances. In our
+particular example, we see that the ``make'' variable contributes most to the
+accuracy of the model with ``sex'' contributing the least. This provides a way
 for the audience to quickly glance at the most relevant variables, and ask
@@
 
-```{r fi, out.width='100%', fig.align = "center", fig.cap="Permutation feature importances for the neural network model.", echo = FALSE}
-knitr::include_graphics("figures/fi-plot.png")
-```
+\begin{figure}
 
-### Partial Dependence Plots
+{\centering \includegraphics[width=1\linewidth]{figures/fi-plot}
 
-"How does the predicted loss cost change, on average, as we change an input?"
+}
+
+\caption{Permutation feature importances for the neural network model.}\label{fig:fi}
+\end{figure}
+
+\hypertarget{partial-dependence-plots}{%
+\subsubsection{Partial Dependence Plots}\label{partial-dependence-plots}}
+
+``How does the predicted loss cost change, on average, as we change an input?''
 
@@
 are available, we can appeal to partial dependence plots (PDP), first proposed
-by @friedmanGreedyFunction2001 for gradient boosting machines (GBM).
+by Friedman (2001) for gradient boosting machines (GBM).
 
-```{r pdp, out.width='70%', fig.align = "center", fig.cap="Partial dependence plot for the neural network model.", echo = FALSE,  fig.env='figure*', fig.pos='h'}
-knitr::include_graphics("figures/pdp-plot.png")
-```
+\begin{figure*}[h]
 
-```{r breakdown, out.width='80%', fig.align = "center", fig.cap="Variable contribution plot for the neural network model.", echo = FALSE, fig.env = 'figure*', fig.pos = 'h'}
-knitr::include_graphics("figures/breakdown-plot.png")
-```
+{\centering \includegraphics[width=0.7\linewidth]{figures/pdp-plot}
 
-To describe PDP, we need to introduce some additional notation. Let $x^j$ denote
+}
+
+\caption{Partial dependence plot for the neural network model.}\label{fig:pdp}
+\end{figure*}
+
+\begin{figure*}[h]
+
+{\centering \includegraphics[width=0.8\linewidth]{figures/breakdown-plot}
+
+}
+
+\caption{Variable contribution plot for the neural network model.}\label{fig:breakdown}
+\end{figure*}
+
+To describe PDP, we need to introduce some additional notation. Let \(x^j\) denote
 the input variable of interest. Then we define the partial dependence function as
@@
 relationship between the predicted response and the variable of interest.
-Empirically, we estimate $h$ by
+Empirically, we estimate \(h\) by
 
@@
 
-where $N$ is the number of records in the dataset.
+where \(N\) is the number of records in the dataset.
 
-In Figure \@ref(fig:pdp), we exhibit the PDP for the "vehicle age" variable.
+In Figure \ref{fig:pdp}, we exhibit the PDP for the ``vehicle age'' variable.
 We see that the average predicted loss cost decreases with vehicle age until
@@
 marginal distribution of the rest of the variables, we may take into account
-unrealistic data (e.g. high vehicle age for a model that is brand new). To
+unrealistic data (e.g.~high vehicle age for a model that is brand new). To
 address this drawback, alternative visualization techniques have been proposed,
 such as accumulated local effect (ALE) plots, which take expectations over
-conditional, rather than marginal, distributions [@apley2016visualizing].
+conditional, rather than marginal, distributions (Apley and Zhu 2016).
 
-### Variable Attribution
+\hypertarget{variable-attribution}{%
+\subsubsection{Variable Attribution}\label{variable-attribution}}
 
-"For a particular policyholder, how does each characteristic contribute to the
-loss cost prediction?"
+``For a particular policyholder, how does each characteristic contribute to the
+loss cost prediction?''
 
@@
 
-Formally, for a fitted model $\widehat{f}$, a given ordering of the variables
-$X^1,\dots,X^p$, where $p$ is the number of predictor variables,
-and a specific instance $x_*$, we
-would like to decompose the model prediction $\widehat{f}(x_*)$ into
+Formally, for a fitted model \(\widehat{f}\), a given ordering of the variables
+\(X^1,\dots,X^p\), where \(p\) is the number of predictor variables,
+and a specific instance \(x_*\), we
+would like to decompose the model prediction \(\widehat{f}(x_*)\) into
 
@@
 
-where $v_0$ denotes the average model response, and $v(j, x_*)$ denotes the 
-contribution of the $j$th variable in instance $x_*$, defined as
+where \(v_0\) denotes the average model response, and \(v(j, x_*)\) denotes the
+contribution of the \(j\)th variable in instance \(x_*\), defined as
 
@@
 
-Hence, the contribution of the $j$th variable to the prediction is the 
-incremental change in the expected model prediction when we set $X^j = x_*^ j$
-assuming the other variables take their values in $x_*$. Note here that this
+Hence, the contribution of the \(j\)th variable to the prediction is the
+incremental change in the expected model prediction when we set \(X^j = x_*^ j\)
+assuming the other variables take their values in \(x_*\). Note here that this
 definition implies that the order in which we consider the variables affects
-the results. Empirically, the expectations in \@ref(eq:var-contrib) are 
+the results. Empirically, the expectations in \eqref{eq:var-contrib} are
 calculated by sampling the test dataset.
 
-In Figure \@ref(fig:breakdown), we exhibit a waterfall plot of variable
-contributions. The "intercept" value denotes the average model prediction and 
-represents the $v_0$ term in Equation \@ref(eq:contrib-decompose). The predicted
+In Figure \ref{fig:breakdown}, we exhibit a waterfall plot of variable
+contributions. The ``intercept'' value denotes the average model prediction and
+represents the \(v_0\) term in Equation \eqref{eq:contrib-decompose}. The predicted
 loss cost for this particular policyholder is slightly less than average; the
@@
 including interactions terms, and averaging over possible orderings. The last
-of these ideas is implemented by @NIPS2017_7062 using Shapley values from 
+of these ideas is implemented by Lundberg and Lee (2017) using Shapley values from
 cooperative game theory, and is referred to as Shapley additive explanations
-(SHAP). These approaches are discussed further in @biecekPMVEE and its references.
-Shapley values were also used in @mango_1998, which has appeared on the CAS syllabus
+(SHAP). These approaches are discussed further in Biecek and Burzykowski (2019) and its references.
+Shapley values were also used in Mango (1998), which has appeared on the CAS syllabus
 starting in 2004, in the context of determining how to allocate catastrophe
@@
 
-## Other Techniques
+\hypertarget{other-techniques}{%
+\subsection{Other Techniques}\label{other-techniques}}
 
@@
 instance-level components for a more granular view into predictor sensitivities
-[@goldstein2015peeking]. To accommodate correlated variables in PDP, accumulated 
+(Goldstein et al. 2015). To accommodate correlated variables in PDP, accumulated
 local effect (ALE) plots computes expected changes in model response over the
 conditional, rather than marginal, distribution of the other variables
-[@apley2016visualizing].
+(Apley and Zhu 2016).
 
-Local interpretable model-agnostic explanations (LIME) [@ribeiro2016should] 
+Local interpretable model-agnostic explanations (LIME) (Ribeiro, Singh, and Guestrin 2016)
 builds simple surrogate models using model predictions, with higher training
@@
 
-# Conclusion {#conclusion}
+\hypertarget{conclusion}{%
+\section{Conclusion}\label{conclusion}}
 
@@
 pricing models. These responsibilities create special challenges for
-communicating machine learning models, which are often seen "black boxes" due
+communicating machine learning models, which are often seen ``black boxes'' due
 in part to their complexity, nonlinearity, flexible construction, and ad hoc
@@
 
-* How does the model impact individual insurance consumers?
-* How are the predictor variables related to expected losses?
+\begin{itemize}
+\tightlist
+\item
+  How does the model impact individual insurance consumers?
+\item
+  How are the predictor variables related to expected losses?
+\end{itemize}
 
@@
 
-Acknowledgments {.unnumbered}
-==========
+\hypertarget{acknowledgments}{%
+\section*{Acknowledgments}\label{acknowledgments}}
+\addcontentsline{toc}{section}{Acknowledgments}
 
@@
 
-References {#references .unnumbered}
-==========
-<div id="refs"></div>
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-tensorflow2015-whitepaper}{}%
+Abadi, Martin, Ashish Agarwal, Paul Barham, Eugene Brevdo, Zhifeng Chen, Craig Citro, Greg S.
+Corrado, et al.
+2015.
+``TensorFlow: Large-Scale Machine Learning on Heterogeneous Systems.'' \url{http://tensorflow.org/}.
+
+\leavevmode\hypertarget{ref-asop_41}{}%
+``Actuarial Standard of Practice No. 41 - Actuarial Communications.'' 2010. Actuarial Standards Board. \url{http://www.actuarialstandardsboard.org/wp-content/uploads/2014/02/asop041_120.pdf}.
+
+\leavevmode\hypertarget{ref-asop_56}{}%
+``Actuarial Standard of Practice No. 56 - Modeling.'' 2019. Actuarial Standards Board. \url{http://www.actuarialstandardsboard.org/wp-content/uploads/2020/01/asop056_195.pdf}.
+
+\leavevmode\hypertarget{ref-anderson_2005}{}%
+Anderson, D., S. Feldblum, C. Modlin, D. Schirmacher, E. Schirmacher, and N. Thandi. 2005. ``A Practitioner's Guide to Generalized Linear Models,'' 4--39.
+
+\leavevmode\hypertarget{ref-apley2016visualizing}{}%
+Apley, Daniel W., and Jingyu Zhu. 2016. ``Visualizing the Effects of Predictor Variables in Black Box Supervised Learning Models.''
+
+\leavevmode\hypertarget{ref-bailey_simon_1960}{}%
+Bailey, Robert A., and LeRoy J.
+Simon.
+1960.
+``Two Studies in Automobile Insurance Ratemaking.'' \emph{ASTIN Bulletin} 1 (4).
+Cambridge University Press: 192--217.
+\url{https://doi.org/10.1017/S0515036100009569}.
+
+\leavevmode\hypertarget{ref-biecekPMVEE}{}%
+Biecek, Przemyslaw, and Tomasz Burzykowski. 2019. ``Predictive Models: Explore, Explain, and Debug.'' \url{https://pbiecek.github.io/PM_VEE/}.
+
+\leavevmode\hypertarget{ref-biranExplanationJustification2017}{}%
+Biran, Or, and Courtenay Cotton. 2017. ``Explanation and Justification in Machine Learning: A Survey.'' In \emph{IJCAI-17 Workshop on Explainable AI (XAI)}, 8:1.
+
+\leavevmode\hypertarget{ref-breimanRandomForests2001}{}%
+Breiman, Leo. 2001. ``Random Forests.'' \emph{Machine Learning} 45 (1): 5--32.
+
+\leavevmode\hypertarget{ref-brown_1988}{}%
+Brown, Robert L. 1988. ``Minimum Bias with Generalized Linear Models,'' 187--217.
+
+\leavevmode\hypertarget{ref-cas_syllabus_2018}{}%
+Casualty Actuarial Society. 2018. ``Syllabus of Basic Education'' 2018. Casualty Actuarial Society. \url{https://www.casact.org/admissions/syllabus/ArchivedSyllabi/2018Syllabus.pdf}.
+
+\leavevmode\hypertarget{ref-doshi-velezRigorousScience2017}{}%
+Doshi-Velez, Finale, and Been Kim. 2017. ``Towards A Rigorous Science of Interpretable Machine Learning.'' \emph{arXiv:1702.08608 {[}Cs, Stat{]}}, February. \url{http://arxiv.org/abs/1702.08608}.
+
+\leavevmode\hypertarget{ref-dugas_2003}{}%
+Dugas, Charles, Y.
+Bengio, Nicolas Chapados, P.
+Vincent, G.
+Denoncourt, and C.
+Fournier.
+2003.
+``Statistical Learning Algorithms Applied to Automobile Insurance Ratemaking,'' December.
+\url{https://doi.org/10.1142/9789812794246_0004}.
+
+\leavevmode\hypertarget{ref-fisherAllModels2018}{}%
+Fisher, Aaron, Cynthia Rudin, and Francesca Dominici.
+2018.
+``All Models Are Wrong, but Many Are Useful: Learning a Variable's Importance by Studying an Entire Class of Prediction Models Simultaneously.'' \emph{arXiv:1801.01489 {[}Stat{]}}, January.
+\url{http://arxiv.org/abs/1801.01489}.
+
+\leavevmode\hypertarget{ref-friedmanGreedyFunction2001}{}%
+Friedman, Jerome H. 2001. ``Greedy Function Approximation: A Gradient Boosting Machine.'' \emph{Annals of Statistics}, 1189--1232.
+
+\leavevmode\hypertarget{ref-gabrielliNeuralNetwork2019}{}%
+Gabrielli, Andrea. 2019. ``A Neural Network Boosted Double over-Dispersed Poisson Claims Reserving Model.'' SSRN Scholarly Paper ID 3365517. Rochester, NY: Social Science Research Network.
+
+\leavevmode\hypertarget{ref-gabrielliNeuralNetwork2019a}{}%
+Gabrielli, Andrea, Ronald Richman, and Mario V. Wüthrich. 2019. ``Neural Network Embedding of the over-Dispersed Poisson Reserving Model.'' \emph{Scandinavian Actuarial Journal}, 1--29.
+
+\leavevmode\hypertarget{ref-gao_2018}{}%
+Gao, Guangyuan, Shengwang Meng, and Mario V. Wuthrich. 2018. ``Claims Frequency Modeling Using Telematics Car Driving Data.'' \emph{Scandinavian Actuarial Journal}.
+
+\leavevmode\hypertarget{ref-gao_2019}{}%
+Gao, Guangyuan, and Mario Wuthrich. 2019. ``Convolutional Neural Network Classification of Telematics Car Driving Data.'' \emph{Risks} 7 (January): 6. \url{https://doi.org/10.3390/risks7010006}.
+
+\leavevmode\hypertarget{ref-gao_2018_2}{}%
+Gao, Guangyuan, and Mario V. Wuthrich. 2018. ``Feature Extraction from Telematics Car Driving Heatmaps.'' \emph{European Actuarial Journal} 8 (2): 383--406.
+
+\leavevmode\hypertarget{ref-goldburd_2016}{}%
+Goldburd, Mark, Anand Khare, and Dan Tevet. 2016. ``Generalized Linear Models for Insurance Rating.'' \emph{Casualty Actuarial Society, CAS Monographs Series}, no. 5.
+
+\leavevmode\hypertarget{ref-goldstein2015peeking}{}%
+Goldstein, Alex, Adam Kapelner, Justin Bleich, and Emil Pitkin.
+2015.
+``Peeking Inside the Black Box: Visualizing Statistical Learning with Plots of Individual Conditional Expectation.'' \emph{Journal of Computational and Graphical Statistics} 24 (1).
+Taylor \& Francis: 44--65.
+
+\leavevmode\hypertarget{ref-henckaerts_2018}{}%
+Henckaerts, Roel, Katrien Antonio, Maxime Clijsters, and Roel Verbelen.
+2018.
+``A Data Driven Binning Strategy for the Construction of Insurance Tariff Classes.'' \emph{Scandinavian Actuarial Journal} 2018 (8): 681--705.
+
+\leavevmode\hypertarget{ref-hiltonConversationalProcesses1990}{}%
+Hilton, Denis J. 1990. ``Conversational Processes and Causal Explanation.'' \emph{Psychological Bulletin} 107 (1): 65.
+
+\leavevmode\hypertarget{ref-kimExamplesAre2016}{}%
+Kim, Been, Rajiv Khanna, and Oluwasanmi O.
+Koyejo.
+2016.
+``Examples Are Not Enough, Learn to Criticize!
+Criticism for Interpretability.'' In \emph{Advances in Neural Information Processing Systems}, 2280--8.
+
+\leavevmode\hypertarget{ref-kuoDeepTriangleDeep2018}{}%
+Kuo, Kevin. 2019. ``DeepTriangle: A Deep Learning Approach to Loss Reserving.'' \emph{Risks} 7 (3). Multidisciplinary Digital Publishing Institute: 97.
+
+\leavevmode\hypertarget{ref-lange_1966}{}%
+Lange, Jeffrey T. 1966. ``General Liability Insurance Ratemaking.'' \emph{Proceedings of the Casualty Actuarial Society} LIII. Arlington, Virginia: 26--53.
+
+\leavevmode\hypertarget{ref-lecunDeepLearning2015}{}%
+LeCun, Yann, Yoshua Bengio, and Geoffrey Hinton. 2015. ``Deep Learning.'' \emph{Nature} 521 (7553): 436.
+
+\leavevmode\hypertarget{ref-NIPS2017_7062}{}%
+Lundberg, Scott M, and Su-In Lee.
+2017.
+``A Unified Approach to Interpreting Model Predictions.'' In \emph{Advances in Neural Information Processing Systems 30}, edited by I.
+Guyon, U.
+V.
+Luxburg, S.
+Bengio, H.
+Wallach, R.
+Fergus, S.
+Vishwanathan, and R.
+Garnett, 4765--74.
+Curran Associates, Inc.
+\url{http://papers.nips.cc/paper/7062-a-unified-approach-to-interpreting-model-predictions.pdf}.
+
+\leavevmode\hypertarget{ref-mango_1998}{}%
+Mango, Donald F. 1998. ``An Application of Game Theory: Property Catastrophe Risk Load.'' In \emph{Proceedings of the Casualty Actuarial Society}, LXXXV:157--86.
+
+\leavevmode\hypertarget{ref-mildenhall_1999}{}%
+Mildenhall, Stephen J.
+1999.
+``A Systematic Relationship Between Minimum Bias and Generalized Linear Models.'' \emph{Proceedings of the Casualty Actuarial Society} LXXXVI.
+Arlington, Virginia: 393--487.
+
+\leavevmode\hypertarget{ref-millerExplanationArtificial2017}{}%
+Miller, Tim. 2017. ``Explanation in Artificial Intelligence: Insights from the Social Sciences.'' \emph{arXiv:1706.07269 {[}Cs{]}}, June. \url{http://arxiv.org/abs/1706.07269}.
+
+\leavevmode\hypertarget{ref-millerExplainableAI2017}{}%
+Miller, Tim, Piers Howe, and Liz Sonenberg.
+2017.
+``Explainable AI: Beware of Inmates Running the Asylum or: How I Learnt to Stop Worrying and Love the Social and Behavioural Sciences.'' \emph{arXiv Preprint arXiv:1712.00547}.
+
+\leavevmode\hypertarget{ref-molnar2018interpretable}{}%
+Molnar, Christoph, and others.
+2018.
+``Interpretable Machine Learning: A Guide for Making Black Box Models Explainable.'' \emph{E-Book At\textless{} Https://Christophm.
+Github.
+Io/Interpretable-Ml-Book/\textgreater{}, Version Dated} 10.
+
+\leavevmode\hypertarget{ref-naic_summer_2017}{}%
+National Association of Insurance Commissioners. 2017. ``2017 Proceedings of the National Association of Insurance Commissioners.''
+
+\leavevmode\hypertarget{ref-naic_white_paper}{}%
+---------.
+2019.
+``Regulatory Review of Predictive Models 10/15/19 Exposure Draft.'' \url{https://content.naic.org/sites/default/files/inline-files/Predictive\%20Model\%20White\%20Paper\%20Exposed\%2010-15-19.docx}.
+
+\leavevmode\hypertarget{ref-nelder_wedderburn_1972}{}%
+Nelder, J. A., and R. W. M. Wedderburn. 1972. ``Generalized Linear Models.'' \emph{Journal of the Royal Statistical Society}, Series a (general), 135 (3): 370--84.
+
+\leavevmode\hypertarget{ref-nilsson_2009}{}%
+Nilsson, Nils J. 2009. \emph{The Quest for Artificial Intelligence}. Cambridge University Press.
+
+\leavevmode\hypertarget{ref-noll_2018}{}%
+Noll, Alexander, Robert Salzmann, and Mario Wuthrich. 2018. ``Case Study: French Motor Third-Party Liability Claims.'' \emph{SSRN Electronic Journal}.
+
+\leavevmode\hypertarget{ref-rlang}{}%
+R Core Team. 2019. \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. \url{https://www.R-project.org/}.
+
+\leavevmode\hypertarget{ref-ribeiro2016should}{}%
+Ribeiro, Marco Tulio, Sameer Singh, and Carlos Guestrin.
+2016.
+``Why Should I Trust You?: Explaining the Predictions of Any Classifier.'' In \emph{Proceedings of the 22nd Acm Sigkdd International Conference on Knowledge Discovery and Data Mining}, 1135--44.
+ACM.
+
+\leavevmode\hypertarget{ref-robertson_2009}{}%
+Robertson, J.P. 2009. ``NCCI's 2007 Hazard Group Mapping.'' \emph{Variance} 3 (2): 194--213.
+
+\leavevmode\hypertarget{ref-roel_2018}{}%
+Roel, Verbelen, Katrien Antonio, and Gerda Claeskens. 2018. ``Unraveling the Predictive Power of Telematics Data in Car Insurance Pricing.'' \emph{Royal Statistical Society}.
+
+\leavevmode\hypertarget{ref-spedicato_2018}{}%
+Spedicato, Giorgio, Christophe Dutang, and Leonardo Petrini.
+2018.
+``Machine Learning Methods to Perform Pricing Optimization: A Comparison with Standard Generalized Linear Models.'' \emph{Variance} 12 (1).
+
+\leavevmode\hypertarget{ref-wang_raj_2017}{}%
+Wang, Haohan, and Bhiksha Raj. 2017. ``On the Origin of Deep Learning.'' \emph{arXiv:1702.07800v4 {[}cs.LG{]}}.
+
+\leavevmode\hypertarget{ref-wuthrich_2017}{}%
+Wuthrich, Mario V. 2017. ``Covariate Selection from Telematics Car Driving Data.'' \emph{European Actuarial Journal} 7 (1): 89--108.
+
+\leavevmode\hypertarget{ref-wuthrichMachineLearning2018}{}%
+Wüthrich, Mario V. 2018. ``Machine Learning in Individual Claims Reserving.'' \emph{Scandinavian Actuarial Journal} 2018 (6): 465--80.
+
+\leavevmode\hypertarget{ref-yang_2018}{}%
+Yang, Yi, Wei Qian, and Hui Zou. 2018. ``Insurance Premium Prediction via Gradient Tree-Boosted Tweedie Compound Poisson Models.'' \emph{Journal of Business and Economic Statistics} 36 (3): 456--70.
 
@@
 
-Appendix {.unnumbered}
-=============
+\hypertarget{appendix}{%
+\section*{Appendix}\label{appendix}}
+\addcontentsline{toc}{section}{Appendix}
 
-```{r variables, results='asis', echo = FALSE}
-library(magrittr)
-variables <- tibble::tribble(
-  ~Variable, ~Type, ~Transformation,
-  "Age range", "Categorical", "One-hot encode",
-  "Sex", "Categorical", "One-hot encode",
-  "Vehicle category", "Categorical", "One-hot encode",
-  "Make", "Categorical", "Embed in $\\mathbb{R}^2$",
-  "Vehicle age", "Numeric", "Center and scale",
-  "Region", "Categorical", "Embed in $\\mathbb{R}^2$"
-)
-knitr::kable(
-  variables,
-  caption = "Input variables and their transformations.",
-  booktabs = TRUE,
-  table.env = "table",
-  escape = FALSE,
-  linesep = ""
-) %>%
-  kableExtra::kable_styling(latex_options = "hold_position")
-```
+\begin{table}[!h]
 
-# Model Development {#model-dev}
+\caption{\label{tab:variables}Input variables and their transformations.}
+\centering
+\begin{tabular}[t]{lll}
+\toprule
+Variable & Type & Transformation\\
+\midrule
+Age range & Categorical & One-hot encode\\
+Sex & Categorical & One-hot encode\\
+Vehicle category & Categorical & One-hot encode\\
+Make & Categorical & Embed in $\mathbb{R}^2$\\
+Vehicle age & Numeric & Center and scale\\
+Region & Categorical & Embed in $\mathbb{R}^2$\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\hypertarget{model-dev}{%
+\section{Model Development}\label{model-dev}}
 
@@
 completeness and reproducibility, we include an overview of the process here.
-Implementation is done using the R [@rlang] interface to TensorFlow 
-[@tensorflow2015-whitepaper]. The model explanation visualizations utilize the 
-implemention by @biecekPMVEE, and the code to reproduce them are available on
-GitHub[^1].
+Implementation is done using the R (R Core Team 2019) interface to TensorFlow
+(Abadi et al. 2015). The model explanation visualizations utilize the
+implemention by Biecek and Burzykowski (2019), and the code to reproduce them are available on
+GitHub\footnote{https://github.com/kasaai/explain-ml-pricing}.
 
-[^1]: https://github.com/kasaai/explain-ml-pricing
+\hypertarget{data}{%
+\subsection{Data}\label{data}}
 
-## Data
-
-We use data from the AUTOSEG ("Automobile Statistics System") of Brazil's 
+We use data from the AUTOSEG (``Automobile Statistics System'') of Brazil's
 Superintendence of Private Insurance (SUSEP). The organization maintains
@@
 
-## Model
+\hypertarget{model}{%
+\subsection{Model}\label{model}}
 
-Table \@ref(tab:variables) shows the input variables to our model and their 
-associated transformations. For "make" and "region", we map each level to a 
-point in $\mathbb{R}^2$ through embedding layers. The model predicts expected
+Table \ref{tab:variables} shows the input variables to our model and their
+associated transformations. For ``make'' and ``region'', we map each level to a
+point in \(\mathbb{R}^2\) through embedding layers. The model predicts expected
 loss cost for all perils combined. The architecture is a feedforward
@@
 does not improve for five epochs.
+
+\end{document}

### 2020-03-05 manuscript/manuscript.Rmd
single column

@@
     base_format: rticles::elsevier_article
-layout: preprint, 3p, twocolumn, letterpaper, 10pt
+layout: preprint, 3p, letterpaper, 10pt
 ---

### 2020-03-05 manuscript/manuscript.Rmd
redactions

@@
 author:
-  - name: Kevin Kuo
-    email: kevin@kasa.ai
+  - name: (redacted)
+    email: (redacted)
     affiliation: a
     footnote: Corresponding Author
-  - name: Daniel Lupton
-    email: dlupton@taylorandmulder.com
+  - name: (redacted)
+    email: (redacted)
     affiliation: b
@@
   - code: a
-    address: Kasa AI, 3040 78th Ave SE \#1271, Mercer Island, WA 98040, USA
+    address: (redacted)
   - code: b
-    address: Taylor & Mulder, 10508 Rivers Bend Lane, Potomac, MD 20854, USA
+    address: (redacted)
 abstract: |
@@
 
-Acknowledgments {.unnumbered}
-==========
-
-We thank Navdeep Gill, Daniel Falbel, Morgan Bugbee, the volunteers of the CAS
-project oversight group, and two anonymous reviewers for helpful discussions
-and/or feedback.
-This work is supported by the Casualty Actuarial Society.
-
 References {#references .unnumbered}
@@
 
-[^1]: https://github.com/kasaai/explain-ml-pricing
+[^1]: (redacted)
 
